---
title: "ICI_culprit_master"
author: "Eric Mukherjee"
date: "2025-06-13"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# INTRODUCTION

This is a unified document to generate the figures for Manuscript 2 -
ICI-culprit interactions. It will be organized by figure.

## Libraries

```{r}
# 📦 Data manipulation and wrangling
library(data.table)
library(dplyr)
library(tidyr)
library(purrr)
library(stringr)
library(readr)

# 📊 Statistical modeling and regression tools
library(stats)
library(broom)
library(car)
library(survival)
library(survminer)

# 📈 Visualization
library(ggplot2)
library(ggpubr)
library(patchwork)
library(scales)
library(RColorBrewer)
library(ggpp)
library(gridExtra)

# 📋 Tables and formatting
library(gt)
library(glue)
library(gsubfn)

# 🌍 Metadata and classification
library(countrycode)
library(reshape2)  # Used for reshaping data (wide ↔ long)
```

## Useful functions

### Create logit summary table

Create a summary table of any logistic regression for publication by
using this function. We choose to create Wald-based confidence
intervals, derived from the standard errors of the model coefficients,
rather that using profile-likelihood-based CIs (which take a while)

```{r}
# This function creates a summary table for logistic regression
# It includes the option for bootstrap confidence intervals and p-values
logit_summary_table <- function(model, data, label_map = NULL) {
  coef_table <- summary(model)$coefficients
  or_table <- exp(cbind(OR = coef(model), confint.default(model)))
  pvals <- coef_table[, 4]

  result_rows <- list()

  predictors <- attr(terms(model), "term.labels")
  response <- all.vars(formula(model))[1]

  for (term in predictors) {
    if (!(term %in% names(data))) next
    var <- data[[term]]
    display_term <- if (!is.null(label_map) && term %in% names(label_map)) label_map[[term]] else term

    # Binary 0/1 or logical
    if (length(unique(na.omit(var))) == 2 && (is.numeric(var) || is.logical(var))) {
      var <- factor(as.integer(var), levels = c(0, 1))
      counts <- table(var)
      sjsten_counts <- table(var[data[[response]] == 1])

      result_rows[[length(result_rows) + 1]] <- data.frame(
        Variable = display_term,
        Level = "0 (ref)",
        N = counts["0"],
        SJS_TEN_Cases = sjsten_counts["0"],
        OR = 1,
        CI = "-",
        P = "-",
        stringsAsFactors = FALSE
      )

      coef_name <- paste0(term, "1")
      i <- grep(paste0("^", coef_name, "$"), rownames(coef_table), fixed = TRUE)
      if (length(i) > 0) {
        or <- round(or_table[i, 1], 2)
        ci <- sprintf("[%.2f–%.2f]", or_table[i, 2], or_table[i, 3])
        p <- signif(pvals[i], 3)
        if (p < 0.001) p <- "<0.001"

        result_rows[[length(result_rows) + 1]] <- data.frame(
          Variable = display_term,
          Level = "1",
          N = counts["1"],
          SJS_TEN_Cases = sjsten_counts["1"],
          OR = or,
          CI = ci,
          P = p,
          stringsAsFactors = FALSE
        )
      }

    # Multilevel factor
    } else if (is.factor(var)) {
      levels_var <- levels(var)
      base <- levels_var[1]
      counts <- table(var)
      sjsten_counts <- table(var[data[[response]] == 1])

      result_rows[[length(result_rows) + 1]] <- data.frame(
        Variable = display_term,
        Level = paste0(base, " (ref)"),
        N = counts[base],
        SJS_TEN_Cases = sjsten_counts[base],
        OR = 1,
        CI = "-",
        P = "-",
        stringsAsFactors = FALSE
      )

      for (lvl in levels_var[-1]) {
        coef_name <- paste0(term, lvl)
        i <- grep(coef_name, rownames(coef_table), fixed = TRUE)
        if (length(i) == 0) next
        or <- round(or_table[i, 1], 2)
        ci <- sprintf("[%.2f–%.2f]", or_table[i, 2], or_table[i, 3])
        p <- signif(pvals[i], 3)
        if (p < 0.001) p <- "<0.001"

        result_rows[[length(result_rows) + 1]] <- data.frame(
          Variable = display_term,
          Level = lvl,
          N = counts[lvl],
          SJS_TEN_Cases = sjsten_counts[lvl],
          OR = or,
          CI = ci,
          P = p,
          stringsAsFactors = FALSE
        )
      }

    # Continuous numeric
    } else if (is.numeric(var)) {
      i <- grep(paste0("^", term, "$"), rownames(coef_table))
      if (length(i) == 0) next
      or <- round(or_table[i, 1], 2)
      ci <- sprintf("[%.2f–%.2f]", or_table[i, 2], or_table[i, 3])
      p <- signif(pvals[i], 3)
      if (p < 0.001) p <- "<0.001"

      result_rows[[length(result_rows) + 1]] <- data.frame(
        Variable = display_term,
        Level = "Per Unit Increase",
        N = nrow(data),
        SJS_TEN_Cases = sum(data[[response]] == 1, na.rm = TRUE),
        OR = or,
        CI = ci,
        P = p,
        stringsAsFactors = FALSE
      )
    }
  }

  result_df <- do.call(rbind, result_rows)
  rownames(result_df) <- NULL
  return(result_df)
}
```

### Change CIs to scientific notation if they have too many digits

```{r}
# Helper function to format each number in CI strings
format_ci <- function(ci_string) {
  nums <- str_extract_all(ci_string, "\\d+\\.?\\d*")[[1]]
  formatted <- map_chr(nums, ~ {
    x <- as.numeric(.x)
    if (!is.na(x) && x > 999) formatC(x, format = "e", digits = 2) else formatC(x, format = "f", digits = 2)
  })
  str_replace(ci_string, "\\[.*\\]", paste0("[", paste(formatted, collapse = "-"), "]"))
}
```

### Create summary table for survival models

Similarly, this function creates summary tables for Cox survival models,
using the model outputs directly.

```{r}
# This function creates a summary table for Cox models
cox_summary_table <- function(model, data) {
  # Extract model info
  coef_table <- summary(model)$coefficients
  ci_table <- summary(model)$conf.int
  pvals <- coef_table[, "Pr(>|z|)"]

  predictors <- attr(terms(model), "term.labels")
  result_rows <- list()

  for (term in predictors) {
    var <- data[[term]]

    if (is.factor(var)) {
      levels_var <- levels(var)
      base <- levels_var[1]
      counts <- table(var)

      result_rows[[length(result_rows) + 1]] <- data.frame(
        Variable = term,
        Level = paste0(base, " (ref)"),
        N = as.character(counts[base]),
        HR = 1,
        CI = "-",
        P = "-",
        stringsAsFactors = FALSE
      )

      for (lvl in levels_var[-1]) {
        term_name <- paste0(term, lvl)
        i <- grep(term_name, rownames(coef_table), fixed = TRUE)
        if (length(i) == 0) next

        or <- round(ci_table[i, "exp(coef)"], 2)
        ci <- sprintf("[%.2f-%.2f]", ci_table[i, "lower .95"], ci_table[i, "upper .95"])
        p <- signif(pvals[i], 3)
        if (!is.na(p) && p < 0.001) p <- "<0.001"

        result_rows[[length(result_rows) + 1]] <- data.frame(
          Variable = term,
          Level = lvl,
          N = as.character(counts[lvl]),
          HR = or,
          CI = ci,
          P = ifelse(is.na(p), "-", as.character(p)),
          stringsAsFactors = FALSE
        )
      }
    }

    else if (is.numeric(var)) {
      i <- which(rownames(coef_table) == term)
      if (length(i) == 0) next

      or <- round(ci_table[i, "exp(coef)"], 2)
      ci <- sprintf("[%.2f-%.2f]", ci_table[i, "lower .95"], ci_table[i, "upper .95"])
      p <- signif(pvals[i], 3)
      if (!is.na(p) && p < 0.001) p <- "<0.001"

      result_rows[[length(result_rows) + 1]] <- data.frame(
        Variable = term,
        Level = "Per Unit Increase",
        N = "-",
        HR = or,
        CI = ci,
        P = ifelse(is.na(p), "-", as.character(p)),
        stringsAsFactors = FALSE
      )
    }
  }

  result_df <- bind_rows(result_rows)
  rownames(result_df) <- NULL
  return(result_df)
}
```


# ASSEMBLING final_dt

## Load files

First, load our sanitized files in the workspace.

```{r}
# Import Drug and Outcome Data
standard_case_drug <- fread("sanitized_drug_data.csv")
standard_case_outcome <- fread("sanitized_outcome_data.csv")
demographics <- fread("sanitized_demographics.csv")
sanitized_ther <- fread("sanitized_ther.csv")
standard_case_outcome_category <- fread("standard_case_outcome_category.csv")

# Dictionaries as well
outcome_dictionary <- fread("sanitized_outcome_dictionary.csv")
drug_dictionary <- fread("Updated_Drug_Dictionary_Fullnames.csv")

# Clean up memory
gc()
```

## Subsetting data to only the years of interest

We'll subset the data to 2013-2023, because ICIs were approved in 2011.

```{r}
# Convert the filename column into full year format
demographics[, year := as.numeric(substr(filename, 5, 6)) + 2000]  # Extract YY, convert to YYYY

# Filter dataset to only include years 2013-2023
demographics <- demographics[year >= 2013 & year <= 2023]

# Print confirmation
cat("✅ Subset complete: Data now includes only years 2013-2023!\n")

# Clean up memory
gc()
```

Subset the drug and outcome information by years as well.

```{r}
# Step 1: Extract the list of composite IDs from sanitized demographics
valid_composite_ids <- unique(demographics$compositeid)

# Step 2: Subset updated_drug_data and updated_outcome_data using these IDs
standard_case_drug <- standard_case_drug[compositeid %in% valid_composite_ids]
standard_case_outcome <- standard_case_outcome[compositeid %in% valid_composite_ids]
sanitized_ther <- sanitized_ther[compositeid %in% valid_composite_ids]

# Print confirmation
cat("✅ Data successfully filtered: Only patients from 2013-2023 are included!\n")
cat("Total records in standard_case_drug:", nrow(standard_case_drug), "\n")
cat("Total records in standard_case_outcome:", nrow(standard_case_outcome), "\n")
cat("total records in sanitized_ther:", nrow(sanitized_ther), "\n")

# Clean up memory
gc()
```

## Initialize final_dt

The data structure we'll use is simple - we'll create one row for each
compositeid present in standard_case_drug and/or standard_case_outcome,
and then use a series of variables (mostly booleans) for age, sex,
cancer, ICI, each specific ICI, culprit drug, each specific culprit
drug, death, and cancer. Each of these will have to be added manually.

First, initialize the data.table.

```{r}
# Get All Unique Composite IDs
all_patient_ids <- unique(c(standard_case_drug$compositeid, standard_case_outcome$compositeid))

# Initialize Final Data Table
final_dt <- data.table(compositeid = all_patient_ids)

# Clean up memory
gc()
```

Then add demographic information (including country information)

```{r}
# Merge age, sex, country, and region from Demographics
demographics_dt <- as.data.table(demographics)[, .(compositeid, age_yr, sex, country)]

# Fill missing or empty 'country' values with "UNKNOWN"
demographics_dt[is.na(country) | trimws(country) == "", country := "UNKNOWN"]

# Replace specific unknown values with NA
demographics_dt[country %in% c("COUNTRY NOT SPECIFIED", "UNKNOWN"), country := NA_character_]

# Map countries to regions using 'countrycode'
demographics_dt[, region := countrycode(country, origin = 'iso2c', destination = 'region')]

# Manually assign region for specific country codes
demographics_dt[country == "RE", region := "Sub-Saharan Africa"]

# Replace any remaining NA values in 'region' with "Unknown"
demographics_dt[is.na(region), region := "Unknown"]

# Convert 'country' and 'region' columns to factors
demographics_dt[, `:=`(
  country = as.factor(country),
  region = as.factor(region)
)]

# Set 'Unknown' as the reference level
demographics_dt[, region := relevel(region, ref = "Unknown")]

# Merge it into the final_dt
final_dt <- merge(final_dt, demographics_dt, by = "compositeid", all.x = TRUE)

head(final_dt)

# Clean up memory
gc()
```

Now add the concomitant drug count

```{r}
# Add Total Drug Count
num_drugs <- standard_case_drug[, .(NumDrugs = .N), by = compositeid]
final_dt <- merge(final_dt, num_drugs, by = "compositeid", all.x = TRUE)
final_dt[, NumDrugs := fifelse(is.na(NumDrugs), 0, NumDrugs)]

print(final_dt)

# Clean up memory
gc()
```

We next flag the ICI and culprit rows.

-   For strong culprits, we'll pick: lamotrigine (705103), TMP/SMX
    (40220482), phenytoin (740910), carbamazepine (740275), and
    allopurinol (1167322). There's also sulfamethoxazole (1836430) and
    trimethoprim (1705674).
-   For weak culprits, we'll pick: azithromycin (1734104), ciprofloxacin
    (1797513), oxcarbazepine (718122), sulfasalazine (964339), acyclovir
    (1703687).
-   For ICI, we'll pick: pembrolizumab (45775965), nivolumab (45892628),
    ipilimumab (40238188), atezolizumab (42629079), durvalumab
    (1594034), cemiplimab (35200783), tremelimumab (741851), dostarlimab
    (1536789).

```{r}
# Define ICI and Culprit Drug Lists
ici_list <- c(
  "Pembrolizumab" = 45775965, "Nivolumab" = 45892628, "Ipilimumab" = 40238188, 
  "Atezolizumab" = 42629079, "Durvalumab" = 1594034, "Cemiplimab" = 35200783, 
  "Tremelimumab" = 741851, "Dostarlimab" = 1536789, "Avelumab" = 1593273
)

strong_culprit_list <- list(
  "Lamotrigine" = c(705103),
  "TMP_SMX" = c(40220482), 
  "Phenytoin" = c(740910, 40238920, 40070682), # Some of these are combination drugs
  "Carbamazepine" = c(740275),
  "Allopurinol" = c(1167322, 40005864, 40005862) # Some of these are combination drugs
)

weak_culprit_list <- c(
  "Azithromycin" = 1734104, "Ciprofloxacin" = 1797513, "Oxcarbazepine" = 718122, 
  "Sulfasalazine" = 964339, "Acyclovir" = 1703687
)

# Identify Patients Exposed to Each ICI & Culprit Drug
ici_cases <- standard_case_drug %>% filter(drug_concept_id %in% ici_list)
strong_culprit_cases <- standard_case_drug %>% filter(drug_concept_id %in% unlist(strong_culprit_list))
weak_culprit_cases <- standard_case_drug %>% filter(drug_concept_id %in% unlist(weak_culprit_list))

# Clean up memory
gc()
```

Now use these case subsets to add additional flags.

```{r}
# Initialize ICI Flags
for (ici_name in names(ici_list)) {
  final_dt[, (paste0("ICI_", ici_name)) := 0]
}

# Populate ICI Flags
for (ici_name in names(ici_list)) {
  ici_id <- ici_list[[ici_name]]
  exposed_patients <- unique(ici_cases$compositeid[ici_cases$drug_concept_id == ici_id])
  final_dt[compositeid %in% exposed_patients, (paste0("ICI_", ici_name)) := 1]
}

# Initialize Strong & Weak Culprit Flags
for (culprit_name in names(strong_culprit_list)) {
  final_dt[, (paste0("Strong_Culprit_", culprit_name)) := 0]
}

for (culprit_name in names(weak_culprit_list)) {
  final_dt[, (paste0("Weak_Culprit_", culprit_name)) := 0]
}

# Populate Strong Culprit Flags (with OR logic across concept IDs)
for (culprit_name in names(strong_culprit_list)) {
  culprit_ids <- strong_culprit_list[[culprit_name]]  # vector of one or more concept IDs
  exposed_patients <- unique(
    standard_case_drug[drug_concept_id %in% culprit_ids, compositeid]
  )
  final_dt[compositeid %in% exposed_patients, 
           (paste0("Strong_Culprit_", culprit_name)) := 1]
}

# Populate Weak Culprit Flags (with OR logic across concept IDs)
for (culprit_name in names(weak_culprit_list)) {
  culprit_ids <- weak_culprit_list[[culprit_name]]  # vector of one or more concept IDs
  exposed_patients <- unique(
    standard_case_drug[drug_concept_id %in% culprit_ids, compositeid]
  )
  final_dt[compositeid %in% exposed_patients, 
           (paste0("Weak_Culprit_", culprit_name)) := 1]
}

## In addition to patients with the TMP-SMX injection code, we'll also add the patients who have individual codes for TMP or SMX individually
# TMP-SMX Detection (Fix OR Logic)
tmp_patients <- unique(standard_case_drug[drug_concept_id == 1705674, compositeid])
smx_patients <- unique(standard_case_drug[drug_concept_id == 1836430, compositeid])
tmp_or_smx_patients <- union(tmp_patients, smx_patients)
pre_combined_patients <- unique(standard_case_drug[drug_concept_id == 40220482, compositeid])
tmp_smx_patients <- unique(c(tmp_or_smx_patients, pre_combined_patients))

# Assign TMP-SMX flag
final_dt[, TMP_SMX := 0]
final_dt[compositeid %in% tmp_smx_patients, TMP_SMX := 1]

# Ensure TMP-SMX is included in StrongCulprit_TMP_SMX
final_dt[, Strong_Culprit_TMP_SMX := pmax(Strong_Culprit_TMP_SMX, TMP_SMX, na.rm = TRUE)]

# Remove TMP_SMX column after assigning it to StrongCulprit_TMP_SMX
final_dt[, TMP_SMX := NULL]

##Now add composite flags for groups of things
# Populate composite flags
final_dt[, ICI := as.integer(rowSums(.SD) > 0), .SDcols = paste0("ICI_", names(ici_list))]
final_dt[, StrongCulprit := as.integer(rowSums(.SD) > 0), .SDcols = paste0("Strong_Culprit_", names(strong_culprit_list))]
final_dt[, WeakCulprit := as.integer(rowSums(.SD) > 0), .SDcols = paste0("Weak_Culprit_", names(weak_culprit_list))]
final_dt[, AllCulprit := as.integer(rowSums(.SD) > 0), 
         .SDcols = c(paste0("Strong_Culprit_", names(strong_culprit_list)), 
                     paste0("Weak_Culprit_", names(weak_culprit_list)))]

# Replace NA with 0 for All Flags
flag_columns <- c("ICI", "StrongCulprit", "WeakCulprit", "AllCulprit",
                  paste0("ICI_", names(ici_list)), 
                  paste0("Strong_Culprit_", names(strong_culprit_list)), 
                  paste0("Weak_Culprit_", names(weak_culprit_list)))

final_dt[, (flag_columns) := lapply(.SD, function(x) ifelse(is.na(x), 0, x)), .SDcols = flag_columns]

# Preview
print(final_dt)
summary(final_dt)

# Clean up memory
gc()
```

Then add flags for specific ICI classes.

```{r}
# Assign PD-1, PD-L1, and CTLA-4 Flags Using OR Logic
pd1_list <- c("Pembrolizumab" = 45775965, "Nivolumab" = 45892628, "Cemiplimab" = 35200783, "Dostarlimab" = 1536789)
pdl1_list <- c("Atezolizumab" = 42629079, "Durvalumab" = 1594034, "Avelumab" = 1593273)
ctla4_list <- c("Ipilimumab" = 40238188, "Tremelimumab" = 741851)

final_dt[, ICI_PD1 := as.integer(rowSums(.SD) > 0), .SDcols = paste0("ICI_", names(pd1_list))]
final_dt[, ICI_PDL1 := as.integer(rowSums(.SD) > 0), .SDcols = paste0("ICI_", names(pdl1_list))]
final_dt[, ICI_CTLA4 := as.integer(rowSums(.SD) > 0), .SDcols = paste0("ICI_", names(ctla4_list))]

# Clean up memory
gc()
```

Now for something a little fancier - add flags for being exposed to more
than one ICI, and one for the number of ICI a patient is exposed to.

```{r}
# Identify Patients Exposed to More Than One ICI (Dual ICI Flag)
final_dt[, Dual_ICI := as.integer(rowSums(.SD) >= 2), .SDcols = paste0("ICI_", names(ici_list))]

# Add Column for Number of ICIs a patient is exposed to
final_dt[, ICI_Count := rowSums(.SD), .SDcols = paste0("ICI_", names(ici_list))]

# Categorize ICI_Count into three levels: 'Zero', 'One', and 'Multiple'
final_dt[, ICI_Count_Factor := fifelse(ICI_Count == 0, "Zero",
                         fifelse(ICI_Count == 1, "One", "Multiple"))]

# Convert ICI_Count to a factor with 'Zero' as the reference level
final_dt[, ICI_Count_Factor := factor(ICI_Count_Factor, levels = c("Zero", "One", "Multiple"))]

# Clean up memory
gc()
```

Add a column for the primary cause (PS).

```{r}
# Subset by PS 
standard_case_drug_PS <- standard_case_drug[role_cod=="PS"]

# Merge with final_dt
final_dt <- merge(final_dt, standard_case_drug_PS, by = "compositeid", all.x = TRUE)

# Filter down by columns
final_dt$drug_seq <- NULL
final_dt$role_cod <- NULL

# Clean up memory
gc()
```

## Deal with combination drugs

We've dealt with a similar issue in the exposure matrix above, but this
time we have to consider the primary suspects. A big issue with
combination drugs is the fact that the only combination code for TMP /
SMX is an injection (it's likely people reported this when they just
gave oral bactrim, because there doesn't seem to be a drug_concept_id
for oral bactrim). We will consider a patient having a primary suspect
of TMP/SMX if both TMP and SMX are coded separately and one is coded as
the primary cause (this is in contrast with TMP/SMX exposure, in which
exposure to either TMP or SMX is considered exposure to TMP/SMX).

```{r}
## In addition to patients with the TMP-SMX injection code, we'll also re-code the primary suspect for those patients whose primary suspect is TMP or SMX - TMP is almost never used alone.

# TMP-SMX Detection (Fix & OR Logic)
tmp_patients_ps <- unique(standard_case_drug_PS[drug_concept_id == 1705674, compositeid])
tmp_patients <- unique(standard_case_drug[drug_concept_id == 1705674, compositeid])

smx_patients_ps <- unique(standard_case_drug_PS[drug_concept_id == 1836430, compositeid])
smx_patients <- unique(standard_case_drug[drug_concept_id == 1836430, compositeid])


# Find the patients who have one of TMP or SMX as a primary cause, and the other one in the drug list
tmp_smx_combined_patients_ps <- union(intersect(tmp_patients_ps, smx_patients), intersect(tmp_patients, smx_patients_ps))

# Find the list of patients for whom TMP / SMX is the primary suspect (those with TMP/SMX already coded as primary, plus those with either TMP or SMX as primary)
pre_combined_patients_ps <- unique(standard_case_drug_PS[drug_concept_id == 40220482, compositeid])
tmp_smx_patients_ps <- unique(c(tmp_smx_combined_patients_ps, pre_combined_patients_ps))

# ✅ Replace the primary suspect drug for identified TMP-SMX patients
final_dt[compositeid %in% tmp_smx_patients_ps, drug_name := "TMP / SMX"]
final_dt[compositeid %in% tmp_smx_patients_ps, drug_concept_id := 40220482]

# Clean up variables
rm(tmp_patients, smx_patients, tmp_smx_patients, pre_combined_patients)
```

##Sanitize age and sex variables

Fix the age_yr and sex variables - the age_yr will be binned in 20-year
increments, sex variable will be coded as F, M, and NS.

```{r}
# Sanitize age_yr and sex variables in final_dt
final_dt <- final_dt %>%
  mutate(
    sex = ifelse(sex %in% c("M", "F"), sex, "NS")  # Replace unknown values with "NS"
  )

# Add age_group variable
# Preferred syntax to avoid shallow copy warnings
setDT(final_dt)
set(final_dt, j = "age_group", value = cut(
  final_dt$age_yr,
  breaks = c(-Inf, 20, 40, 60, 80, Inf),
  labels = c("0-20", "21-40", "41-60", "61-80", "80+"),
  right = TRUE
))

# Replace NA with "Unknown"
final_dt[is.na(age_group), age_group := "Unknown"]

# Convert to Proper Data Types (Ensure Categorical Variables Are Factors)
final_dt <- final_dt %>%
  mutate(sex = as.factor(sex),  # Ensure sex is a factor
    age_group = as.factor(age_group)  # Ensure outcome variable is a factor
  )

# Clean up memory
gc()
```

##Add mortality information

We'll need a function to combine isr and primaryid into one type of id -
the compositeid. For most of the files, this has already been done, but
it hasn't been for the file with mortality information
(standard_case_outcome_category).

```{r}
# Function to create compositeids based on primaryid and isr, with NA handling
create_compositeid <- function(data) {
    # Ensure data is a data.table
    data <- as.data.table(data)
    
    # Check if both primaryid and isr columns exist
    if ("primaryid" %in% names(data) & "isr" %in% names(data)) {
        # Create compositeid using primaryid if it's not NA, otherwise use isr
        data[, compositeid := ifelse(!is.na(primaryid), paste0("pid.", primaryid), paste0("isr.", isr))]
    
    # If only primaryid exists
    } else if ("primaryid" %in% names(data)) {
        data[, compositeid := paste0("pid.", primaryid)]
    
    # If only isr exists
    } else if ("isr" %in% names(data)) {
        data[, compositeid := paste0("isr.", isr)]
    
    # If neither column exists, throw an error
    } else {
        stop("Neither 'primaryid' nor 'isr' columns found in the data.")
    }
    
    # Return the updated data table
    return(data)
}

# Clean up memory
gc()
```

Then apply this function to our new table:

```{r}
# Add a compositeid and remove the other columns
standard_case_outcome_category <- create_compositeid(standard_case_outcome_category)
standard_case_outcome_category$primaryid <- NULL
standard_case_outcome_category$isr <- NULL
standard_case_outcome_category$snomed_concept_id <- NULL

# Add mortality information to final_dt
final_dt <- merge(final_dt, standard_case_outcome_category, by = "compositeid", all.x = TRUE)

# Determine if any 'DE' exists per compositeid
final_dt[, Death := if (any(outc_code == "DE", na.rm = TRUE)) {
  1L
} else if (all(is.na(outc_code))) {
  NA_integer_
} else {
  0L
}, by = compositeid]

# Remove duplicate rows while keeping only one row per compositeid
final_dt <- unique(final_dt, by = "compositeid")

# Print the cleaned dataset
print(final_dt)
summary(final_dt)

# Clean up memory
gc()
```

## Add biologic vs small molecule classification

Add a column classifying whether the causative drug is a biologic or not
(small molecule). Biologics will be those drugs that contain a word
ending with "mab" or "cept".

```{r}
# Classify drugs as biologics or small molecules
final_dt <- final_dt %>%
  mutate(drug_class = ifelse(grepl("mab$|cept$", drug_name, ignore.case = TRUE),
                             "Biologic", "Small Molecule"))

# Clean up memory
gc()
```

##Add SJS/TEN flag

Then add a flag for SJS/TEN cases.

```{r}
# Flag SJS/TEN Cases
SJSTEN_IDs <- c(36009724, 43010942, 36009754)
sjs_ten_cases <- standard_case_outcome %>%
  filter(outcome_concept_id %in% SJSTEN_IDs) %>%
  select(compositeid) %>%
  distinct() %>%
  mutate(SJSTEN = 1)

final_dt <- merge(final_dt, sjs_ten_cases, by = "compositeid", all.x = TRUE)
final_dt[, SJSTEN := fifelse(is.na(SJSTEN), 0, SJSTEN)]

# Convert SJSTEN variable to factors
final_dt <- final_dt %>%
  mutate(SJSTEN = as.factor(SJSTEN))

print(final_dt)
summary(final_dt)

# Clean up memory
gc()
```

##Add TTE information

Next, we can left joint combined_ther to the standard_case_drug table to
get the dates we needed for TTE analysis.

```{r}
# Perform a left join using base R, ensuring that non-matching rows get NA
drug_exposure <- merge(standard_case_drug_PS, sanitized_ther, 
                       by = c("compositeid", "drug_seq"), 
                       all.x = TRUE)

# View the result
print(drug_exposure)
```

The patient demographics table has the start date of the given reaction.
Now we can add that data to the outcome_data.

```{r}
# Perform a left join to add the demographic columns to outcome data
# Use merge with all.x = TRUE to preserve all rows of outcome data
updated_outcome_data <- merge(standard_case_outcome, demographics, by = "compositeid", all.x = TRUE)

# View the updated data to verify the join
print(updated_outcome_data)
```

We'll need the event_dt from the updated_outcome_data, and the start_dt
from drug_exposure. We can first subset both of these to just SJS/TEN.

```{r}
# Subset outcome data to just SJS/TEN
outcome_data_sjsten <- as.data.table(updated_outcome_data)[, .(compositeid, event_dt, fda_dt, rept_dt)]
outcome_data_sjsten <- outcome_data_sjsten[compositeid %in% sjs_ten_cases$compositeid]

# Subset drug_exposure
exposure_data_sjsten <- drug_exposure[compositeid %in% sjs_ten_cases$compositeid]

# Merge these two
drug_TTE_sjsten <- merge(outcome_data_sjsten, exposure_data_sjsten, by = "compositeid", all = TRUE)
```

Then, we can calculate the time-to-event (TTE) from these numbers.

```{r}
# Function to check if date is complete (YYYYMMDD format)
is_full_date <- function(date) {
  nchar(as.character(date)) == 8  # Only keep dates with 8 characters
}

# Subset rows with complete dates
drug_TTE_sjsten <- drug_TTE_sjsten[
  is_full_date(start_dt) & is_full_date(event_dt)  # Filter for full dates only
]

# Convert to Date format
drug_TTE_sjsten[, start_dt := as.Date(as.character(start_dt), format = "%Y%m%d")]
drug_TTE_sjsten[, event_dt := as.Date(as.character(event_dt), format = "%Y%m%d")]

# Calculate TTE for each row (Time-to-Event in days)
drug_TTE_sjsten[, TTE := as.numeric(event_dt - start_dt)]

# Filter TTE between 0 and 180 days
drug_TTE_sjsten <- drug_TTE_sjsten %>% filter(TTE > 0 & TTE <= 180)

# Keep row with lowest TTE per compositeid
drug_TTE_sjsten <- drug_TTE_sjsten %>%
  group_by(compositeid) %>%
  slice_min(order_by = TTE, with_ties = FALSE) %>%
  ungroup()
```

Now, we have the TTE information for each drug. We also need to fix the
TMP/SMX rows, like we did with final_dt.

```{r}
# Ensure both are data.tables
setDT(drug_TTE_sjsten)
setDT(final_dt)

# Create a clean mapping table (compositeid -> updated drug fields)
final_updates <- unique(final_dt[, .(compositeid, drug_name, drug_concept_id)])

# Rename to avoid conflict in merge
setnames(final_updates, c("drug_name", "drug_concept_id"), c("updated_drug_name", "updated_drug_concept_id"))

# Perform left join
drug_TTE_sjsten <- merge(
  drug_TTE_sjsten,
  final_updates,
  by = "compositeid",
  all.x = TRUE,
  suffixes = c("", ".final")
)

# Apply updates only if non-NA
drug_TTE_sjsten[!is.na(updated_drug_name), drug_name := updated_drug_name]
drug_TTE_sjsten[!is.na(updated_drug_concept_id), drug_concept_id := updated_drug_concept_id]

# Remove the temporary columns
drug_TTE_sjsten[, c("updated_drug_name", "updated_drug_concept_id") := NULL]
```

## Add cancer information

The next column to add is a boolean for the presence or absence of a
cancer diagnosis.

In order to filter this down to just cancer patients, we need to
distinguish cancer from non-cancer. We will do this with keywords.

```{r}
# Load your indications file
indications_index_dt <- fread("table_of_indications.csv")
indications_index_dt[, indication_clean := toupper(trimws(indication))]

cancer_keywords <- c(
  # General terms
  "CANCER", "CARCINOMA", "MALIGNANT", "MALIGNANCY", "NEOPLASM", "TUMOR", "TUMOUR", "METASTASIS", "METASTASES", "METASTATIC", "BLASTOMA",
  "PARANEOPLASTIC", "ONCOLOGIC COMPLICATION", "CHEMOTHERAPY", "NEOADJUVANT THERAPY", "ADJUVANT THERAPY", "METAPLASIA", "\\bMASS\\b", "RADIOIMMUNOTHERAPY", "\\bPAPILLOMA\\b",
  
  # Skin cancers
  "MELANOMA", "LENTIGO MALIGNA", "MYCOSIS FUNGOIDES", "SEZARY SYNDROME", 
  
  # Hematologic malignancies
  "LYMPHOMA", "MYELOMA", "LEUKEMIA", "LEUKAEMIA", "MYELODYSPLASTIC", "MYELOPROLIFERATIVE",
  "POLYCYTHEMIA VERA", "POLYCYTHAEMIA VERA", "ESSENTIAL THROMBOCYTHEMIA", "ESSENTIAL THROMBOCYTHAEMIA", "HODGKIN",
  "ESSENTIAL THROMBOCYTOSIS", "MYELOFIBROSIS", "MACROGLOBULINEMIA", "MACROGLOBULINAEMIA",
  "POST TRANSPLANT LYMPHOPROLIFERATIVE DISORDER", "BLAST CELL CRISIS", "CHLOROMA", "MASTOCYTOSIS",

  # Brain and CNS
  "GLIOMA", "GLIOBLASTOMA", "ASTROCYTOMA", "EPENDYMOMA", "MENINGIOMA", "CRANIOPHARYNGIOMA", "CHORDOMA", "GERMINOMA", "HEMANGIOPERICYTOMA",

  # Soft tissue and bone
  "SARCOMA", "FIBROMA", "SCHWANNOMA", "BOWEN'S DISEASE", "RETINAL MELANOCYTOMA", "CHONDROMA", "LIPOMA",

  # GI and endocrine tumors
  "INSULINOMA", "LINITS PLASTICA", "CHROMOCYTOMA", "CARCINOID",

  # Other specific types
  "TERATOMA", "ADENOMA", "MESOTHELIOMA", "BONE MARROW INFILTRATION", "THYMOMA",
  "RICHTER'S SYNDROME",  "PAGET'S DISEASE OF THE VULVA", "PAGET'S DISEASE OF THE PENIS", "EXTRAMAMMARY PAGET'S DISEASE")


# Build audit table separately
# This will tell us which indications were mapped as cancer
audit_dt <- copy(indications_index_dt)[, .(indication, indication_concept_id, indication_clean)]

# Build keyword flags
for (kw in cancer_keywords) {
  audit_dt[, paste0("keyword_", kw) := str_detect(indication_clean, kw)]
}

# Build exclusion flags
audit_dt[, exclude_flag := str_detect(indication_clean, "HYPERTENSION MALIGNANT|HYPERTHERMIA MALIGNANT|BODY MASS|VACCINATION SITE MASS|BONE MASS DECREASED|HUMAN PAPILLOMA VIRUS|PAPILLOMA VIRAL|CHONDROMALACIA")]
audit_dt[, tumor_nf_flag := str_detect(indication_clean, "TUMOR NECROSIS FACTOR")]

# Initialize cancer flag
indications_index_dt[, cancer_flag := FALSE]

# Apply rule logic using audit table
normal_idx <- which(audit_dt$tumor_nf_flag == FALSE & audit_dt$exclude_flag == FALSE)
normal_cols <- paste0("keyword_", cancer_keywords)
normal_hits <- rowSums(audit_dt[normal_idx, ..normal_cols]) > 0
set(indications_index_dt, i = normal_idx[normal_hits], j = "cancer_flag", value = TRUE)

# Tumor necrosis factor cases (exclude "TUMOR")
keywords_excluding_tumor <- setdiff(cancer_keywords, "TUMOR")
tnf_idx <- which(audit_dt$tumor_nf_flag == TRUE & audit_dt$exclude_flag == FALSE)
tnf_cols <- paste0("keyword_", keywords_excluding_tumor)
tnf_hits <- rowSums(audit_dt[tnf_idx, ..tnf_cols]) > 0
set(indications_index_dt, i = tnf_idx[tnf_hits], j = "cancer_flag", value = TRUE)

# Explicitly set excluded rows to FALSE
exclude_idx <- which(audit_dt$exclude_flag == TRUE)
set(indications_index_dt, i = exclude_idx, j = "cancer_flag", value = FALSE)

# Output summary
cat("Cancer-related indications:", sum(indications_index_dt$cancer_flag), "\n")
cat("Non-cancer indications:", sum(!indications_index_dt$cancer_flag), "\n")

# Export results
fwrite(indications_index_dt, "table_of_indications_with_cancer_flag.csv")
fwrite(audit_dt, "keyword_audit_table.csv")
```

Now we'll use this information to subset to those patients with cancer
indication in their data. We'll add the cancer flag to our final_dt.

```{r}
# Read in standard_case_indication file to build cancer flag
standard_case_indication <- fread("standard_case_indication.csv")
standard_case_indication[, compositeid := ifelse(!is.na(isr),
                                                   paste0("isr.", isr),
                                                   paste0("pid.", primaryid))]
# Clean up extra columns
standard_case_indication$isr <- NULL
standard_case_indication$primaryid <- NULL
```

Merge the cancer flag into the standard_case_indication from
indications_index_dt

```{r}
# Merge cancer_flag from indications_index_dt
standard_case_indication <- merge(
  standard_case_indication,
  indications_index_dt[, .(indication_concept_id, cancer_flag)],
  by = "indication_concept_id",
  all.x = TRUE
)

# Collapse to case-level cancer flag (any cancer indication → cancer case)
case_cancer_flag <- standard_case_indication[, .(cancer_flag = any(cancer_flag == TRUE)), by = compositeid]

# Merge back into final_dt
final_dt <- merge(final_dt, case_cancer_flag, by = "compositeid", all.x = TRUE)
final_dt[is.na(cancer_flag), cancer_flag := FALSE]
```

## Ensure all variables are factors

R requires that variables that are used for logistic regression are in
factor form. We’ll transform them now.

```{r}
# Ensure final_dt is a data.table
final_dt <- as.data.table(final_dt)

# Assign the cancer flag as a factor with two levels using 0/1 encoding
final_dt <- final_dt %>%
  mutate(cancer_flag = factor(as.integer(cancer_flag), levels = c(0, 1)))

# Character factors with specified levels
final_dt$sex <- factor(final_dt$sex, levels = c("F", "M", "NS"))
final_dt$ICI_Count_Factor <- factor(final_dt$ICI_Count_Factor, levels = c("Zero", "One", "Multiple"))

# Age
final_dt$age_group <- factor(
  final_dt$age_group,
  levels = c("0-20", "21-40", "41-60", "61-80", "80+", "Unknown")
)

final_dt$drug_class <- factor(final_dt$drug_class)
final_dt$region <- factor(final_dt$region)
final_dt$country <- factor(final_dt$country)
final_dt$outc_code <- factor(final_dt$outc_code)
final_dt$Death <- factor(final_dt$Death, levels = c(0, 1))
final_dt$SJSTEN <- factor(final_dt$SJSTEN, levels = c(0, 1))  # Ensures correct reference

# Binary 0/1 numeric columns to factors
binary_cols <- c("ICI_Pembrolizumab", "ICI_Nivolumab", "ICI_Ipilimumab", "ICI_Atezolizumab", "ICI_Durvalumab", "ICI_Cemiplimab", "ICI_Tremelimumab", "ICI_Dostarlimab", "ICI_Avelumab", "Strong_Culprit_Lamotrigine", "Strong_Culprit_TMP_SMX", "Strong_Culprit_Phenytoin", "Strong_Culprit_Carbamazepine", "Strong_Culprit_Allopurinol", "Weak_Culprit_Azithromycin", "Weak_Culprit_Ciprofloxacin", "Weak_Culprit_Oxcarbazepine", "Weak_Culprit_Sulfasalazine", "Weak_Culprit_Acyclovir", "ICI", "StrongCulprit", "WeakCulprit", "AllCulprit", "ICI_PD1", "ICI_PDL1", "ICI_CTLA4", "Dual_ICI")

final_dt[, (binary_cols) := lapply(.SD, function(x) factor(x, levels = c(0, 1))), .SDcols = binary_cols]

# Then fix final_dt to change NA in the drug_name column to "No Primary Suspect Listed"
final_dt <- final_dt %>% mutate(drug_name = ifelse(is.na(drug_name), "No Primary Suspect Listed", drug_name))

#Clean up memory
gc()
```

## Cancer auditing

There are several cases of ICI-exposed patients that have 0 for a cancer
flag, which is contradictory. Let's audit this further.

First, lets summarize the issue:

```{r}
# 🔁 Copy original data
dt_temp <- copy(final_dt)

# 🧼 Ensure consistent format for both flags
dt_temp[, ICI := as.character(ICI)]
dt_temp[, cancer_flag := as.character(cancer_flag)]

# 📊 Count combination frequencies
ici_cancer_table <- dt_temp[, .N, by = .(ICI, cancer_flag)]

# 🧱 Reshape to wide format
ici_cancer_wide <- dcast(
  ici_cancer_table,
  ICI ~ cancer_flag,
  value.var = "N",
  fill = 0
)

# 🏷 Label rows
rownames(ici_cancer_wide) <- fifelse(ici_cancer_wide$ICI == "0", "No ICI", "ICI")

# 🔁 Convert to matrix, selecting only "0" and "1" columns
ici_cancer_matrix <- as.matrix(ici_cancer_wide[, c("0", "1")])

# 🏷 Rename columns for readability
colnames(ici_cancer_matrix) <- c("No Cancer", "Cancer")

# ✅ Print final 2x2 matrix
print(ici_cancer_matrix)
```

So there are a substantial number of patients exposed to ICI without a
cancer flag - 22426 in fact, compared to 167330 ICI-exposed people with
cancer (meaning 11.8% are misclassified). Let's see what indications ICI
are given for in these patients.

```{r}
# Get compositeids from patients with ICI exposure but no cancer flag
ici_nocancer_ids <- final_dt[ICI == 1 & cancer_flag == 0, compositeid]

# Subset standard_case_drug and standard_case_indication to those patients
ici_nocancer_drugs <- standard_case_drug[compositeid %in% ici_nocancer_ids]
ici_nocancer_indications <- standard_case_indication[compositeid %in% ici_nocancer_ids]

# Merge on compositeid and drug_seq
ici_indications_merged <- merge(
  ici_nocancer_drugs,
  ici_nocancer_indications,
  by.x = c("compositeid", "drug_seq"),
  by.y = c("compositeid", "indi_drug_seq"),
  all.x = TRUE
)

# Filter down to ICI drugs only
ici_only_nocancer <- ici_indications_merged[drug_concept_id %in% ici_list]

# Summarize indications
ici_only_nocancer[, .N, by = indi_pt][order(-N)]
```

Manually inspecting that file reveals that 14974 cases use ICI for an
unknown indication, and 10468 don't have an indication listed. That's
the vast majority.

To fix this - we'll flag patients that receive a drug only used for
cancer as cancer patients. To figure out which drugs those are, we have
to audit the drugs manually. The rules we'll use are:

-   If the drug is used, either on-label or off-label, only for the
    direct treatment of neoplasms, it will be **included** as a cancer
    drug. This includes all checkpoint inhibitors, almost alkylating
    agents (except cyclophosphamide), and almost all kinase inhibitors
    (except JAKi).

-   If the drug is used, either on-label or off-label, for the treatment
    of autoimmune conditions (rituximab, JAK inhibitor), it will **NOT
    be included** as a cancer drug.

-   If the drug is used, either on-label of off-label, for endocrine
    indications (central precocious puberty, endometriosis, transgender
    therapy), it will **not be included**. This one is more tricky,
    because these are often off-label and center-dependent, but we will
    be conservative about what to call a cancer drug.

-   If the drug is mainly or only used for some cancer-adjacent or
    chemotherapy-induced condition, rather than cancer itself
    (filgrastim, mesna, drugs used for bone marrow depletion before
    HSCT, meds for pain or nausea), it will be **excluded** as a cancer
    drug.

```{r}
# 🔁 Merge drug and indication tables on compositeid + drug_seq
drug_indication_merged <- merge(
  standard_case_drug,
  standard_case_indication,
  by.x = c("compositeid", "drug_seq"),
  by.y = c("compositeid", "indi_drug_seq"),
  all = FALSE
)

# 🧼 Filter to rows where cancer_flag == TRUE
cancer_drug_usage <- drug_indication_merged[cancer_flag == TRUE]

# 📊 Count how often each drug appears with a cancer indication
cancer_drug_counts <- cancer_drug_usage[
  , .N, by = .(drug_name, drug_concept_id)
][order(-N)]

# Export to CSV
fwrite(cancer_drug_counts, "drugs_with_cancer_indications.csv")
```

I have manually annotated the top 1000 drugs given for cancer-related
indications, of which 252 were marked as cancer drugs. Now, we will add
the cancer flag to any patient exposed to those drugs.

```{r}
# 📥 Load manually annotated drug list
manual_cancer_drugs <- fread("drugs_with_cancer_indications_top1000.csv")

# ✅ Get drug_concept_ids manually marked as cancer drugs
confirmed_cancer_drugs <- manual_cancer_drugs[cancer_drug_manual == TRUE, drug_concept_id]

# 🔍 Find compositeids exposed to these drugs
cancer_patients_from_drugs <- standard_case_drug[
  drug_concept_id %in% confirmed_cancer_drugs,
  unique(compositeid)
]

# ✅ Clean up internal references before column assignment
setDT(final_dt)

# 🧪 Store original flag for comparison (as character)
final_dt[, prev_cancer_flag := as.character(cancer_flag)]

# ✅ Update to "1" (the cancer-positive level) where appropriate
final_dt[compositeid %in% cancer_patients_from_drugs, cancer_flag := "1"]

# 🔢 Count how many new updates occurred
new_flags_added <- final_dt[
  compositeid %in% cancer_patients_from_drugs &
  prev_cancer_flag == "0" & cancer_flag == "1",
  .N
]

# 🧼 Clean up and restore factor type
final_dt[, prev_cancer_flag := NULL]
final_dt[, cancer_flag := factor(cancer_flag, levels = c("0", "1"))]

# 📊 Report
cat("Number of new cancer_flag entries added based on manual drug annotation:", new_flags_added, "\n")

```

Let's see how that matrix looks now:

```{r}
# 🔁 Copy original data
dt_temp <- copy(final_dt)

# 🧼 Ensure consistent format for both flags
dt_temp[, ICI := as.character(ICI)]
dt_temp[, cancer_flag := as.character(cancer_flag)]

# 📊 Count combination frequencies
ici_cancer_table <- dt_temp[, .N, by = .(ICI, cancer_flag)]

# 🧱 Reshape to wide format
ici_cancer_wide <- dcast(
  ici_cancer_table,
  ICI ~ cancer_flag,
  value.var = "N",
  fill = 0
)

# 🏷 Label rows
rownames(ici_cancer_wide) <- fifelse(ici_cancer_wide$ICI == "0", "No ICI", "ICI")

# 🔁 Convert to matrix, selecting only "0" and "1" columns
ici_cancer_matrix <- as.matrix(ici_cancer_wide[, c("0", "1")])

# 🏷 Rename columns for readability
colnames(ici_cancer_matrix) <- c("No Cancer", "Cancer")

# ✅ Print final 2x2 matrix
print(ici_cancer_matrix)

```

Excellent - all ICI cases are now cancer-related.

# FIGURE GENERATION

Now that we've generated the final_dt and the TTE information, we can
use it to generate the figures directly - they contain all the
information needed.

## FIGURE OUTLINE

This section describes the structure and purpose of each figure
generated for the analysis.

-   **Figure 1 – SJS/TEN Cases in FAERS**\
    Summary of the \~17,000 SJS/TEN cases in FAERS, including age, sex,
    and causative drug.

-   **Figure S1 – Demographics**\
    Descriptive plots of age, sex, region, and suspected culprit drug
    class across all SJS/TEN cases in FAERS from 2013–2023.

-   **Figure 2 – Disproportionality Analysis**\
    Heatmaps of reporting odds ratios (RORs) for SJS/TEN in association
    with ICI-small molecule drug pairs, highlighting known high-risk
    combinations.

-   **Figure S2 – Allopurinol Logit**\
    Logistic regression evaluating the interaction of allopurinol and
    ICI exposure on SJS/TEN development, adjusting for age, sex, drug
    count, and cancer.

-   **Figure S3 – TMP-SMX Logit**\
    Logistic regression evaluating the interaction of TMP-SMX and ICI
    exposure on SJS/TEN development, adjusting for age, sex, drug count,
    and cancer.

-   **Figure 3 – Development and Mortality Logistic Models (Full
    FAERS)**\
    Multivariate logistic regressions for predictors of SJS/TEN
    development and associated mortality in the full FAERS dataset,
    including strong and weak culprit exposure, cancer, ICI exposure,
    and demographic covariates.

-   **Figure S4 – Restricted Logit Model**\
    A model based on Figure 3, but excluding culprit exposure variables,
    to explore the independent relationship of cancer and ICI with
    SJS/TEN development in a more restrictive fashion.

-   **Figure 4 – Small-Molecule SJS/TEN Logistic Models**\
    Multivariable logistic models evaluating predictors of both SJS/TEN
    development and mortality, restricted to reports with small molecule
    drugs as the primary suspect. Models include drug identity, ICI
    exposure (in three levels), age, sex, and drug count.

-   **Figure S5 – Small-Molecule SJS/TEN Logistic Models (Cancer
    Patients)**\
    Replication of Figure 4 within the subset of patients with cancer
    indications in FAERS. The reason it was subsetted in this way,
    instead of just including cancer as a variable in Figure 4, is
    because the set of causative drugs is different in the cancer
    population than the general population.

-   **Figure S6 and S7 – ICI-Induced SJS/TEN Logit**\
    Logistic regression among reports where ICIs were the primary
    suspect drug, evaluating the role of co-prescribed small molecules
    in mediating SJS/TEN risk.

-   **Figure 5 – Latency Analysis (Full FAERS)**\
    Median time-to-onset for SJS/TEN by drug, stratified by ICI exposure
    and cancer diagnosis.

-   **Figure S8 – Latency Analysis (Small-Molecule SJS/TEN)**\
    Subset of latency analysis restricted to patients with
    small-molecule-induced SJS/TEN, stratified by ICI exposure.

-   **Figure S9 – Latency Analysis (Small-Molecule SJS/TEN, Cancer
    Patients)**\
    Further subset of latency analysis restricted to cancer patients
    with small-molecule-induced, stratified by ICI exposure.

-   **Figure S10 – VIF Heatmap**\
    Heatmap summarizing generalized variance inflation factors (GVIF)
    across all major regression models to assess collinearity between
    covariates.

## FIGURE 1 - SJS/TEN SUMMARY

### FIGURE 1A - SUMMARY TABLE

This generates a table summarizing SJS/TEN cases in FAERS by various
demographic variables. We use the gt package to generate HTML tables
with a clean look.

```{r}
# Subset final_dt to just SJS/TEN cases
final_SJSTEN_dt <- final_dt[SJSTEN == 1]

# ─── Prepare Components ─────────────────────────────────────────────

# Total patients
total_row <- tibble(
  Group = "Overall",
  Subgroup = "Total Patients",
  Value = as.character(nrow(final_SJSTEN_dt))
)

# Age summary
age_row <- final_SJSTEN_dt %>%
  summarise(
    Group = "Overall",
    Subgroup = "Age (Median [IQR])",
    Value = sprintf("%.1f [%d–%d]",
                    median(age_yr, na.rm = TRUE),
                    quantile(age_yr, 0.25, na.rm = TRUE),
                    quantile(age_yr, 0.75, na.rm = TRUE))
  )

# Sex breakdown
sex_rows <- final_SJSTEN_dt %>%
  count(sex) %>%
  mutate(
    Group = "Sex",
    Subgroup = as.character(sex),
    Value = as.character(n)
  ) %>%
  select(Group, Subgroup, Value)

# Region breakdown
region_rows <- final_SJSTEN_dt %>%
  count(region) %>%
  mutate(
    Group = "Region",
    Subgroup = as.character(region),
    Value = as.character(n)
  ) %>%
  select(Group, Subgroup, Value)

# Mortality outcome
death_rows <- final_SJSTEN_dt %>%
  count(Death) %>%
  mutate(
    Group = "Outcomes",
    Subgroup = case_when(
      Death == 1 ~ "Dead",
      Death == 0 ~ "Alive",
      TRUE ~ "Unknown"
    ),
    Value = as.character(n)
  ) %>%
  select(Group, Subgroup, Value)

# Top 10 Causative Drugs
top_drugs_rows <- final_SJSTEN_dt %>%
  count(drug_name, sort = TRUE) %>%
  mutate(
    Group = "Top 10 Causative Drugs",
    Subgroup = drug_name,
    Value = as.character(n)
  ) %>%
  slice_head(n = 10) %>%
  select(Group, Subgroup, Value)

# ICI Exposure
ici_exposure_rows <- final_SJSTEN_dt %>%
  mutate(ICI_exposed = ifelse(ICI == 1, "Yes", "No")) %>%
  count(ICI_exposed) %>%
  mutate(
    Group = "ICI Exposure",
    Subgroup = ICI_exposed,
    Value = as.character(n)
  ) %>%
  select(Group, Subgroup, Value)

# Cancer Diagnosis
cancer_rows <- final_SJSTEN_dt %>%
  mutate(cancer_flag = ifelse(cancer_flag == 1, "Yes", "No")) %>%
  count(cancer_flag) %>%
  mutate(
    Group = "Cancer Diagnosis",
    Subgroup = as.character(cancer_flag),
    Value = as.character(n)
  ) %>%
  select(Group, Subgroup, Value)

# ─── Combine Table ─────────────────────────────────────────────────

summary_table <- bind_rows(
  total_row,
  age_row,
  sex_rows,
  region_rows,
  death_rows,
  top_drugs_rows,
  ici_exposure_rows,
  cancer_rows
)

# ─── Render Table with `gt` ────────────────────────────────────────

Figure1A <- summary_table %>%
  gt(groupname_col = "Group") %>%
  tab_header(
    title = md("**Baseline Characteristics of SJS/TEN Cases in FAERS, 2013-2023**"),
    subtitle = "Demographic, Exposure, and Drug Summary"
  ) %>%
  cols_label(
    Subgroup = "",
    Value = "N or Summary"
  ) %>%
  tab_options(
    table.font.names = "Arial",
    table.font.size = 12,
    heading.title.font.size = 14,
    heading.subtitle.font.size = 12,
    row_group.as_column = TRUE,
    data_row.padding = px(4)
  )

#Display tabel
Figure1A 
```

### FIGURE 1B - STACKED BAR CHART OF ICI-EXPOSED PATIENTS

This generates a stacked bar chart showing the number of SJS/TEN cases
involving immune checkpoint inhibitors (ICIs), separated by drug or drug
class and categorized by exposure type. Each bar represents the total
number of SJS/TEN cases in which a given ICI was either reported as the
*primary suspect* (dark red) or as *concomitant exposure* (light blue).

```{r}
# Helper function to count exposed + primary cause
get_counts <- function(df, colname, label, concept_id) {
  df %>%
    summarise(
      total_exposed = sum(get(colname) == 1, na.rm = TRUE),
      primary_suspect = sum(drug_concept_id == concept_id, na.rm = TRUE)
    ) %>%
    mutate(group = label)
}

# Individual ICI counts
individual_counts <- bind_rows(
  get_counts(final_SJSTEN_dt, "ICI_Pembrolizumab", "Pembrolizumab", 45775965),
  get_counts(final_SJSTEN_dt, "ICI_Nivolumab", "Nivolumab", 45892628),
  get_counts(final_SJSTEN_dt, "ICI_Ipilimumab", "Ipilimumab", 40238188),
  get_counts(final_SJSTEN_dt, "ICI_Atezolizumab", "Atezolizumab", 42629079),
  get_counts(final_SJSTEN_dt, "ICI_Durvalumab", "Durvalumab", 1594034),
  get_counts(final_SJSTEN_dt, "ICI_Cemiplimab", "Cemiplimab", 35200783),
  get_counts(final_SJSTEN_dt, "ICI_Tremelimumab", "Tremelimumab", 741851),
  get_counts(final_SJSTEN_dt, "ICI_Dostarlimab", "Dostarlimab", 1536789), 
  get_counts(final_SJSTEN_dt, "ICI_Avelumab", "Avelumab", 1593273)
)

# Grouped categories (PD-1, PD-L1, etc)
group_counts <- tibble(
  group = c("All ICIs", "PD-1", "PD-L1", "CTLA-4", "Dual ICI"),
  total_exposed = c(
    sum(final_SJSTEN_dt$ICI == 1, na.rm = TRUE),
    sum(final_SJSTEN_dt$ICI_PD1 == 1, na.rm = TRUE),
    sum(final_SJSTEN_dt$ICI_PDL1 == 1, na.rm = TRUE),
    sum(final_SJSTEN_dt$ICI_CTLA4 == 1, na.rm = TRUE),
    sum(final_SJSTEN_dt$Dual_ICI == 1, na.rm = TRUE)
  ),
  primary_suspect = c(
    sum(final_SJSTEN_dt$drug_concept_id %in% ici_list, na.rm = TRUE),
    sum(final_SJSTEN_dt$drug_concept_id %in% pd1_list, na.rm = TRUE),
    sum(final_SJSTEN_dt$drug_concept_id %in% pdl1_list, na.rm = TRUE),
    sum(final_SJSTEN_dt$drug_concept_id %in% ctla4_list, na.rm = TRUE),
    sum(final_SJSTEN_dt$Dual_ICI == 1 & final_SJSTEN_dt$drug_concept_id %in% ici_list, na.rm = TRUE)
  )
)

# Combine and reshape data
plot_data <- bind_rows(individual_counts, group_counts) %>%
  mutate(other_exposed = total_exposed - primary_suspect) %>%
  select(group, primary_suspect, other_exposed) %>%
  tidyr::pivot_longer(cols = c("primary_suspect", "other_exposed"),
                      names_to = "type", values_to = "count") %>%
  mutate(type = factor(type, levels = c("other_exposed", "primary_suspect"),
                       labels = c("Exposed", "Primary Suspect")))

# Plot with bar borders and legend enhancements
Figure1B <- ggplot(plot_data, aes(x = reorder(group, -count), y = count, fill = type)) +
  geom_bar(stat = "identity", color = "black", linewidth = 0.3) +  # Black border
  scale_fill_manual(
    values = c("Exposed" = "#B0C4DE", "Primary Suspect" = "#DC143C"),
    name = "Exposure Type"
  ) +
  labs(
    title = "ICI Exposure in SJS/TEN Cases",
    x = "Drug or Class",
    y = "Number of Cases"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "right",
    legend.box.background = element_rect(color = "black", fill = NA),
    legend.title = element_text(face = "bold")
  )

# Display the figure
Figure1B
```

## FIGURE S1 - DEMOGRAPHICS

In this supplementary figure, we'll compare two sets - ICI-primary
suspect cases, and small-molecule primary suspect cases. We'll also
include a panel showing what percentage of these cases have strong
culprit exposure.

```{r figure-s1, fig.width=12, fig.height=10}
# Subset
ici_cases <- final_SJSTEN_dt[drug_concept_id %in% unname(ici_list)]
sm_cases  <- final_SJSTEN_dt[drug_class == "Small Molecule"]

# Label groups
ici_cases$group <- "ICI-induced"
sm_cases$group  <- "SM-induced"

# Combine
demographic_compare_dt <- rbind(ici_cases, sm_cases)

# 🧼 Handle missing
demographic_compare_dt$region[is.na(demographic_compare_dt$region)] <- "Unknown"

# 🟨 Panel A: Age Group (Stacked)
panel_age <- demographic_compare_dt %>%
  count(group, age_group) %>%
  group_by(group) %>%
  mutate(percentage = n / sum(n)) %>%
  ggplot(aes(x = group, y = percentage, fill = age_group)) +
  geom_bar(stat = "identity", color = "black") +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Age Distribution", x = "", y = "Proportion", fill = "Age (Years)") +
  theme_minimal(base_size = 12)

# 🟦 Panel B: Sex Distribution (Stacked)
panel_sex <- demographic_compare_dt %>%
  count(group, sex) %>%
  group_by(group) %>%
  mutate(percentage = n / sum(n)) %>%
  ggplot(aes(x = group, y = percentage, fill = sex)) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Sex Distribution", x = "", y = "Proportion", fill = "Sex") +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal(base_size = 12)

# 🟩 Panel C: Region Distribution (Stacked)
panel_region <- demographic_compare_dt %>%
  count(group, region) %>%
  group_by(group) %>%
  mutate(percentage = n / sum(n)) %>%
  ggplot(aes(x = group, y = percentage, fill = region)) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_brewer(palette = "Pastel1") +
  labs(title = "Region Distribution", x = "", y = "Proportion", fill = "Region") +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal(base_size = 12)

# 🟥 Panel D: Strong Culprit Classification (Three categories)
ici_ids <- unname(ici_list)
strong_ids <- unlist(strong_culprit_list)

# ICI group
ici_bar <- final_SJSTEN_dt[drug_concept_id %in% ici_ids][
  , .(Group = "ICI-induced", Subgroup = ifelse(StrongCulprit == 1, "Strong Culprit Present (Not PS)", "No Strong Culprit"))
]

# Small Molecule group
sm_bar <- final_SJSTEN_dt[drug_class == "Small Molecule"][
  , Subgroup := fifelse(
      drug_concept_id %in% strong_ids, 
      "Strong Culprit (PS)",
      fifelse(StrongCulprit == 1, "Strong Culprit Present (Not PS)", "No Strong Culprit")
  )
][, .(Group = "SM-induced", Subgroup)]

# Combine
culprit_plot_dt <- rbind(ici_bar, sm_bar)

# Plot
panel_culprit <- culprit_plot_dt %>%
  count(Group, Subgroup) %>%
  group_by(Group) %>%
  mutate(percentage = n / sum(n)) %>%
  ggplot(aes(x = Group, y = percentage, fill = factor(Subgroup,
    levels = c("Strong Culprit (PS)", "Strong Culprit Present (Not PS)", "No Strong Culprit")))) +
  geom_bar(stat = "identity", color = "black") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(
    values = c(
      "Strong Culprit (PS)" = "#DC143C",
      "Strong Culprit Present (Not PS)" = "#B0C4DE",
      "No Strong Culprit" = "gray80"
    )
  ) +
  labs(
    title = "Culprit Classification by Drug Type",
    x = "",
    y = "Proportion",
    fill = "Classification"
  ) +
  theme_minimal(base_size = 12)

# 🧩 2x2 Final Layout
Figure_S1 <- (
  panel_age + panel_sex + panel_region + panel_culprit
) + plot_layout(ncol = 2)

# Display the figure
Figure_S1
```

## FIGURE 2 - DISPROPORTIONALITY ANALYSIS

### Function to Compute Disproportionality

We'll use a unified function that calculates disproportionality metrics
from final_dt, including filtering down the denominator by an exposure
if needed.

```{r}
# This function can compute disproportionality as described above
compute_disproportionality <- function(
  final_dt,
  additional_exposure_col,
  label_exposure,
  background_exposure_col = NULL
) {
  # Convert factor -> character -> integer safely
  outcome_flag <- as.integer(as.character(final_dt$SJSTEN))
  exposure_flag <- as.integer(as.character(final_dt[[additional_exposure_col]]))
  
  background_flag <- if (!is.null(background_exposure_col)) {
    as.integer(as.character(final_dt[[background_exposure_col]]))
  } else {
    rep(1L, nrow(final_dt))
  }
  
  # Subsets
  exposed_rows    <- which(exposure_flag == 1 & background_flag == 1)
  background_rows <- which(exposure_flag == 0 & background_flag == 1)
  
  # Contingency table, coerced to numeric to prevent overflow
  a <- as.numeric(sum(outcome_flag[exposed_rows] == 1, na.rm = TRUE))
  b <- as.numeric(sum(outcome_flag[exposed_rows] == 0, na.rm = TRUE))
  c <- as.numeric(sum(outcome_flag[background_rows] == 1, na.rm = TRUE))
  d <- as.numeric(sum(outcome_flag[background_rows] == 0, na.rm = TRUE))
  
  # Fail-safe
  if ((a + b) == 0 | (c + d) == 0 | b == 0 | c == 0 | d == 0) {
    return(data.frame(
      Exposure = label_exposure,
      a = a, b = b, c = c, d = d,
      PRR = NA, PRR_lower = NA, PRR_upper = NA,
      ROR = NA, ROR_lower = NA, ROR_upper = NA
    ))
  }
  
  # PRR and SE
  prr <- (a / (a + b)) / (c / (c + d))
  prr_se <- sqrt((b / (a * (a + b))) + (d / (c * (c + d))))
  prr_lower <- exp(log(prr) - 1.96 * prr_se)
  prr_upper <- exp(log(prr) + 1.96 * prr_se)
  
  # ROR and SE
  ror <- (a * d) / (b * c)
  ror_se <- sqrt(1 / a + 1 / b + 1 / c + 1 / d)
  ror_lower <- exp(log(ror) - 1.96 * ror_se)
  ror_upper <- exp(log(ror) + 1.96 * ror_se)
  
  # Output
  return(data.frame(
    Exposure = label_exposure,
    a = a, b = b, c = c, d = d,
    PRR = prr, PRR_lower = prr_lower, PRR_upper = prr_upper,
    ROR = ror, ROR_lower = ror_lower, ROR_upper = ror_upper
  ))
}
```

### FIGURE 2A - ICI AS BACKGROUND

The first heatmap will display ICIs on the y and culprits on the X. ICI,
ICI classes, and FAERS as a whole will be the background populations.

```{r}
# Define reference populations (Y-axis)
individual_ici_cols <- grep("^ICI_.*mab$", names(final_dt), value = TRUE)
individual_ici_labels <- tools::toTitleCase(gsub("^ICI_", "", individual_ici_cols))

reference_cols <- c(
  NULL, "ICI", "Dual_ICI", "ICI_CTLA4", "ICI_PD1", "ICI_PDL1", individual_ici_cols
)
reference_labels <- c(
  "All FAERS", "All ICIs", "Dual ICI", "CTLA-4", "PD-1", "PD-L1", individual_ici_labels
)

# Define exposure columns (X-axis)
# Define all exposure (culprit) groups
base_additional <- list(
  "All Culprits" = "AllCulprit",
  "All Strong Culprits" = "StrongCulprit",
  "All Weak Culprits" = "WeakCulprit"
)
strong_cols <- grep("^Strong_Culprit_", names(final_dt), value = TRUE)
strong_named <- setNames(strong_cols, tools::toTitleCase(gsub("Strong_Culprit_", "", strong_cols)))
weak_cols <- grep("^Weak_Culprit_", names(final_dt), value = TRUE)
weak_named <- setNames(weak_cols, tools::toTitleCase(gsub("Weak_Culprit_", "", weak_cols)))
additional_groups <- c(base_additional, strong_named, weak_named)

# Loop for All FAERS
heatmap_results <- list()
pb <- txtProgressBar(min = 0, max = length(additional_groups), style = 3)
step <- 0

for (add_label in names(additional_groups)) {
  add_col <- additional_groups[[add_label]]
  step <- step + 1

  res <- compute_disproportionality(
    final_dt,
    additional_exposure_col = add_col,
    label_exposure = add_label,
    background_exposure_col = NULL
  )

  res$Reference <- "All FAERS"
  heatmap_results[[length(heatmap_results) + 1]] <- res
  setTxtProgressBar(pb, step)
}
close(pb)

# Loop for other reference populations (ICIs)
ici_cols <- c("ICI", "Dual_ICI", "ICI_CTLA4", "ICI_PD1", "ICI_PDL1",
              grep("^ICI_.*mab$", names(final_dt), value = TRUE))
ici_labels <- c("All ICIs", "Dual ICI", "CTLA-4", "PD-1", "PD-L1",
                tools::toTitleCase(gsub("^ICI_", "", grep("^ICI_.*mab$", names(final_dt), value = TRUE))))

pb <- txtProgressBar(min = 0, max = length(ici_cols) * length(additional_groups), style = 3)
step <- 0

for (i in seq_along(ici_cols)) {
  ref_col <- ici_cols[i]
  ref_label <- ici_labels[i]

  for (add_label in names(additional_groups)) {
    add_col <- additional_groups[[add_label]]
    step <- step + 1

    res <- compute_disproportionality(
      final_dt,
      additional_exposure_col = add_col,
      label_exposure = add_label,
      background_exposure_col = ref_col
    )

    res$Reference <- ref_label
    heatmap_results[[length(heatmap_results) + 1]] <- res
    setTxtProgressBar(pb, step)
  }
}
close(pb)

# Combine all
heatmap_dt_2A <- data.table::rbindlist(heatmap_results, fill = TRUE)
```

Now that the dt is made, add some additional columns needed to make the
heatmap. We need logROR for the heatmap scale (the actual cells will
show the ROR).

```{r}
# logROR and standard error
heatmap_dt_2A[, `:=`(
  logROR = log(ROR),
  logROR_se = sqrt(1 / a + 1 / b + 1 / c + 1 / d)
)]

# Wald p-value using logROR
heatmap_dt_2A[, p_raw := 2 * pnorm(-abs(logROR) / logROR_se)]

# BH correction and stars
heatmap_dt_2A[, p_adj := p.adjust(p_raw, method = "BH")]
heatmap_dt_2A[, stars := fifelse(is.na(p_adj), "",
                        fifelse(p_adj < 0.001, "***",
                        fifelse(p_adj < 0.01, "**",
                        fifelse(p_adj < 0.05, "*", ""))))]

# Confidence interval with brackets and stars
heatmap_dt_2A[, CI_text := sprintf("[%.1f–%.1f] %s", ROR_lower, ROR_upper, stars)]

# Combine into multi-line label for ggplot
heatmap_dt_2A[, Label := paste0(
  "n = ", a, "\n",
  "ROR = ", round(ROR, 1), "\n",
  CI_text
)]

#Clean up memory
gc()
```

Then make the plot from this new dt

```{r heatmap_Fig2A, fig.width = 14, fig.height = 10, out.width = "100%"}
# Define fixed Y-axis order
fixed_y_order <- c(
  "All FAERS", "All ICIs", "Dual ICI", "CTLA-4", "PD-1", "PD-L1"
)

# Extract remaining individual ICIs and order by total 'a'
ici_extras <- heatmap_dt_2A[
  !Reference %in% fixed_y_order, 
  .(a_total = sum(a, na.rm = TRUE)), 
  by = Reference
][order(-a_total)]$Reference

# Combine for full Y-axis order
heatmap_dt_2A[, Reference := factor(Reference, levels = c(fixed_y_order, ici_extras))]

# Preserve X-axis order as entered
heatmap_dt_2A[, Exposure := factor(Exposure, levels = unique(Exposure))]

# Filter out zero-count cells
heatmap_plot_dt <- heatmap_dt_2A[a > 0]

# Now plot it
Figure2A <- ggplot(heatmap_plot_dt, aes(x = Exposure, y = Reference, fill = logROR)) +
  geom_tile(color = "white") +
  geom_text(aes(label = Label), size = 3, lineheight = 1.1, na.rm = TRUE) +
  scale_fill_gradient2(
    low = "blue", mid = "white", high = "red", 
    midpoint = 0, na.value = "grey90",
    name = "logROR"
  ) +
  labs(
    title = "Additive Effect of Known Culprit Medications on Patients Recieving ICIs",
    x = "Additional Drug",
    y = "Reference Population"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(size = 10),
    panel.grid = element_blank()
  )

# Show it
Figure2A
```

### FIGURE 2B - CULPRIT AS BACKGROUND

This time, we'll use patients exposed to culprits as the background
population, and ICI exposure as additional factors. FAERS will still be
used as a background population.

```{r}
# Setup: ICI exposures (X-axis)
ici_cols <- c("ICI", "Dual_ICI", "ICI_CTLA4", "ICI_PD1", "ICI_PDL1",
              grep("^ICI_.*mab$", names(final_dt), value = TRUE))
ici_labels <- c("All ICIs", "Dual ICI", "CTLA-4", "PD-1", "PD-L1",
                tools::toTitleCase(gsub("^ICI_", "", grep("^ICI_.*mab$", names(final_dt), value = TRUE))))

# Setup: Culprit-based reference groups (Y-axis)
base_reference <- list(
  "All Culprits" = "AllCulprit",
  "All Strong Culprits" = "StrongCulprit",
  "All Weak Culprits" = "WeakCulprit"
)

strong_cols <- grep("^Strong_Culprit_", names(final_dt), value = TRUE)
strong_named <- setNames(as.list(strong_cols), tools::toTitleCase(gsub("^Strong_Culprit_", "", strong_cols)))

weak_cols <- grep("^Weak_Culprit_", names(final_dt), value = TRUE)
weak_named <- setNames(as.list(weak_cols), tools::toTitleCase(gsub("^Weak_Culprit_", "", weak_cols)))

reference_groups <- c(base_reference, strong_named, weak_named)

heatmap_results <- list()
pb <- txtProgressBar(min = 0, max = length(ici_cols), style = 3)
step <- 0

for (i in seq_along(ici_cols)) {
  exp_col <- ici_cols[i]
  exp_label <- ici_labels[i]
  step <- step + 1

  res <- compute_disproportionality(
    final_dt,
    additional_exposure_col = exp_col,
    label_exposure = exp_label,
    background_exposure_col = NULL  # All FAERS
  )

  res$Reference <- "All FAERS"
  heatmap_results[[length(heatmap_results) + 1]] <- res
  setTxtProgressBar(pb, step)
}
close(pb)


pb <- txtProgressBar(min = 0, max = length(reference_groups) * length(ici_cols), style = 3)
step <- 0

for (ref_label in names(reference_groups)) {
  ref_col <- reference_groups[[ref_label]]

  for (i in seq_along(ici_cols)) {
    exp_col <- ici_cols[i]
    exp_label <- ici_labels[i]
    step <- step + 1

    res <- compute_disproportionality(
      final_dt,
      additional_exposure_col = exp_col,
      label_exposure = exp_label,
      background_exposure_col = ref_col
    )

    res$Reference <- ref_label
    heatmap_results[[length(heatmap_results) + 1]] <- res
    setTxtProgressBar(pb, step)
  }
}
close(pb)

# Combine results
heatmap_dt_2B <- data.table::rbindlist(heatmap_results, fill = TRUE)

# Calculate logROR and SE separately to avoid eval order issues
heatmap_dt_2B[, `:=`(
  logROR = log(ROR),
  logROR_se = sqrt(1 / a + 1 / b + 1 / c + 1 / d)
)]

# p-value and adjustment
heatmap_dt_2B[, p_raw := 2 * pnorm(-abs(logROR) / logROR_se)]
heatmap_dt_2B[, p_adj := p.adjust(p_raw, method = "BH")]

# Stars and formatted labels
heatmap_dt_2B[, stars := fifelse(is.na(p_adj), "",
                        fifelse(p_adj < 0.001, "***",
                        fifelse(p_adj < 0.01, "**",
                        fifelse(p_adj < 0.05, "*", ""))))]

heatmap_dt_2B[, CI_text := sprintf("[%.1f–%.1f] %s", ROR_lower, ROR_upper, stars)]
heatmap_dt_2B[, Label := paste0("n = ", a, "\nROR = ", round(ROR, 1), "\n", CI_text)]

#Clean up memory
gc()
```

Now make the plot

```{r heatmap_Fig2B, fig.width = 14, fig.height = 10, out.width = "100%"}
# Define fixed X-axis order for ICI exposures
fixed_ici_order <- c("All ICIs", "Dual ICI", "CTLA-4", "PD-1", "PD-L1")

# Get individual ICIs ordered by descending 'a'
ici_extras <- heatmap_dt_2B[
  !Exposure %in% fixed_ici_order,
  .(a_total = sum(a, na.rm = TRUE)),
  by = Exposure
][order(-a_total)]$Exposure

# Apply full X-axis order
heatmap_dt_2B[, Exposure := factor(Exposure, levels = c(fixed_ici_order, ici_extras))]

# Y-axis (reference populations): use exact input order from the loop
reference_order <- c(
  "All FAERS",
  "All Culprits",
  "All Strong Culprits",
  "All Weak Culprits",
  names(reference_groups)  # individual strong + weak culprits
)

# Ensure no duplicates
reference_order <- unique(reference_order)

heatmap_dt_2B[, Reference := factor(Reference, levels = reference_order)]

# Filter out n = 0 cells
heatmap_plot_dt <- heatmap_dt_2B[a > 0]

#Plot it
Figure2B <- ggplot(heatmap_plot_dt, aes(x = Exposure, y = Reference, fill = logROR)) +
  geom_tile(color = "white") +
  geom_text(aes(label = Label), size = 3, lineheight = 1.1, na.rm = TRUE) +
  scale_fill_gradient2(
    low = "blue", mid = "white", high = "red",
    midpoint = 0, na.value = "grey90", name = "logROR"
  ) +
  labs(
    title = "Additive Effect of ICI on Patients Recieving Known Culprit Medications",
    x = "ICI Exposure",
    y = "Reference Drug Population"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(size = 10),
    panel.grid = element_blank()
  )

#Show it
Figure2B
```

## FIGURE S2 - ALLOPURINOL LOGIT

This logistic regression subsets the entire dataset to just those
patients exposed to allopurinol, and asks whether co-exposure to ICI is
associated with the development of SJS/TEN. There are 43 patients
co-exposed to allopurinol and an ICI (which we determined from Figure
2), which is sufficient for a logistic regression.

```{r}
# Subset allopurinol cases from final_dt
final_dt_allopurinol_any <- final_dt[Strong_Culprit_Allopurinol == 1]

# Fit the logistic regression model
logit_allopurinol_any_ICI <- glm(
  SJSTEN ~ age_group + sex + NumDrugs + ICI + cancer_flag,
  data = final_dt_allopurinol_any,
  family = binomial
)

# Tidy the results
summary_tbl_FigS2 <- logit_summary_table(logit_allopurinol_any_ICI, final_dt_allopurinol_any)  # Converts log-odds to ORs

# Transform into gt table
FigureS2 <- summary_tbl_FigS2 %>%
  mutate(
    Variable = case_when(
      Variable == "age_group" ~ "Age (Years)",
      Variable == "sex" ~ "Sex",
      Variable == "NumDrugs" ~ "Number of Concomittant Drugs",
      Variable == "ICI" ~ "ICI Exposure",
      Variable == "cancer_flag" ~ "Cancer Diagnosis",
      TRUE ~ Variable
    ),
    Level = case_when(
      Level == "M" ~ "Male",
      Level == "F" ~ "Female",
      Level == "F (ref)" ~ "Female (ref)",
      Level == "NS" ~ "Not Specified",
      Level == "0 (ref)" ~ "No (ref)",
      Level == "1" ~ "Yes",
      Level == "0" ~ "No",
      TRUE ~ Level
    )
  ) %>%
  gt(groupname_col = "Variable") %>%
  tab_header(
    title = md("**SJS/TEN in Patients Exposed to Allopurinol**"),
    subtitle = "Multivariable Logistic Regression, FAERS 2013-2023",
  ) %>%
  cols_label(
    Level = "",
    N = "N (Total)",
    SJS_TEN_Cases = "N (SJS/TEN)",
    OR = "Odds Ratio",
    CI = "95% CI",
    P = "P-value"
  ) %>%
  sub_missing(columns = everything(), missing_text = "-") %>%
  tab_options(
    table.font.names = "Times",
    table.font.size = 12,
    heading.title.font.size = 14,
    heading.subtitle.font.size = 12,
    row_group.as_column = TRUE,
    data_row.padding = px(4)
  )

# Display the table
FigureS2
```

## FIGURE S3 - TMP-SMX LOGIT

This logistic regression subsets the entire dataset to just those
patients exposed to TMP-SMX, and asks whether co-exposure to ICI is
associated with the development of SJS/TEN. There are 43 co-exposed
patients in this analysis.

```{r}
#Subset TMP/SMX cases from final_dt
final_dt_tmpsmx_any <- final_dt[Strong_Culprit_TMP_SMX == 1]

# Fit the logistic regression model
logit_tmpsmx_any_ICI <- glm(
  SJSTEN ~ age_group + sex + NumDrugs + ICI + cancer_flag,
  data = final_dt_tmpsmx_any,
  family = binomial
)

# Tidy the results
summary_tbl_FigS3 <- logit_summary_table(logit_tmpsmx_any_ICI, final_dt_tmpsmx_any)  # Converts log-odds to ORs

# Create gt table by renaming variables to human-readable names
FigureS3 <- summary_tbl_FigS3 %>%
  mutate(
    Variable = case_when(
      Variable == "age_group" ~ "Age (Years)",
      Variable == "sex" ~ "Sex",
      Variable == "NumDrugs" ~ "Number of Concomittant Drugs",
      Variable == "ICI" ~ "ICI Exposure",
      Variable == "cancer_flag" ~ "Cancer Diagnosis",
      TRUE ~ Variable
    ),
    Level = case_when(
      Level == "M" ~ "Male",
      Level == "F" ~ "Female",
      Level == "F (ref)" ~ "Female (ref)",
      Level == "NS" ~ "Not Specified",
      Level == "0 (ref)" ~ "No (ref)",
      Level == "1" ~ "Yes",
      Level == "0" ~ "No",
      TRUE ~ Level
    )
  ) %>%
  gt(groupname_col = "Variable") %>%
  tab_header(
    title = md("**SJS/TEN in Patients Exposed to TMP-SMX**"),
    subtitle = "Multivariable Logistic Regression, FAERS 2013-2023"
  ) %>%
  cols_label(
    Level = "",
    N = "N (Total)",
    SJS_TEN_Cases = "N (SJS/TEN)",
    OR = "Odds Ratio",
    CI = "95% CI",
    P = "P-value"
  ) %>%
  sub_missing(columns = everything(), missing_text = "-") %>%
  tab_options(
    table.font.names = "Times",
    table.font.size = 12,
    heading.title.font.size = 14,
    heading.subtitle.font.size = 12,
    row_group.as_column = TRUE,
    data_row.padding = px(4)
  )

# Display the figure
FigureS3
```

## FIGURE 3 - DEVELOPMENT AND MORTALITY LOGIT, ENTIRE FAERS

This figure is a pair of logistic regressions for development and
mortality of SJS/TEN in the entire dataset. The first (development) one
will use all of final_dt, and have SJS/TEN as the outcome. The second
(mortality) will use final_SJSTEN_dt (final_dt restricted to just
SJS/TEN), and use death as the outcome. For covariates, we will use ICI,
strong culprit exposure, weak culprit exposure, age, sex, and cancer
diagnosis. We will create both logistic regressions then unify them into
one table.

```{r}
## Conduct the logistic regression
# Development of SJS/TEN, logistic regression
Fig3_development <- glm(SJSTEN ~ ICI + StrongCulprit + WeakCulprit + age_group + sex + cancer_flag + NumDrugs, data = final_dt, family = binomial())

# Mortality in SJS/TEN patients, logistic regression
Fig3_mortality <- glm(Death ~ ICI + StrongCulprit + WeakCulprit + age_group + sex + cancer_flag + NumDrugs, data = final_SJSTEN_dt, family = binomial())
```

Now create the tables:

```{r}
## Generate summary tables
Fig3_development_tbl <- logit_summary_table(Fig3_development, final_dt)
Fig3_development_tbl

Fig3_mortality_tbl <- logit_summary_table(Fig3_mortality, final_SJSTEN_dt)
Fig3_mortality_tbl

## Combine the tables
Fig3_combined_tbl <- Fig3_development_tbl %>%
    full_join(Fig3_mortality_tbl, by = c("Variable", "Level"), suffix = c("_dev", "_mort"))

# Format the table into gt format
Figure3_final <- Fig3_combined_tbl %>%
    mutate(
      Variable = case_when(
        Variable %in% c("age_group") ~ "Age (Years)",
        Variable %in% c("ICI", "ICI_Count_Factor") ~ "ICI Exposure",
        Variable == "sex" ~ "Sex",
        Variable == "NumDrugs" ~ "Number of Concomittant Drugs",
        Variable == "cancer_flag" ~ "Cancer Diagnosis",
        Variable == "WeakCulprit" ~ "Exposure to Weak Culprit",
        Variable == "StrongCulprit" ~ "Exposure to Strong Culprit",
        TRUE ~ Variable
      ),
      Level = case_when(
        Variable == "Number of Concomittant Drugs" ~ "Per Unit Increase",
        Level == "0 (ref)" ~ "No (ref)",
         Level == "1" ~ "Yes",
        Level == "F (ref)" ~ "Female (ref)",
        Level == "M" ~ "Male",
        Level == "NS" ~ "Not Specified",
        is.na(Level) | Level == "" ~ "-",
        TRUE ~ Level
      )
    ) %>%
    arrange(Variable, desc(str_detect(Level, "\\(ref\\)")))

# Create gt table
Figure3_final <- Figure3_final %>%
    select(
      Variable, Level,
      `N (Total)` = N_dev,
      `N (SJS/TEN)` = SJS_TEN_Cases_dev,
      `N (Deaths)` = SJS_TEN_Cases_mort,
      `OR (Dev)` = OR_dev,
      `95% CI (Dev)` = CI_dev,
      `P (Dev)` = P_dev,
      `OR (Mort)` = OR_mort,
      `95% CI (Mort)` = CI_mort,
      `P (Mort)` = P_mort
    ) %>%
    gt(groupname_col = "Variable") %>%
    tab_header(
      title = md("**Predictors of SJS/TEN Development and Mortality**"), subtitle = "Multivariate Logistic Regression on All FAERS Reports, 2013-2023"
    ) %>%
    cols_align(align = "left", columns = Level) %>%
    cols_width(everything() ~ px(100)) %>%
    tab_options(
    table.font.names = "Times",
    table.font.size = 12,
    heading.title.font.size = 14,
    heading.subtitle.font.size = 12,
    row_group.as_column = TRUE,
    data_row.padding = px(4)
    ) %>%
    sub_missing(columns = everything(), missing_text = "-")

# Show the table
Figure3_final
```

## FIGURE S4 - RESTRICTED LOGIT (EXPOSURE VARIABLES REMOVED)

One of the unusual findings in Figure 3 is that cancer is negatively
associated with the development of SJS/TEN. Previous literature has
reported the opposite association, however those studies don't account
for other exposures a lot of the time. We can figure this out by
removing variables by logistic regression (namely, the culprit exposure
and ICI variables), and seeing if this association changes.

```{r}
## Conduct the logistic regression (on fewer variables)
# Development of SJS/TEN, logistic regression
FigS4_development <- glm(SJSTEN ~ age_group + sex + cancer_flag + NumDrugs, data = final_dt, family = binomial())

# Mortality in SJS/TEN patients, logistic regression
FigS4_mortality <- glm(Death ~ age_group + sex + cancer_flag + NumDrugs, data = final_SJSTEN_dt, family = binomial())
```

Now create the tables:

```{r}
## Generate summary tables
FigS4_development_tbl <- logit_summary_table(FigS4_development, final_dt)
FigS4_development_tbl

FigS4_mortality_tbl <- logit_summary_table(FigS4_mortality, final_SJSTEN_dt)
FigS4_mortality_tbl

## Combine the tables
FigS4_combined_tbl <- FigS4_development_tbl %>%
    full_join(FigS4_mortality_tbl, by = c("Variable", "Level"), suffix = c("_dev", "_mort"))

# Format the table into gt format
FigureS4_final <- FigS4_combined_tbl %>%
    mutate(
      Variable = case_when(
        Variable %in% c("age_group") ~ "Age (Years)",
        Variable == "sex" ~ "Sex",
        Variable == "NumDrugs" ~ "Number of Concomittant Drugs",
        Variable == "cancer_flag" ~ "Cancer Diagnosis",
        TRUE ~ Variable
      ),
      Level = case_when(
        Variable == "Number of Concomittant Drugs" ~ "Per Unit Increase",
        Level == "0 (ref)" ~ "No (ref)",
         Level == "1" ~ "Yes",
        Level == "F (ref)" ~ "Female (ref)",
        Level == "M" ~ "Male",
        Level == "NS" ~ "Not Specified",
        is.na(Level) | Level == "" ~ "-",
        TRUE ~ Level
      )
    ) %>%
    arrange(Variable, desc(str_detect(Level, "\\(ref\\)")))

# Create gt table
FigureS4_final <- FigureS4_final %>%
    select(
      Variable, Level,
      `N (Total)` = N_dev,
      `N (SJS/TEN)` = SJS_TEN_Cases_dev,
      `N (Deaths)` = SJS_TEN_Cases_mort,
      `OR (Dev)` = OR_dev,
      `95% CI (Dev)` = CI_dev,
      `P (Dev)` = P_dev,
      `OR (Mort)` = OR_mort,
      `95% CI (Mort)` = CI_mort,
      `P (Mort)` = P_mort
    ) %>%
    gt(groupname_col = "Variable") %>%
    tab_header(
      title = md("**Predictors of SJS/TEN Development and Mortality in FAERS**"), subtitle = "Multivariate Logistic Regression on All FAERS Reports, 2013–2023"
    ) %>%
    cols_align(align = "left", columns = Level) %>%
    cols_width(everything() ~ px(100)) %>%
    tab_options(
    table.font.names = "Times",
    table.font.size = 12,
    heading.title.font.size = 14,
    heading.subtitle.font.size = 12,
    row_group.as_column = TRUE,
    data_row.padding = px(4)
    ) %>%
    sub_missing(columns = everything(), missing_text = "-")

# Show the table
FigureS4_final
```

## FIGURE 4 - SMALL-MOLECULE LOGIT, DEVELOPMENT AND MORTALITY

This figure will show the same two multivariable logistic regression
models (development and mortality), restricted to SJS/TEN cases in which
a small molecule drug was listed as the primary suspect. For this
analysis, we will explicitly include the top 10 (by frequency) primary
suspect drugs as a covariate, and thus cannot use strong and weak
culprit flags as covariates. We will also use the number of ICI as a
covariate, rather than mere ICI exposure.

```{r}
# Subset final_dt to just those with small molecule culprits
final_dt_sm <- final_dt[drug_class == "Small Molecule"]

## Label the top 10 drugs in each cohort
## First, in the small molecule cohort
# Label the top 10 drugs
top_drugs <- final_dt_sm %>%
    filter(SJSTEN == 1) %>%
    count(drug_name, sort = TRUE) %>%
    slice_head(n = 10) %>%
    pull(drug_name)

# Add a drug_group column
final_dt_sm <- final_dt_sm %>%
    mutate(drug_group = ifelse(drug_name %in% top_drugs, drug_name, "Other"))

# Relevel the drug_group factor
final_dt_sm$drug_group <- relevel(as.factor(final_dt_sm$drug_group), ref = "Other")

# Subset this to SJS/TEN cases for the mortality logit
final_SJSTEN_dt_sm <- final_dt_sm[SJSTEN == 1]

# Run logistic models
Figure4_dev_model <- glm(SJSTEN ~ age_group + sex + NumDrugs + ICI_Count_Factor + drug_group, data = final_dt_sm, family = binomial)
Figure4_mort_model <- glm(Death ~ age_group + sex + NumDrugs + ICI_Count_Factor + drug_group, data = final_SJSTEN_dt_sm, family = binomial)
```

Now create the tables:

```{r}
## Generate summary tables
Fig4_development_tbl <- logit_summary_table(Figure4_dev_model, final_dt_sm)
Fig4_development_tbl

Fig4_mortality_tbl <- logit_summary_table(Figure4_mort_model, final_SJSTEN_dt_sm)
Fig4_mortality_tbl
```

And reformat it:

```{r}
## Combine the tables
Fig4_combined_tbl <- Fig4_development_tbl %>%
    full_join(Fig4_mortality_tbl, by = c("Variable", "Level"), suffix = c("_dev", "_mort"))

# Format the table into gt format
Figure4_final <- Fig4_combined_tbl %>%
    mutate(
      Variable = case_when(
        Variable == "age_group" ~ "Age (Years)",
        Variable == "sex" ~ "Sex",
        Variable == "NumDrugs" ~ "Number of Concomittant Drugs",
        Variable == "ICI_Count_Factor" ~ "ICI Exposure",
        Variable == "drug_group" ~ "Causative Drug",
        TRUE ~ Variable
      ),
      Level = case_when(
        Variable == "Number of Concomittant Drugs" ~ "Per Unit Increase",
        Level == "0 (ref)" ~ "No (ref)",
        Level == "1" ~ "Yes",
        Level == "F (ref)" ~ "Female (ref)",
        Level == "M" ~ "Male",
        Level == "NS" ~ "Not Specified",
        is.na(Level) | Level == "" ~ "-",
        TRUE ~ Level
      )
    ) %>%
    arrange(Variable, desc(str_detect(Level, "\\(ref\\)")))

# Create gt table
Figure4_final <- Figure4_final %>%
    select(
      Variable, Level,
      `N (Total)` = N_dev,
      `N (SJS/TEN)` = SJS_TEN_Cases_dev,
      `N (Deaths)` = SJS_TEN_Cases_mort,
      `OR (Dev)` = OR_dev,
      `95% CI (Dev)` = CI_dev,
      `P (Dev)` = P_dev,
      `OR (Mort)` = OR_mort,
      `95% CI (Mort)` = CI_mort,
      `P (Mort)` = P_mort
    ) %>%
    gt(groupname_col = "Variable") %>%
    tab_header(
      title = md("**Predictors of Small-Molecule-Induced SJS/TEN Development and Mortality**"), subtitle = "Multivariate Logistic Regression on FAERS, 2013-2023 (Restricted to Small Molecule Primary Suspects)"
    ) %>%
    cols_align(align = "left", columns = Level) %>%
    cols_width(everything() ~ px(100)) %>%
      tab_options(
      table.font.names = "Times",
      table.font.size = 12,
      heading.title.font.size = 14,
      heading.subtitle.font.size = 12,
      row_group.as_column = TRUE,
      data_row.padding = px(4)
    ) %>%
    sub_missing(columns = everything(), missing_text = "-")

# Show the table
Figure4_final
```

## FIGURE S5 - SMALL-MOLECULE LOGIT, CANCER PATIENTS ONLY, DEVELOPMENT AND MORTALITY

This figure is a repeat of Figure 4, but isolated to cancer patients.
The reason we used a sensitivity analysis, rather than including cancer
as a covariate in Figure 4, is that the causative drugs are very
different between cancer patients and non-cancer patients. We will again
show two multivariable logistic regression models, restricted to SJS/TEN
cases in which a small molecule drug was listed as the primary suspect.

```{r}
# Subset final_dt to just those with small molecule culprits
final_dt_sm_cancer <- final_dt[drug_class == "Small Molecule" & cancer_flag == 1]

## Label the top 10 drugs in each cohort
## First, in the small molecule cohort
# Label the top 10 drugs
top_drugs <- final_dt_sm_cancer %>%
    filter(SJSTEN == 1) %>%
    count(drug_name, sort = TRUE) %>%
    slice_head(n = 10) %>%
    pull(drug_name)

# Add a drug_group column
final_dt_sm_cancer <- final_dt_sm_cancer %>%
    mutate(drug_group = ifelse(drug_name %in% top_drugs, drug_name, "Other"))

# Relevel the drug_group factor
final_dt_sm_cancer$drug_group <- relevel(as.factor(final_dt_sm_cancer$drug_group), ref = "Other")

# Subset this to SJS/TEN cases for the mortality logit
final_SJSTEN_dt_sm_cancer <- final_dt_sm_cancer[SJSTEN == 1]

# Run logistic models
FigureS5_dev_model <- glm(SJSTEN ~ age_group + sex + NumDrugs + ICI_Count_Factor + drug_group, data = final_dt_sm_cancer, family = binomial)
FigureS5_mort_model <- glm(Death ~ age_group + sex + NumDrugs + ICI_Count_Factor + drug_group, data = final_SJSTEN_dt_sm_cancer, family = binomial)
```

Now create the tables:

```{r}
## Generate summary tables
FigS5_development_tbl <- logit_summary_table(FigureS5_dev_model, final_dt_sm_cancer)
FigS5_development_tbl

FigS5_mortality_tbl <- logit_summary_table(FigureS5_mort_model, final_SJSTEN_dt_sm_cancer)
FigS5_mortality_tbl
```

And reformat it:

```{r}
## Combine the tables
FigS5_combined_tbl <- FigS5_development_tbl %>%
    full_join(FigS5_mortality_tbl, by = c("Variable", "Level"), suffix = c("_dev", "_mort"))

# Format the table into gt format
FigureS5_final <- FigS5_combined_tbl %>%
    mutate(
      Variable = case_when(
        Variable == "age_group" ~ "Age (Years)",
        Variable == "sex" ~ "Sex",
        Variable == "NumDrugs" ~ "Number of Concomittant Drugs",
        Variable == "ICI_Count_Factor" ~ "ICI Exposure",
        Variable == "drug_group" ~ "Causative Drug",
        TRUE ~ Variable
      ),
      Level = case_when(
        Variable == "Number of Concomittant Drugs" ~ "Per Unit Increase",
        Level == "0 (ref)" ~ "No (ref)",
        Level == "1" ~ "Yes",
        Level == "F (ref)" ~ "Female (ref)",
        Level == "M" ~ "Male",
        Level == "NS" ~ "Not Specified",
        is.na(Level) | Level == "" ~ "-",
        TRUE ~ Level
      )
    ) %>%
    arrange(Variable, desc(str_detect(Level, "\\(ref\\)")))

# Create gt table
FigureS5_final <- FigureS5_final %>%
    select(
      Variable, Level,
      `N (Total)` = N_dev,
      `N (SJS/TEN)` = SJS_TEN_Cases_dev,
      `N (Deaths)` = SJS_TEN_Cases_mort,
      `OR (Dev)` = OR_dev,
      `95% CI (Dev)` = CI_dev,
      `P (Dev)` = P_dev,
      `OR (Mort)` = OR_mort,
      `95% CI (Mort)` = CI_mort,
      `P (Mort)` = P_mort
    ) %>%
    gt(groupname_col = "Variable") %>%
    tab_header(
      title = md("**Predictors of Small-Molecule-Induced SJS/TEN Development and Mortality (Cancer Patients)**"), subtitle = "Multivariate Logistic Regression on FAERS, 2013-2023 (Restricted to Small Molecule Primary Suspects With Cancer Diagnosis)"
    ) %>%
    cols_align(align = "left", columns = Level) %>%
    cols_width(everything() ~ px(100)) %>%
      tab_options(
      table.font.names = "Times",
      table.font.size = 12,
      heading.title.font.size = 14,
      heading.subtitle.font.size = 12,
      row_group.as_column = TRUE,
      data_row.padding = px(4)
    ) %>%
    sub_missing(columns = everything(), missing_text = "-")

# Show the table
FigureS5_final
```

## FIGURE S6 - ICI-INDUCED SJS/TEN LOGIT

Figure S6 presents the multivariate logistic regression results for
predictors of ICI-induced SJS/TEN development and mortality. The
analysis is restricted to cases where an immune checkpoint inhibitor
(ICI) is the primary suspect drug. Covariates in the model include age
group, sex, number of concomitant drugs, and culprit drug
classification.

Odds ratios and 95% confidence intervals are shown for both the
likelihood of developing SJS/TEN and the risk of mortality among those
cases. Scientific notation is selectively applied for values exceeding
999 to improve readability. Unlike the other logits, categorical
variables were explicitly re-coded so that the "Unknown" age group
serves as the reference level, because the 0-20 level has no entries.

```{r}
# Subset final_dt to only cases in which ICI are the primary suspect
final_dt_ici_cases_only <- final_dt[drug_concept_id %in% ici_list]
final_dt_ici_cases_only$drug_name <- factor(final_dt_ici_cases_only$drug_name)

# Subset final_dt to only cases in which ICI are the primary suspect
final_dt_ici_cases_only$age_group <- relevel(final_dt_ici_cases_only$age_group, ref = "Unknown")

# Subset this to SJS/TEN cases for the mortality logit
final_SJSTEN_dt_ici_cases_only <- final_dt_ici_cases_only[SJSTEN == 1]

# Create model
FigureS6_dev_model <- glm(SJSTEN ~ age_group + sex + StrongCulprit + WeakCulprit + drug_name + NumDrugs,
family = binomial(), data = final_dt_ici_cases_only)
FigureS6_mort_model <- glm(Death ~ age_group + sex + StrongCulprit + WeakCulprit + drug_name + NumDrugs, data = final_SJSTEN_dt_ici_cases_only, family = binomial)
```

Then generate the tables

```{r}
## Generate summary tables
FigS6_development_tbl <- logit_summary_table(FigureS6_dev_model, final_dt_ici_cases_only)
FigS6_development_tbl

FigS6_mortality_tbl <- logit_summary_table(FigureS6_mort_model, final_SJSTEN_dt_ici_cases_only)
FigS6_mortality_tbl
```

And reformat it:

```{r}
## Combine the tables
FigS6_combined_tbl <- FigS6_development_tbl %>%
    full_join(FigS6_mortality_tbl, by = c("Variable", "Level"), suffix = c("_dev", "_mort"))

# Rename variables
FigureS6_final <- FigS6_combined_tbl %>%
    mutate(
      Variable = case_when(
        Variable == "age_group" ~ "Age (Years)",
        Variable == "sex" ~ "Sex",
        Variable == "NumDrugs" ~ "Number of Concomittant Drugs",
        Variable == "WeakCulprit" ~ "Exposure to Weak Culprit",
        Variable == "StrongCulprit" ~ "Exposure to Strong Culprit",
        Variable == "drug_name" ~ "Causative Drug",
        TRUE ~ Variable
      ),
      Level = case_when(
        Variable == "Number of Concomittant Drugs" ~ "Per Unit Increase",
        Level == "0 (ref)" ~ "No (ref)",
        Level == "1" ~ "Yes",
        Level == "F (ref)" ~ "Female (ref)",
        Level == "M" ~ "Male",
        Level == "NS" ~ "Not Specified",
        is.na(Level) | Level == "" ~ "-",
        TRUE ~ Level
      )
    ) %>%
    arrange(Variable, desc(str_detect(Level, "\\(ref\\)")))

# Apply formatting
FigureS6_final <- FigureS6_final %>%
  select(
    Variable, Level,
    `N (Total)` = N_dev,
    `N (SJS/TEN)` = SJS_TEN_Cases_dev,
    `N (Deaths)` = SJS_TEN_Cases_mort,
    `OR (Dev)` = OR_dev,
    `95% CI (Dev)` = CI_dev,
    `P (Dev)` = P_dev,
    `OR (Mort)` = OR_mort,
    `95% CI (Mort)` = CI_mort,
    `P (Mort)` = P_mort
  ) %>%
  mutate(
    `OR (Dev)` = ifelse(
      `OR (Dev)` > 999,
      formatC(`OR (Dev)`, format = "e", digits = 2),
      formatC(`OR (Dev)`, format = "f", digits = 2)
    ),
    `OR (Mort)` = ifelse(
      `OR (Mort)` > 999,
      formatC(`OR (Mort)`, format = "e", digits = 2),
      formatC(`OR (Mort)`, format = "f", digits = 2)
    ),
    `95% CI (Dev)` = map_chr(`95% CI (Dev)`, format_ci),
    `95% CI (Mort)` = map_chr(`95% CI (Mort)`, format_ci)
  ) %>%
  gt(groupname_col = "Variable") %>%
  tab_header(
    title = md("**Predictors of ICI-Induced SJS/TEN Development and Mortality**"),
    subtitle = "Multivariate Logistic Regression on FAERS, 2013–2023 (Restricted to ICI Primary Suspects)"
  ) %>%
  cols_align(align = "left", columns = Level) %>%
  cols_width(everything() ~ px(100)) %>%
  tab_options(
    table.font.names = "Times",
    table.font.size = 12,
    heading.title.font.size = 14,
    heading.subtitle.font.size = 12,
    row_group.as_column = TRUE,
    data_row.padding = px(4)
  ) %>%
  sub_missing(columns = everything(), missing_text = "-")


# Show the table
FigureS6_final
```

## FIGURE S7 - ICI-INDUCED (GROUPED) SJS/TEN LOGIT

We repeat the construction of figure S6, BUT, with grouping the
causative ICI by mechanism. If the patient is exposed to more than one
ICI and they have different targets, then it's coded as "multiple".

```{r}
# Create new ICI category
final_dt_ici_cases_only <- final_dt_ici_cases_only %>%
  mutate(
    ICI_causative = case_when(
      rowSums(across(c(ICI_PD1, ICI_PDL1, ICI_CTLA4)) == 1, na.rm = TRUE) >= 2 ~ "Multiple",
      ICI_PD1 == 1 ~ "PD-1",
      ICI_PDL1 == 1 ~ "PD-L1",
      ICI_CTLA4 == 1 ~ "CTLA-4",
      TRUE ~ "Other"
    )
  )

final_dt_ici_cases_only$ICI_causative <- factor(
  final_dt_ici_cases_only$ICI_causative,
  levels = c("CTLA-4", "PD-L1", "PD-1", "Multiple"),
  exclude = NULL
)

final_dt_ici_cases_only$ICI_causative <- relevel(final_dt_ici_cases_only$ICI_causative, ref = "CTLA-4")


# Subset this to SJS/TEN cases for the mortality logit
final_SJSTEN_dt_ici_cases_only <- final_dt_ici_cases_only[SJSTEN == 1]

# Create model
FigureS7_dev_model <- glm(SJSTEN ~ age_group + sex + StrongCulprit + WeakCulprit + ICI_causative + NumDrugs,
family = binomial(), data = final_dt_ici_cases_only)
FigureS7_mort_model <- glm(Death ~ age_group + sex + StrongCulprit + WeakCulprit + ICI_causative + NumDrugs, data = final_SJSTEN_dt_ici_cases_only, family = binomial)
```

Then generate the tables

```{r}
## Generate summary tables
FigS7_development_tbl <- logit_summary_table(FigureS7_dev_model, final_dt_ici_cases_only)
FigS7_development_tbl

FigS7_mortality_tbl <- logit_summary_table(FigureS7_mort_model, final_SJSTEN_dt_ici_cases_only)
FigS7_mortality_tbl
```

And reformat it:

```{r}
## Combine the tables
FigS7_combined_tbl <- FigS7_development_tbl %>%
    full_join(FigS7_mortality_tbl, by = c("Variable", "Level"), suffix = c("_dev", "_mort"))

FigureS7_final <- FigS7_combined_tbl %>%
    mutate(
      Variable = case_when(
        Variable == "age_group" ~ "Age (Years)",
        Variable == "sex" ~ "Sex",
        Variable == "NumDrugs" ~ "Number of Concomittant Drugs",
        Variable == "WeakCulprit" ~ "Exposure to Weak Culprit",
        Variable == "StrongCulprit" ~ "Exposure to Strong Culprit",
        Variable == "ICI_causative" ~ "Causative ICI Class",
        TRUE ~ Variable
      ),
      Level = case_when(
        Variable == "Number of Concomittant Drugs" ~ "Per Unit Increase",
        Level == "0 (ref)" ~ "No (ref)",
        Level == "1" ~ "Yes",
        Level == "F (ref)" ~ "Female (ref)",
        Level == "M" ~ "Male",
        Level == "NS" ~ "Not Specified",
        is.na(Level) | Level == "" ~ "-",
        TRUE ~ Level
      )
    ) %>%
    arrange(Variable, desc(str_detect(Level, "\\(ref\\)")))

# Apply formatting
FigureS7_final <- FigureS7_final %>%
  select(
    Variable, Level,
    `N (Total)` = N_dev,
    `N (SJS/TEN)` = SJS_TEN_Cases_dev,
    `N (Deaths)` = SJS_TEN_Cases_mort,
    `OR (Dev)` = OR_dev,
    `95% CI (Dev)` = CI_dev,
    `P (Dev)` = P_dev,
    `OR (Mort)` = OR_mort,
    `95% CI (Mort)` = CI_mort,
    `P (Mort)` = P_mort
  ) %>%
  mutate(
    `OR (Dev)` = ifelse(
      `OR (Dev)` > 999,
      formatC(`OR (Dev)`, format = "e", digits = 2),
      formatC(`OR (Dev)`, format = "f", digits = 2)
    ),
    `OR (Mort)` = ifelse(
      `OR (Mort)` > 999,
      formatC(`OR (Mort)`, format = "e", digits = 2),
      formatC(`OR (Mort)`, format = "f", digits = 2)
    ),
    `95% CI (Dev)` = map_chr(`95% CI (Dev)`, format_ci),
    `95% CI (Mort)` = map_chr(`95% CI (Mort)`, format_ci)
  ) %>%
  gt(groupname_col = "Variable") %>%
  tab_header(
    title = md("**Predictors of ICI-Induced SJS/TEN Development and Mortality**"),
    subtitle = "Multivariate Logistic Regression on FAERS, 2013–2023 (Restricted to ICI Primary Suspects, Grouped by Class)"
  ) %>%
  cols_align(align = "left", columns = Level) %>%
  cols_width(everything() ~ px(100)) %>%
  tab_options(
    table.font.names = "Times",
    table.font.size = 12,
    heading.title.font.size = 14,
    heading.subtitle.font.size = 12,
    row_group.as_column = TRUE,
    data_row.padding = px(4)
  ) %>%
  sub_missing(columns = everything(), missing_text = "-")


# Show the table
FigureS7_final
```

## FIGURE 5 - LATENCY, ALL SJS/TEN

This section generates latency models and Kaplan-Meier curves. We will
generate three figures. Each figure will have a Kaplan-Meier curve as
the "A" panel, and a table of Cox proportional hazards model results in
the "B" panel.

-   One of the whole dataset, using ICI exposure, cancer diagnosis, and
    culprit exposures as covariates (similar to Figure 3)

-   One of small-molecule causes, using ICI exposure and causative drug
    as covariates (similar to Figure 4)

-   One of small-molecule causes isolated to cancer diagnoses (similar
    to Figure S5).

First, merge TTE information with final_dt.

```{r}
# Create version of final_dt with TTE information
final_dt_tte <- merge(
   final_dt[, .(compositeid, age_group, sex, NumDrugs, ICI_Count_Factor, drug_name, drug_class, SJSTEN, ICI, cancer_flag, StrongCulprit, WeakCulprit)],
  drug_TTE_sjsten[, .(compositeid, TTE)],
    by = "compositeid")
  
# Filter down to SJS/TEN cases with intact TTE information
sjsten_tte_data <- final_dt_tte %>% filter(SJSTEN == 1, !is.na(TTE))
```

Now generate the additional two cohorts.

```{r}
#Small-molecule cohort (Figure S8)
sjsten_tte_data_sm <- sjsten_tte_data[drug_class == "Small Molecule"]
                                      
#Small-molecule cancer cohort (Figure S9)
sjsten_tte_data_sm_cancer <- sjsten_tte_data_sm[cancer_flag == 1]
```

Create the drug_group factor for each cohort.

```{r}
## Fix the drug_group for the whole cohort
# Do the top drugs thing
top_drugs <- sjsten_tte_data %>% 
    count(drug_name, sort = TRUE) %>%
    slice_head(n = 10) %>%
    pull(drug_name)

# Create the drug_group factor
sjsten_tte_data <- sjsten_tte_data %>%
    mutate(drug_group = ifelse(drug_name %in% top_drugs, drug_name, "Other"))

# Relevel the top drugs so "Other" is the reference  
sjsten_tte_data$drug_group <- relevel(as.factor(sjsten_tte_data$drug_group), ref = "Other")

## Fix the drug_group for the sm cohort
# Do the top drugs thing
top_drugs_sm <- sjsten_tte_data_sm %>% 
    count(drug_name, sort = TRUE) %>%
    slice_head(n = 10) %>%
    pull(drug_name)

# Create the drug_group factor
sjsten_tte_data_sm <- sjsten_tte_data_sm %>%
    mutate(drug_group = ifelse(drug_name %in% top_drugs_sm, drug_name, "Other"))

# Relevel the top drugs so "Other" is the reference  
sjsten_tte_data_sm$drug_group <- relevel(as.factor(sjsten_tte_data_sm$drug_group), ref = "Other")

## Fix the drug_group for the sm+cancer cohort
# Do the top drugs thing
top_drugs_sm_cancer <- sjsten_tte_data_sm_cancer %>% 
    count(drug_name, sort = TRUE) %>%
    slice_head(n = 10) %>%
    pull(drug_name)

# Create the drug_group factor
sjsten_tte_data_sm_cancer <- sjsten_tte_data_sm_cancer %>%
    mutate(drug_group = ifelse(drug_name %in% top_drugs_sm_cancer, drug_name, "Other"))

# Relevel the top drugs so "Other" is the reference  
sjsten_tte_data_sm_cancer$drug_group <- relevel(as.factor(sjsten_tte_data_sm_cancer$drug_group), ref = "Other")
```

Then use these to generate the figures.

```{r}
# Create a survival object  
surv_obj_Fig5 <- Surv(time = sjsten_tte_data$TTE, event = rep(1, nrow(sjsten_tte_data)))
  
# Create the model
Figure5_model <- coxph(
    surv_obj_Fig5 ~ ICI + age_group + sex + NumDrugs + StrongCulprit + WeakCulprit + cancer_flag,
    data = sjsten_tte_data, model = TRUE
)

# Extract the results
Fig5_results_tbl <- cox_summary_table(Figure5_model, sjsten_tte_data)
Fig5_results_tbl
```

Then we can turn this into a gt table.

```{r}
# Format the table like the others
Figure5B_final <- Fig5_results_tbl %>%
  mutate(
    Variable = case_when(
      Variable == "age_group" ~ "Age (Years)",
      Variable == "sex" ~ "Sex",
      Variable == "NumDrugs" ~ "Number of Concomitant Drugs",
      Variable == "drug_group" ~ "Causative Drug",
      Variable == "ICI" ~ "ICI Exposure",
      Variable == "cancer_flag" ~ "Cancer Diagnosis",
      Variable == "StrongCulprit" ~ "Exposure to Strong Culprit",
      Variable == "WeakCulprit" ~ "Exposure to Weak Culprit",
      TRUE ~ Variable
    ),
    Level = case_when(
      Variable == "Number of Concomitant Drugs" ~ "Per Unit Increase",
      Level == "0 (ref)" ~ "No (ref)",
      Level == "1" ~ "Yes",
      Level == "F (ref)" ~ "Female (ref)",
      Level == "M" ~ "Male",
      Level == "NS" ~ "Not Specified",
      is.na(Level) | Level == "" ~ "-",
      TRUE ~ Level
    )
  ) %>%
  arrange(Variable, desc(str_detect(Level, "\\(ref\\)")))  # Ensure reference levels show first

# Create gt table
Figure5B_final <- Figure5B_final %>%
  select(
    Variable, Level,
    `N` = N,
    `HR` = HR,
    `95% CI` = CI,
    `P` = P
  ) %>%
  gt(groupname_col = "Variable") %>%
  tab_header(
    title = md("**Latency Model - SJS/TEN**"),
    subtitle = "Cox Proportional Hazards Model for Time-to-Onset of SJS/TEN"
  ) %>%
  cols_align(align = "left", columns = Level) %>%
  cols_width(everything() ~ px(100)) %>%
  tab_options(
    table.font.names = "Times",
    table.font.size = 12,
    heading.title.font.size = 14,
    heading.subtitle.font.size = 12,
    row_group.as_column = TRUE,
    data_row.padding = px(4)
  ) %>%
  sub_missing(columns = everything(), missing_text = "-")

# Display table
Figure5B_final
```

Then make the Kaplan-Meier curves also.

```{r}
# Re-create a variable to combine ICI and cancer, for ease
sjsten_tte_data <- sjsten_tte_data %>%
  mutate(KM_group = case_when(
    ICI == 1 ~ "ICI",
    ICI == 0 & cancer_flag == 0 ~ "No ICI, No Cancer",
    ICI == 0 & cancer_flag == 1 ~ "No ICI, Cancer"
  )) %>%
  mutate(KM_group = factor(KM_group, levels = c("No ICI, No Cancer", "No ICI, Cancer", "ICI")))

# Refit survival object with updated ICI levels
fit_km_Fig5A <- survfit(surv_obj_Fig5 ~ KM_group, data = sjsten_tte_data)

# Plot with reordered strata
Figure5A_plot <- ggsurvplot(
  fit_km_Fig5A,
  data = sjsten_tte_data,
  pval = TRUE,
  risk.table = TRUE,
  conf.int = TRUE,
  risk.table.col = "strata",
  legend.title = "Group",
  legend.labs = c("No ICI, No Cancer", "No ICI, Cancer", "ICI"),
  title = "Time to Onset in SJS/TEN",
  xlab = "Days to Onset",
  ylab = "Survival Probability",
  risk.table.height = 0.25,
  ggtheme = theme_minimal(base_size = 12),
  break.time.by = 20
)
```

Modify the plot to add a table:

```{r Figure5A, fig.width = 14, fig.height = 8, out.width = "100%"}
# 🧮 Create summary stats table
summary_table_5A <- sjsten_tte_data[, .(
  `Median [IQR], days` = sprintf(
    "%.1f [%.1f–%.1f]",
    median(TTE),
    quantile(TTE, 0.25),
    quantile(TTE, 0.75)
  ),
  `Mean [95%% CI], days` = {
    m <- mean(TTE)
    se <- sd(TTE) / sqrt(.N)
    sprintf("%.1f [%.1f–%.1f]", m, m - 1.96 * se, m + 1.96 * se)
  }
), by = KM_group]

# 🏷️ Rename for display
setnames(summary_table_5A, "KM_group", "Group")

# 📐 Reorder rows to match plot legend
summary_table_5A <- summary_table_5A[match(c("No ICI, No Cancer", "No ICI, Cancer", "ICI"), Group)]

# 🎨 Table theme (simple black text)
table_theme <- ttheme_default(
  core = list(fg_params = list(col = "black")),
  colhead = list(fg_params = list(fontface = 2))
)

# 🧩 Add table to main plot (not risk table)
Figure5A_plot$plot <- Figure5A_plot$plot +
  geom_table(
    data = data.frame(x = 80, y = 0.9),
    aes(x = x, y = y, label = list(summary_table_5A)),
    table.theme = table_theme,
    hjust = 0,
    vjust = 1
  )

# 📈 Final display
Figure5A_plot
```

## FIGURE S8 - LATENCY BY CAUSATIVE DRUG

```{r}
# ici_list is a named vector of ICI names (e.g., c("12345" = "nivolumab"))
ici_drug_names <- tolower(names(ici_list))

# Prepare data
km_data <- copy(sjsten_tte_data)

# Normalize drug_name to lowercase for matching
km_data[, drug_name_lower := tolower(drug_name)]

# Define causative group based on lowercase drug_name
km_data[, causative_group := fcase(
  drug_name_lower %in% ici_drug_names,      "ICI",
  drug_class == "Biologic",                 "Non-ICI Biologic",
  drug_class == "Small Molecule",           "Small Molecule"
)]

# Change to factor
km_data[, causative_group := factor(causative_group, levels = c("Small Molecule", "Non-ICI Biologic", "ICI"))]


# All rows are SJS/TEN cases, so event = 1
km_data[, event := 1L]

# Fit Cox model (Figure S8B)
FigureS8_model <- coxph(Surv(TTE, event) ~ causative_group + age_group + sex + cancer_flag + NumDrugs,
                   data = km_data)

# Summarize model and make a GT table
FigureS8_results_tbl <- cox_summary_table(FigureS8_model, km_data)
FigureS8_results_tbl
```

Then modify the table:

```{r}
# Format the table
FigureS8B_final <- FigureS8_results_tbl %>%
  mutate(
    Variable = case_when(
      Variable == "age_group" ~ "Age (Years)",
      Variable == "sex" ~ "Sex",
      Variable == "NumDrugs" ~ "Number of Concomitant Drugs",
      Variable == "causative_group" ~ "Causative Drug Class",
      Variable == "cancer_flag" ~ "Cancer Diagnosis",
      TRUE ~ Variable
    ),
    Level = case_when(
      Variable == "Number of Concomitant Drugs" ~ "Per Unit Increase",
      Level == "0 (ref)" ~ "No (ref)",
      Level == "1" ~ "Yes",
      Level == "F (ref)" ~ "Female (ref)",
      Level == "M" ~ "Male",
      Level == "NS" ~ "Not Specified",
      is.na(Level) | Level == "" ~ "-",
      TRUE ~ Level
    )
  ) %>%
  arrange(Variable, desc(str_detect(Level, "\\(ref\\)")))  # Ensure reference levels show first

# Create gt table
FigureS8B_final <- FigureS8B_final %>%
  select(
    Variable, Level,
    `N` = N,
    `HR` = HR,
    `95% CI` = CI,
    `P` = P
  ) %>%
  gt(groupname_col = "Variable") %>%
  tab_header(
    title = md("**Latency Model - SJS/TEN By Causative Drug Class**"),
    subtitle = "Cox Proportional Hazards Model for Time-to-Onset of SJS/TEN"
  ) %>%
  cols_align(align = "left", columns = Level) %>%
  cols_width(everything() ~ px(100)) %>%
  tab_options(
    table.font.names = "Times",
    table.font.size = 12,
    heading.title.font.size = 14,
    heading.subtitle.font.size = 12,
    row_group.as_column = TRUE,
    data_row.padding = px(4)
  ) %>%
  sub_missing(columns = everything(), missing_text = "-")

# Display table
FigureS8B_final
```

Perform Kaplan-Meier analysis between the 3 groups:

```{r}
# Fit Kaplan–Meier model
km_fit_figS8 <- survfit(Surv(TTE, event) ~ causative_group, data = km_data)

# Plot KM curves (Figure S8A)
FigureS8A_plot <- ggsurvplot(
  km_fit_figS8,
  data = km_data,
  pval = TRUE,
  conf.int = TRUE,
  risk.table = TRUE,
  risk.table.col = "strata",
  xlab = "Days to Onset",
  ylab = "Survival Probability",
  legend.title = "Causative Drug Group",
  legend.labs = c("Small Molecule", "Non-ICI Biologic", "ICI"),
  title = "Time to Onset in SJS/TEN by Causative Drug Class",
  risk.table.height = 0.25,
  ggtheme = theme_minimal(base_size = 12),
  break.time.by = 20
)
```

Now we'll add the summary table

```{r FigureS8A, fig.width = 14, fig.height = 8, out.width = "100%"}
# 🧮 Create the summary stats table
summary_table_S8A <- km_data[, .(
  `Median [IQR], days` = sprintf(
    "%.1f [%.1f–%.1f]",
    median(TTE),
    quantile(TTE, 0.25),
    quantile(TTE, 0.75)
  ),
  `Mean [95%% CI], days` = {
    m <- mean(TTE)
    se <- sd(TTE) / sqrt(.N)
    sprintf("%.1f [%.1f–%.1f]", m, m - 1.96 * se, m + 1.96 * se)
  }
), by = causative_group]

# 🏷️ Rename column for display
setnames(summary_table_S8A, "causative_group", "Causative Drug")

# 📐 Optional: order rows to match plot legend
summary_table_S8A <- summary_table_S8A[match(c("Small Molecule", "Non-ICI Biologic", "ICI"), `Causative Drug`)]

# 🎨 Basic table theme (black text)
table_theme <- ttheme_default(
  core = list(fg_params = list(col = "black")),
  colhead = list(fg_params = list(fontface = 2))
)

# 🧩 Add summary table to main plot (not risk table)
FigureS8A_plot$plot <- FigureS8A_plot$plot +
  geom_table(
    data = data.frame(x = 80, y = 0.9),
    aes(x = x, y = y, label = list(summary_table_S8A)),
    table.theme = table_theme,
    hjust = 0,
    vjust = 1
  )

# 📈 Print final figure (KM + risk table + stat table)
FigureS8A_plot
```

## FIGURE S9 - LATENCY, SMALL-MOLECULE SJS/TEN

This figure is the latency model for small-molecule SJS/TEN.

```{r}
# Create a survival object  
surv_obj_FigS9 <- Surv(time = sjsten_tte_data_sm$TTE, event = rep(1, nrow(sjsten_tte_data_sm)))
  
# Create the model
FigureS9_model <- coxph(
    surv_obj_FigS9 ~ ICI + age_group + sex + NumDrugs + drug_group,
    data = sjsten_tte_data_sm, model = TRUE
)

# Extract the results
FigS9_results_tbl <- cox_summary_table(FigureS9_model, sjsten_tte_data_sm)
FigS9_results_tbl
```

Then we can turn this into a gt table.

```{r}
# Format the table
FigureS9B_final <- FigS9_results_tbl %>%
  mutate(
    Variable = case_when(
      Variable == "age_group" ~ "Age (Years)",
      Variable == "sex" ~ "Sex",
      Variable == "NumDrugs" ~ "Number of Concomitant Drugs",
      Variable == "drug_group" ~ "Causative Drug",
      Variable == "ICI" ~ "ICI Exposure",
      TRUE ~ Variable
    ),
    Level = case_when(
      Variable == "Number of Concomitant Drugs" ~ "Per Unit Increase",
      Level == "0 (ref)" ~ "No (ref)",
      Level == "1" ~ "Yes",
      Level == "F (ref)" ~ "Female (ref)",
      Level == "M" ~ "Male",
      Level == "NS" ~ "Not Specified",
      is.na(Level) | Level == "" ~ "-",
      TRUE ~ Level
    )
  ) %>%
  arrange(Variable, desc(str_detect(Level, "\\(ref\\)")))  # Ensure reference levels show first

# Create gt table
FigureS9B_final <- FigureS9B_final %>%
  select(
    Variable, Level,
    `N` = N,
    `HR` = HR,
    `95% CI` = CI,
    `P` = P
  ) %>%
  gt(groupname_col = "Variable") %>%
  tab_header(
    title = md("**Latency Model - Small-Molecule-Induced SJS/TEN**"),
    subtitle = "Cox Proportional Hazards Model for Time-to-Onset of SJS/TEN"
  ) %>%
  cols_align(align = "left", columns = Level) %>%
  cols_width(everything() ~ px(100)) %>%
  tab_options(
    table.font.names = "Times",
    table.font.size = 12,
    heading.title.font.size = 14,
    heading.subtitle.font.size = 12,
    row_group.as_column = TRUE,
    data_row.padding = px(4)
  ) %>%
  sub_missing(columns = everything(), missing_text = "-")

# Display table
FigureS9B_final
```

Make the Kaplan-Meier Curve

```{r}
# Relevel ICI so that "1" (ICI) comes first and "0" (No ICI) second
sjsten_tte_data_sm$ICI <- factor(sjsten_tte_data_sm$ICI, levels = c(1, 0), labels = c("ICI", "No ICI"))

# Refit survival object with updated ICI levels
fit_km_FigS9A <- survfit(surv_obj_FigS9 ~ ICI, data = sjsten_tte_data_sm)

# Plot with reordered strata
FigureS9A_plot <- ggsurvplot(
  fit_km_FigS9A,
  data = sjsten_tte_data_sm,
  pval = TRUE,
  risk.table = TRUE,
  conf.int = TRUE,
  risk.table.col = "strata",
  legend.title = "ICI Exposure",
  legend.labs = c("ICI", "No ICI"),
  title = "Time to Onset in Small-Molecule-Induced SJS/TEN",
  xlab = "Days to Onset",
  ylab = "Survival Probability",
  risk.table.height = 0.25,
  ggtheme = theme_minimal(base_size = 12),
  break.time.by = 20
)
```

Modify the plots:

```{r FigureS9A, fig.width = 14, fig.height = 8, out.width = "100%"}
# 🧮 Create summary stats table
summary_table_S9A <- sjsten_tte_data_sm[, .(
  `Median [IQR], days` = sprintf(
    "%.1f [%.1f–%.1f]",
    median(TTE),
    quantile(TTE, 0.25),
    quantile(TTE, 0.75)
  ),
  `Mean [95%% CI], days` = {
    m <- mean(TTE)
    se <- sd(TTE) / sqrt(.N)
    sprintf("%.1f [%.1f–%.1f]", m, m - 1.96 * se, m + 1.96 * se)
  }
), by = ICI]

# 🏷️ Rename for display
setnames(summary_table_S9A, "ICI", "ICI Exposure")

# 📐 Optional: reorder rows to match legend
summary_table_S9A <- summary_table_S9A[match(c("ICI", "No ICI"), `ICI Exposure`)]

# 🎨 Table theme (simple black text)
table_theme <- ttheme_default(
  core = list(fg_params = list(col = "black")),
  colhead = list(fg_params = list(fontface = 2))
)

# 🧩 Add table to main plot (not risk table)
FigureS9A_plot$plot <- FigureS9A_plot$plot +
  geom_table(
    data = data.frame(x = 80, y = 0.9),
    aes(x = x, y = y, label = list(summary_table_S9A)),
    table.theme = table_theme,
    hjust = 0,
    vjust = 1
  )

# 📈 Final display
FigureS9A_plot
```

## FIGURE S10 - LATENCY, SMALL-MOLECULE SJS/TEN, CANCER PATIENTS

This figure is the latency model for the cancer cohort.

```{r}
# Create a survival object  
surv_obj_FigS10 <- Surv(time = sjsten_tte_data_sm_cancer$TTE, event = rep(1, nrow(sjsten_tte_data_sm_cancer)))
  
# Create the model
FigureS10_model <- coxph(
    surv_obj_FigS10 ~ ICI + age_group + sex + NumDrugs + drug_group,
    data = sjsten_tte_data_sm_cancer, model = TRUE
)

# Extract the results
FigS10_results_tbl <- cox_summary_table(FigureS10_model, sjsten_tte_data_sm_cancer)
FigS10_results_tbl
```

Then we can turn this into a gt table

```{r}
# Format the table
FigureS10B_final <- FigS10_results_tbl %>%
  mutate(
    Variable = case_when(
      Variable == "age_group" ~ "Age (Years)",
      Variable == "sex" ~ "Sex",
      Variable == "NumDrugs" ~ "Number of Concomitant Drugs",
      Variable == "drug_group" ~ "Causative Drug",
      Variable == "ICI" ~ "ICI Exposure",
      TRUE ~ Variable
    ),
    Level = case_when(
      Variable == "Number of Concomitant Drugs" ~ "Per Unit Increase",
      Level == "0 (ref)" ~ "No (ref)",
      Level == "1" ~ "Yes",
      Level == "F (ref)" ~ "Female (ref)",
      Level == "M" ~ "Male",
      Level == "NS" ~ "Not Specified",
      is.na(Level) | Level == "" ~ "-",
      TRUE ~ Level
    )
  ) %>%
  arrange(Variable, desc(str_detect(Level, "\\(ref\\)")))  # Ensure reference levels show first

# Create gt table
FigureS10B_final <- FigureS10B_final %>%
  select(
    Variable, Level,
    `N` = N,
    `HR` = HR,
    `95% CI` = CI,
    `P` = P
  ) %>%
  gt(groupname_col = "Variable") %>%
  tab_header(
    title = md("**Latency Model - Small-Molecule-Induced SJS/TEN in Cancer Patients**"),
    subtitle = "Cox Proportional Hazards Model for Time-to-Onset of SJS/TEN"
  ) %>%
  cols_align(align = "left", columns = Level) %>%
  cols_width(everything() ~ px(100)) %>%
  tab_options(
    table.font.names = "Times",
    table.font.size = 12,
    heading.title.font.size = 14,
    heading.subtitle.font.size = 12,
    row_group.as_column = TRUE,
    data_row.padding = px(4)
  ) %>%
  sub_missing(columns = everything(), missing_text = "-")

# Display table
FigureS10B_final
```

Then make the Kaplan-Meier curves also.

```{r}
# Relevel ICI so that "1" (ICI) comes first and "0" (No ICI) second
sjsten_tte_data_sm_cancer$ICI <- factor(sjsten_tte_data_sm_cancer$ICI, levels = c(1, 0), labels = c("ICI", "No ICI"))

# Refit survival object with updated ICI levels
fit_km_FigS10A <- survfit(surv_obj_FigS10 ~ ICI, data = sjsten_tte_data_sm_cancer)

# Plot with reordered strata
FigureS10A_plot <- ggsurvplot(
  fit_km_FigS10A,
  data = sjsten_tte_data_sm_cancer,
  pval = TRUE,
  risk.table = TRUE,
  conf.int = TRUE,
  risk.table.col = "strata",
  legend.title = "ICI Exposure",
  legend.labs = c("ICI", "No ICI"),
  title = "Time to Onset in Small-Molecule-Induced SJS/TEN (Cancer Patients)",
  xlab = "Days to Onset",
  ylab = "Survival Probability",
  risk.table.height = 0.25,
  ggtheme = theme_minimal(base_size = 12),
  break.time.by = 20
)
```
Then modify the plot to add a table:

```{r FigureS10A, fig.width = 14, fig.height = 8, out.width = "100%"}
# 🧮 Create summary stats table
summary_table_S10A <- sjsten_tte_data_sm_cancer[, .(
  `Median [IQR], days` = sprintf(
    "%.1f [%.1f–%.1f]",
    median(TTE),
    quantile(TTE, 0.25),
    quantile(TTE, 0.75)
  ),
  `Mean [95%% CI], days` = {
    m <- mean(TTE)
    se <- sd(TTE) / sqrt(.N)
    sprintf("%.1f [%.1f–%.1f]", m, m - 1.96 * se, m + 1.96 * se)
  }
), by = ICI]

# 🏷️ Rename for display
setnames(summary_table_S10A, "ICI", "ICI Exposure")

# 📐 Optional: reorder rows to match legend
summary_table_S10A <- summary_table_S10A[match(c("ICI", "No ICI"), `ICI Exposure`)]

# 🎨 Table theme (simple black text)
table_theme <- ttheme_default(
  core = list(fg_params = list(col = "black")),
  colhead = list(fg_params = list(fontface = 2))
)

# 🧩 Add table to main plot (not risk table)
FigureS10A_plot$plot <- FigureS10A_plot$plot +
  geom_table(
    data = data.frame(x = 80, y = 0.9),
    aes(x = x, y = y, label = list(summary_table_S10A)),
    table.theme = table_theme,
    hjust = 0,
    vjust = 1
  )

# 📈 Final display
FigureS10A_plot
```



## FIGURE S11 - VIF HEATMAP

Figure S11 presents a heatmap of generalized variance inflation factors
(GVIFs) across all multivariable models used in the study. Each row
corresponds to a covariate, and each column to a specific logistic
model. No GVIFs exceed typical concern thresholds (\>5), suggesting low
multicollinearity and good stability in the regression models.

```{r}
# 📊 Run and store VIFs
vif_list <- list(
  "Figure S7 – Mortality"     = car::vif(FigureS7_mort_model)[,1],
  "Figure S7 – Development"   = car::vif(FigureS7_dev_model)[,1],
  "Figure S6 – Mortality"     = car::vif(FigureS6_mort_model)[,1],
  "Figure S6 – Development"   = car::vif(FigureS6_dev_model)[,1],
  "Figure S5 – Mortality"     = car::vif(FigureS5_mort_model)[,1],
  "Figure S5 – Development"   = car::vif(FigureS5_dev_model)[,1],
  "Figure S4 – Mortality"     = car::vif(FigS4_mortality)[,1],
  "Figure S4 – Development"   = car::vif(FigS4_development)[,1],
  "Figure S3 (TMP-SMX)"       = car::vif(logit_tmpsmx_any_ICI)[,1],
  "Figure S2 (Allopurinol)"   = car::vif(logit_allopurinol_any_ICI)[,1],
  "Figure 4 – Mortality"      = car::vif(Figure4_mort_model)[,1],
  "Figure 4 – Development"    = car::vif(Figure4_dev_model)[,1],
  "Figure 3 – Mortality"      = car::vif(Fig3_mortality)[,1],
  "Figure 3 – Development"    = car::vif(Fig3_development)[,1]
)

# 🧼 Combine and clean
vif_df <- bind_rows(lapply(vif_list, function(x) as.data.frame(x) %>% tibble::rownames_to_column("Variable")),
                    .id = "Model") %>%
  rename(GVIF = x)

# 🔁 Custom variable label mapping
label_map <- c(
  "ICI" = "ICI Exposure (Binary)",
  "ICI_Count_Factor" = "ICI Exposure (Count)",
  "ICI_causative" = "ICI Class (Causative Agent)",
  "NumDrugs" = "Number of Concommitant Drugs",
  "StrongCulprit" = "Exposure to Strong Culprit",
  "WeakCulprit" = "Exposure to Weak Culprit",
  "age_group" = "Age",
  "cancer_flag" = "Cancer Diagnosis",
  "sex" = "Sex",
  "drug_group" = "Causative Drug",
  "drug_name" = "Causative Drug"
)

# Apply label map
vif_df$Variable <- dplyr::recode(vif_df$Variable, !!!label_map)

# 🧹 Remove duplicate “Causative Drug” rows from merged drug_name/group
vif_df <- vif_df %>% group_by(Model, Variable) %>% summarise(GVIF = max(GVIF), .groups = "drop")
```

Now plot it:

```{r Figure_S11_HeatmapGVIF, fig.width = 14, fig.height = 8, out.width = "100%"}
# Heatmap plot
FigureS11 <- ggplot(vif_df, aes(x = Model, y = Variable, fill = GVIF)) +
  geom_tile(color = "black", linewidth = 0.4) +
  scale_fill_distiller(palette = "Blues", direction = 1, name = "GVIF") +
  geom_text(aes(label = sprintf("%.3f", GVIF)), size = 3) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(face = "bold"),
    panel.grid = element_blank()
  ) +
  labs(
    title = "GVIF Heatmap Across Models",
    x = "Model",
    y = "Variable"
  )

# Plot it
FigureS11
```

# END
