---
title: "FAERS_TTE_Demography"
author: "Eric Mukherjee"
date: "2024-10-15"
output: html_document
---

```{r message=FALSE, warning=FALSE, include=FALSE, r,results='hide'}
library(data.table)
library(readr)
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(patchwork)
library(ggplot2)
library(ggpubr)
library(survminer)
library(survival)
```

##INTRODUCTION

Import the data into R:

```{r}
# Import drug, outcome, demographics, and ther tables
standard_case_drug <- read_csv("sanitized_drug_data.csv")
standard_case_outcome <- read_csv("sanitized_outcome_data.csv")
patient_demographics <- read_csv("sanitized_demographics.csv")
combined_ther <- read_csv("sanitized_ther.csv")

#Import the drug dictionary and outcome dictionary
#Import Updated Drug Dictionary
drug_dictionary <- read_csv("drug_dictionary.csv")
outcome_dictionary <- read_csv("sanitized_outcome_dictionary.csv")
```

For all analysis, we will stick with primary suspects (PS) only.

```{r}
# Filter down standard_case_drug to primary suspects only
standard_case_drug_PS <- standard_case_drug[standard_case_drug$role_cod == "PS",]
```

Next, we can left joint combined_ther to the standard_case_drug table:

```{r}
# Perform a left join using base R, ensuring that non-matching rows get NA
drug_exposure <- merge(standard_case_drug_PS, combined_ther, 
                       by = c("compositeid", "drug_seq"), 
                       all.x = TRUE)

# View the result
print(drug_exposure)
```

The patient demographics table has the start date of the given reaction. Now we can add that data to the outcome_data:

```{r}
# Perform a left join to add the demographic columns to outcome data
# Use merge with all.x = TRUE to preserve all rows of outcome data
updated_outcome_data <- merge(standard_case_outcome, patient_demographics, by = "compositeid", all.x = TRUE)

# View the updated data to verify the join
print(updated_outcome_data)
```

Now, the updated_outcome_data has the date of the reaction, and the drug_exposure data.table has the date of each drug exposure. We can now build a new table with the information we need:

```{r}
# Step 0: Make sure the relevant tables are formatted
updated_outcome_data <- as.data.table(updated_outcome_data)
standard_case_drug <- as.data.table(standard_case_drug)
drug_exposure <- as.data.table(drug_exposure)

# Step 1: Create a base table with all unique compositeid from standard_case_drug and updated_outcome_data
all_compositeids <- unique(c(standard_case_drug$compositeid, updated_outcome_data$compositeid))
final_table <- data.table(compositeid = all_compositeids)

# Step 2: Collapse reac_pt_list in updated_outcome_data for each compositeid with case-insensitive deduplication
outcome_data_collapsed <- updated_outcome_data[, .(
  reac_pt_list = paste(unique(tolower(pt)), collapse = "; ")
), by = compositeid]

# Step 3: Collapse drug names in standard_case_drug for each compositeid with case-insensitive deduplication
# Note this uses all drugs, not just PS!
drug_list_collapsed <- standard_case_drug[, .(
  drug_pt_list = paste(unique(tolower(drug_name)), collapse = "; ")), by = compositeid]

# Step 4: Extract only the PS role (causative drug) and its start date from drug_exposure
causative_drug_data <- drug_exposure[role_cod == "PS", .(compositeid, causative_drug = drug_name, start_dt)]

# Step 5: Perform separate left joins for each data source, ensuring all compositeids are preserved
final_table <- merge(final_table, patient_demographics, by = "compositeid", all.x = TRUE)
final_table <- merge(final_table, causative_drug_data, by = "compositeid", all.x = TRUE)
final_table <- merge(final_table, outcome_data_collapsed, by = "compositeid", all.x = TRUE)
final_table <- merge(final_table, drug_list_collapsed, by = "compositeid", all.x = TRUE)

# Deduplicate the final table
final_table <- unique(final_table)

# View the final table
print(final_table)

gc()  # Run garbage collection to free up memory
```
As a caveat - there are repeated rows with different start dates. This is likely due to how the administration is recorded. For example there's a patient who has 52 separate rows, with a vorinostat administration done repeatedly over a period of time. 

##FILTERING BY SCAR

Create a version that's filtered down to just DRESS:

```{r}
DRESS_IDs <- c(43053854, 37179294, 35104825)
compositeid_DRESS <- unique(updated_outcome_data[outcome_concept_id %in% DRESS_IDs, .(compositeid)])
final_table_DRESS_only <- final_table[compositeid %in% compositeid_DRESS$compositeid]
```

Lets do the same for SJS/TEN:

```{r}
SJSTEN_IDs <- c(36009724, 43010942, 36009754)
compositeid_SJSTEN <- unique(updated_outcome_data[outcome_concept_id %in% SJSTEN_IDs, .(compositeid)])
final_table_SJSTEN_only <- final_table[compositeid %in% compositeid_SJSTEN$compositeid]
```

And AGEP:

```{r}
AGEP_IDs <- c(36211150, 37179294)
compositeid_AGEP <- unique(updated_outcome_data[outcome_concept_id %in% AGEP_IDs, .(compositeid)])
final_table_AGEP_only <- final_table[compositeid %in% compositeid_AGEP$compositeid]
```

And GBFDE:

```{r}
GBFDE_IDs <- c(35327030)
compositeid_GBFDE <- unique(updated_outcome_data[outcome_concept_id %in% GBFDE_IDs, .(compositeid)])
final_table_GBFDE_only <- final_table[compositeid %in% compositeid_GBFDE$compositeid]
```

Also create one that contains all SCAR:

```{r}
SCAR_IDs <- c(43053854, 36009724, 43010942, 36009754, 36211150, 37179294, 35327030, 35104825)
compositeid_SCAR <- unique(updated_outcome_data[outcome_concept_id %in% SCAR_IDs, .(compositeid)])
final_table_SCAR_only <- final_table[compositeid %in% compositeid_SCAR$compositeid]
```

##ANALYZE TTE

###Preparing Data

For TTE analysis, we need to keep only the rows that have valid start_dt and event_dt, then calculate TTE.

```{r}
# Ensure final_table is a data.table
final_table <- as.data.table(final_table)

# Function to check if date is complete (YYYYMMDD format)
is_full_date <- function(date) {
  nchar(as.character(date)) == 8  # Only keep dates with 8 characters
}

# Step 1: Subset rows with complete dates
final_table_TTE <- final_table[
  is_full_date(start_dt) & is_full_date(event_dt)  # Filter for full dates only
]

# Step 2: Convert to Date format
final_table_TTE[, start_dt := as.Date(as.character(start_dt), format = "%Y%m%d")]
final_table_TTE[, event_dt := as.Date(as.character(event_dt), format = "%Y%m%d")]

# Step 3: Calculate TTE (Time-to-Event in days)
final_table_TTE[, TTE := as.numeric(event_dt - start_dt)]
```

Now subset this table by SCAR patients, for the initial analysis. And write them to disk.

```{r}
# Filter and write final tables to disk

# Write SJSTEN table
final_table_TTE_SJSTEN <- final_table_TTE[compositeid %in% compositeid_SJSTEN$compositeid]
write.csv(final_table_TTE_SJSTEN, "final_table_TTE_SJSTEN.csv", row.names = FALSE)

# Write DRESS table
final_table_TTE_DRESS <- final_table_TTE[compositeid %in% compositeid_DRESS$compositeid]
write.csv(final_table_TTE_DRESS, "final_table_TTE_DRESS.csv", row.names = FALSE)

# Write AGEP table
final_table_TTE_AGEP <- final_table_TTE[compositeid %in% compositeid_AGEP$compositeid]
write.csv(final_table_TTE_AGEP, "final_table_TTE_AGEP.csv", row.names = FALSE)

# Write GBFDE table
final_table_TTE_GBFDE <- final_table_TTE[compositeid %in% compositeid_GBFDE$compositeid]
write.csv(final_table_TTE_GBFDE, "final_table_TTE_GBFDE.csv", row.names = FALSE)

# Write SCAR table
final_table_TTE_SCAR <- final_table_TTE[compositeid %in% compositeid_SCAR$compositeid]
write.csv(final_table_TTE_SCAR, "final_table_TTE_SCAR.csv", row.names = FALSE)
```


###Per SCAR TTE

Now for plotting:

```{r}
# Define consistent colors for all SCAR types
scar_colors <- setNames(brewer.pal(5, "Set1"),  # Use Set1 with 5 distinct colors
                        c("SJS-TEN", "DRESS", "AGEP", "GBFDE", "Total SCAR"))

# Function to prepare data with TTE filtering and IQR-based outlier removal
prepare_TTE_data <- function(data, scar_name) {
  # Remove rows with missing TTE
  data <- data %>% filter(!is.na(TTE))
  
  # Step 1: Filter TTE between 0 and 180 days
  filtered_data <- data %>% filter(TTE > 0 & TTE <= 180)
  
  # Step 2: Calculate IQR for further outlier removal
  Q1 <- quantile(filtered_data$TTE, 0.25)
  Q3 <- quantile(filtered_data$TTE, 0.75)
  IQR <- Q3 - Q1
  
  # Define bounds for outlier detection
  lower_bound <- Q1 - 1.5 * IQR
  upper_bound <- Q3 + 1.5 * IQR
  
  # Step 3: Apply IQR-based outlier removal
  filtered_data <- filtered_data %>% filter(TTE >= lower_bound & TTE <= upper_bound)
  
  # Add SCAR group name as a column
  filtered_data <- filtered_data %>% mutate(SCAR = scar_name)
  
  ## REMOVE DUPLICATES AT THE END: Keep row with lowest TTE per compositeid
  filtered_data <- filtered_data %>%
    group_by(compositeid) %>%
    slice_min(order_by = TTE, with_ties = FALSE) %>%
    ungroup()  
  
  return(filtered_data)
}
```

Use this function to prepare the data.

```{r}
# Prepare data for each SCAR group
SJSTEN_data <- prepare_TTE_data(final_table_TTE_SJSTEN, "SJS-TEN")
DRESS_data <- prepare_TTE_data(final_table_TTE_DRESS, "DRESS")
AGEP_data <- prepare_TTE_data(final_table_TTE_AGEP, "AGEP")
GBFDE_data <- prepare_TTE_data(final_table_TTE_GBFDE, "GBFDE")
SCAR_data <- prepare_TTE_data(final_table_TTE_SCAR, "Total SCAR")

# Combine all datasets into one
combined_data <- bind_rows(SJSTEN_data, DRESS_data, AGEP_data, GBFDE_data, SCAR_data)

# Generate the violin plot
p4.1 <- ggplot(combined_data, aes(x = SCAR, y = TTE, fill = SCAR)) +
  geom_violin(trim = TRUE, alpha = 0.7) +   # Violin plot with trimming
  geom_boxplot(width = 0.1, outlier.shape = NA) +  # Boxplot inside violin for summary
  
  # Add pairwise comparisons with only significant stars and no brackets for ns
  stat_compare_means(
    method = "wilcox.test",  # Wilcoxon test for non-parametric data
    comparisons = list(
      c("SJS-TEN", "DRESS"), 
      c("DRESS", "AGEP"), 
      c("SJS-TEN", "AGEP"),
      c("GBFDE", "SJS-TEN"),  
      c("GBFDE", "DRESS")
    ),  # GBFDE vs AGEP removed
    label = "p.signif",  # Display only stars (***, **, *)
    hide.ns = TRUE,      # Hide non-significant comparisons entirely
    label.y = c(60, 65, 70, 75, 78)  # Adjust vertical positions
  ) +
  coord_cartesian(ylim = c(0, 80)) +  # Set maximum y-axis to 80
  theme_minimal() +
  labs(
    y = "Latency (Days)", x = NULL
  ) +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 10, angle = 45, hjust = 1),  # Rotate x-axis labels
    legend.position = "none"  # Hide legend since SCAR groups are already labeled
  ) +
  scale_fill_manual(values = scar_colors)  # Use scar_colors for fill

# Display the plot
print(p4.1)
```

GBFDE only has 5 rows pass filter, so it's harder to dive into that data. For reference, it's AGEP (n = 1541), DRESS (n = 5021), SJS-TEN (n = 7230), GBFDE (n = 5), and total SCAR (n = 13633)

Lets also do a TTE graph in case that looks better:

```{r}
# Create a new table
combined_data_noscar <- bind_rows(SJSTEN_data, DRESS_data, AGEP_data, GBFDE_data)
combined_data_noscar$SCAR <- as.factor(combined_data_noscar$SCAR)

# Fit survival curve by SCAR
fit_scar <- survfit(Surv(TTE) ~ SCAR, data = combined_data_noscar)

# Generate survival plot with custom colors
p4.1alt <- ggsurvplot(
  fit_scar,
  data = combined_data_noscar,
  pval = TRUE,
  pval.coord = c(50, 0.95),
  conf.int = FALSE,
  risk.table = TRUE,
  risk.table.height = 0.3,
  risk.table.fontsize = 3.5,  # Adjust font size for risk table
  surv.median.line = "hv",
  risk.table.col = "black",
  risk.table.y.text.col = TRUE,
  risk.table.y.text = TRUE,
  break.time.by = 10,
  ggtheme = theme_minimal(),
  xlim = c(0, 80),
  title = "",
  xlab = "",
  ylab = "",
  legend.title = "",
  legend.labs = levels(combined_data_noscar$SCAR),
  font.legend = list(size = 14, color = "black", face = "bold"),
  palette = scar_colors  # Apply custom colors
)
```
###Per Drug TTE

Now we want to compare TTE for various drugs within each SCAR.

```{r}
# Function to filter, clean, and prepare data for plotting
prepare_top_drugs_TTE <- function(data, scar_name, top_n = 5) {
  # Remove rows with missing TTE or causative drugs
  data <- data %>% filter(!is.na(TTE), !is.na(causative_drug))
  
  # Filter TTE to 0–180 days and remove outliers based on IQR
  filtered_data <- data %>% filter(TTE > 0 & TTE <= 180)
  Q1 <- quantile(filtered_data$TTE, 0.25)
  Q3 <- quantile(filtered_data$TTE, 0.75)
  IQR <- Q3 - Q1
  lower_bound <- Q1 - 1.5 * IQR
  upper_bound <- Q3 + 1.5 * IQR
  filtered_data <- filtered_data %>% filter(TTE >= lower_bound & TTE <= upper_bound)
  
  # Select top 'n' causative drugs by frequency
  top_drugs <- filtered_data %>%
    count(causative_drug, sort = TRUE) %>%
    top_n(top_n, n) %>%
    pull(causative_drug)
  
  # Filter dataset for only top drugs
  filtered_data <- filtered_data %>% filter(causative_drug %in% top_drugs)
  
  # Add SCAR group label
  filtered_data <- filtered_data %>% mutate(SCAR = scar_name)
  
  return(filtered_data)
}

```


Using this function, we can prepare the data:

```{r}
# Prepare data for each SCAR group
SJS_TEN_TTE <- prepare_top_drugs_TTE(final_table_TTE_SJSTEN, "SJS-TEN", top_n = 40)
DRESS_TTE <- prepare_top_drugs_TTE(final_table_TTE_DRESS, "DRESS", top_n = 40)
AGEP_TTE <- prepare_top_drugs_TTE(final_table_TTE_AGEP, "AGEP", top_n = 40)

# Combine datasets and enforce SCAR order
combined_top_drugs <- bind_rows(SJS_TEN_TTE, DRESS_TTE, AGEP_TTE) %>%
  mutate(SCAR = factor(SCAR, levels = c("SJS-TEN", "DRESS", "AGEP")))

# Function to calculate overall statistics
calculate_overall_stats <- function(data, scar_name) {
  data <- data %>% filter(!is.na(TTE))
  filtered_data <- data %>% filter(TTE > 0 & TTE <= 180)
  
  # IQR filtering
  Q1 <- quantile(filtered_data$TTE, 0.25)
  Q3 <- quantile(filtered_data$TTE, 0.75)
  IQR <- Q3 - Q1
  lower_bound <- Q1 - 1.5 * IQR
  upper_bound <- Q3 + 1.5 * IQR
  filtered_data <- filtered_data %>% filter(TTE >= lower_bound & TTE <= upper_bound)
  
  # Compute statistics
  data.frame(
    SCAR = scar_name,
    causative_drug = "All Drugs",
    median_TTE = median(filtered_data$TTE),
    lower_CI = quantile(filtered_data$TTE, 0.25),
    upper_CI = quantile(filtered_data$TTE, 0.75)
  )
}

# Compute overall stats for each SCAR type
overall_stats <- bind_rows(
  calculate_overall_stats(final_table_TTE_SJSTEN, "SJS-TEN"),
  calculate_overall_stats(final_table_TTE_DRESS, "DRESS"),
  calculate_overall_stats(final_table_TTE_AGEP, "AGEP")
) %>% mutate(SCAR = factor(SCAR, levels = c("SJS-TEN", "DRESS", "AGEP")))

# Calculate summary stats and append overall stats
summary_stats <- combined_top_drugs %>%
  group_by(SCAR, causative_drug) %>%
  summarise(
    median_TTE = median(TTE, na.rm = TRUE),
    lower_CI = quantile(TTE, probs = 0.25, na.rm = TRUE),
    upper_CI = quantile(TTE, probs = 0.75, na.rm = TRUE),
    .groups = "drop"  # Ensure it returns an ungrouped data frame
  )

# Append overall stats to summary stats and renumber rows
summary_stats <- bind_rows(summary_stats, overall_stats) %>%
  mutate(row_id = row_number())  # Renumber rows
```

Now for the plot

```{r}
# Plot: Double Lollipop with "All Drugs" as a diamond and black line
p4.2 <- ggplot(summary_stats, aes(x = reorder(causative_drug, as.numeric(causative_drug != "All Drugs")), 
                          y = median_TTE, color = SCAR)) +
  
      # Add vertical dashed line at "All Drugs" median for each SCAR
  geom_hline(data = subset(summary_stats, causative_drug == "All Drugs"), 
             aes(yintercept = median_TTE), 
             linetype = "dashed", color = "gray", size = 0.8) +  # Vertical dashed line
  
  # Error bars for CI
  geom_segment(aes(xend = causative_drug, y = lower_CI, yend = upper_CI), size = 1.2) + 
  geom_point(size = 3) +  # Median marker
  
  # Add "All Drugs" as a diamond and black line
  geom_segment(data = subset(summary_stats, causative_drug == "All Drugs"), 
               aes(xend = causative_drug, y = lower_CI, yend = upper_CI), 
               size = 1.2, color = "black") +  # Black CI line
  
  geom_point(data = subset(summary_stats, causative_drug == "All Drugs"),
             aes(x = causative_drug, y = median_TTE), shape = 18, size = 5, color = "black") + # Black diamond
  
  # Layout adjustments
  coord_flip() +
  facet_wrap(~SCAR, scales = "free") +
  theme_minimal() +
  labs(x = "Causative Drug",
    y = "Median Latency, 25-75th %ile (Days)"
  ) +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10),
    strip.text = element_text(size = 12, face = "bold"),
    legend.position = "none"
  ) +
  scale_color_manual(values = scar_colors) +
  scale_y_continuous(breaks = seq(0, 180, 10))

plot(p4.2)
```


Create the TTE plot:


```{r}
# Combine p4.1 and p4.2 side by side
pFig4 <- p4.1 + p4.2 + plot_layout(widths = c(1, 1.5))

# Display the combined plot
pFig4
```

Another (using the Kaplan-Meier plot instead of the Violin plot):

```{r}
p4.1altfull <- p4.1alt$plot / (p4.1alt$table + xlab("Days")) + plot_layout(heights = c(8,2))
pFig4_alt2 <- p4.1altfull|p4.2
plot(pFig4_alt2)
```

