---
title: "FAERS_Mortality_Analysis"
author: "Eric Mukherjee"
date: "2025-02-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PACKAGES

```{r message=FALSE, warning=FALSE, include=FALSE, r,results='hide'}
library(readr)
library(reshape2)
library(dplyr)
library(progress)
library(RColorBrewer)
library(patchwork)
library(ggpubr)
library(gridExtra)
library(grid)
library(pdp)
library(Hmisc)
library(ggplot2)
library(caret)
library(ranger)
library(caret)
library(doParallel)
library(scales)
library(data.table)
library(pROC)
library(gt)
library(scales)
library(vivid)
library(pdp)

```

##INTRODUCTION

This document is to analyze the determinants of mortality in SCAR.

We will use pre-sanitized data from the FAERS_Survival_Analysis sheet (clean_data_SCAR.csv). Note this contains only those rows with age between 0 and 120 (with NAs replaced by the median age), and removes rows with outlier TTE (either greater than 180 days or less than 1 day)

```{r}
# Import Data
clean_data_SCAR <- read_csv("clean_data_SCAR.csv")
```

###Get outcome categories

We also need the outcome category data from FAERS, which we haven't really used yet

```{r}
standard_case_outcome_category <- read_csv("standard_case_outcome_category.csv")
```

We'll need a function to combine isr and primaryid into one type of id - the compositeid.

```{r}
# Function to create compositeids based on primaryid and isr, with NA handling
create_compositeid <- function(data) {
    # Ensure data is a data.table
    data <- as.data.table(data)
    
    # Check if both primaryid and isr columns exist
    if ("primaryid" %in% names(data) & "isr" %in% names(data)) {
        # Create compositeid using primaryid if it's not NA, otherwise use isr
        data[, compositeid := ifelse(!is.na(primaryid), paste0("pid.", primaryid), paste0("isr.", isr))]
    
    # If only primaryid exists
    } else if ("primaryid" %in% names(data)) {
        data[, compositeid := paste0("pid.", primaryid)]
    
    # If only isr exists
    } else if ("isr" %in% names(data)) {
        data[, compositeid := paste0("isr.", isr)]
    
    # If neither column exists, throw an error
    } else {
        stop("Neither 'primaryid' nor 'isr' columns found in the data.")
    }
    
    # Return the updated data table
    return(data)
}
```

Then apply this function to our new table:

```{r}
# Add a compositeid and remove the other columns
standard_case_outcome_category <- create_compositeid(standard_case_outcome_category)
standard_case_outcome_category$primaryid <- NULL
standard_case_outcome_category$isr <- NULL
standard_case_outcome_category$snomed_concept_id <- NULL
```

Excellent. Now join this data to our extant data

```{r}
clean_data_SCAR_mortality <- merge(clean_data_SCAR, standard_case_outcome_category, by = "compositeid", all.x = TRUE)
```

Fix the new table by having a Death boolean:

```{r}
# Convert to data.table
clean_data_SCAR_mortality <- as.data.table(clean_data_SCAR_mortality)

# Create a binary "Death" column:
# - If any row for a given compositeid has "DE", set Death = 1
# - If no "DE" but other outcomes exist, set Death = 0
# - If all `outc_code` values are NA for a compositeid, set Death = NA
clean_data_SCAR_mortality[, Death := fifelse(
  any(outc_code == "DE", na.rm = TRUE), 1,   # If at least one "DE", set to 1
  fifelse(all(is.na(outc_code)), NA_integer_, 0)  # If all NA, keep NA; otherwise, 0
), by = compositeid]

# Remove duplicate rows while keeping only one row per compositeid
clean_data_SCAR_mortality <- unique(clean_data_SCAR_mortality, by = "compositeid")

# Drop the original outc_code column if no longer needed
clean_data_SCAR_mortality[, outc_code := NULL]

# Convert Death to a factor for classification
clean_data_SCAR_mortality[, Death := factor(Death, levels = c(0, 1), labels = c("No", "Yes"))]

# Print the cleaned dataset
print(clean_data_SCAR_mortality)
```

Now that data is clean! Let's run our analysis.

##ANALYSIS

###Descriptive Analysis

Let's look at mortality rate per SCAR:

```{r}
# Filter out rows where Death is NA
clean_data_filtered <- clean_data_SCAR_mortality[!is.na(Death)]

#Subset data
clean_data_filtered_SJSTEN<- clean_data_filtered[clean_data_filtered$SJSTEN,]
clean_data_filtered_DRESS<- clean_data_filtered[clean_data_filtered$DRESS,]
clean_data_filtered_AGEP<- clean_data_filtered[clean_data_filtered$AGEP,]
```

We use this filtered version of the data for all subsequent analyses.

```{r}
# Compute mortality rate for each SCAR type separately
mortality_by_SCAR <- data.table(
  SCAR = c("SJS-TEN", "DRESS", "AGEP", "GBFDE"),
  
  # Total patients per SCAR
  Total_Patients = c(
    sum(clean_data_filtered$SJSTEN, na.rm = TRUE),
    sum(clean_data_filtered$DRESS, na.rm = TRUE),
    sum(clean_data_filtered$AGEP, na.rm = TRUE),
    sum(clean_data_filtered$GBFDE, na.rm = TRUE)
  ),
  
  # Count deaths per SCAR (Death == "Yes")
  Deaths = c(
    sum(clean_data_filtered$Death == "Yes" & clean_data_filtered$SJSTEN == 1, na.rm = TRUE),
    sum(clean_data_filtered$Death == "Yes" & clean_data_filtered$DRESS == 1, na.rm = TRUE),
    sum(clean_data_filtered$Death == "Yes" & clean_data_filtered$AGEP == 1, na.rm = TRUE),
    sum(clean_data_filtered$Death == "Yes" & clean_data_filtered$GBFDE == 1, na.rm = TRUE)
  )
)

# Calculate mortality rate as percentage
mortality_by_SCAR[, Mortality_Rate := (Deaths / Total_Patients) * 100]

# Mortality by Age Group
mortality_by_age <- clean_data_filtered[, .(
  Total_Patients = .N,
  Deaths = sum(Death == "Yes"),
  Mortality_Rate = (sum(Death == "Yes") / .N) * 100
), by = age_group][order(-Mortality_Rate)]  # Sort by highest mortality rate

# Mortality by Sex
mortality_by_sex <- clean_data_filtered[, .(
  Total_Patients = .N,
  Deaths = sum(Death == "Yes"),
  Mortality_Rate = (sum(Death == "Yes") / .N) * 100
), by = sex][order(-Mortality_Rate)]  # Sort by highest mortality rate

# Mortality by Region
mortality_by_region <- clean_data_filtered[, .(
  Total_Patients = .N,
  Deaths = sum(Death == "Yes"),
  Mortality_Rate = (sum(Death == "Yes") / .N) * 100
), by = region][order(-Mortality_Rate)]  # Sort by highest mortality rate

# Print the final tables
print("Mortality by SCAR Type:")
print(mortality_by_SCAR)

print("Mortality by Age Group:")
print(mortality_by_age)

print("Mortality by Sex:")
print(mortality_by_sex)

print("Mortality by Region:")
print(mortality_by_region)
```

Let's plot mortality curves:

```{r}
library(Hmisc)  # For adaptive binning
library(ggplot2)
library(data.table)

# Ensure Death is binary (1 = Death, 0 = No Death)
clean_data_filtered[, Death := as.numeric(Death == "Yes")]

# 🔹 Adaptive Binning (Equal Frequency Bins)
clean_data_filtered[, age_bin := cut2(age_yr, g = 10)]  # 10 bins based on data distribution
clean_data_filtered[, TTE_bin := cut2(TTE, g = 10)]
clean_data_filtered[, concomitant_bin := cut2(concomitant_count, g = 10)]

# Compute mortality rates by **adaptive binned age**
mortality_age <- clean_data_filtered[, .(
  Total = .N,
  Deaths = sum(Death, na.rm = TRUE),
  Mortality_Rate = (sum(Death, na.rm = TRUE) / .N) * 100
), by = age_bin]

# Compute mortality rates by **adaptive binned concomitant drugs**
mortality_concomitant <- clean_data_filtered[, .(
  Total = .N,
  Deaths = sum(Death, na.rm = TRUE),
  Mortality_Rate = (sum(Death, na.rm = TRUE) / .N) * 100
), by = concomitant_bin]

# Compute mortality rates by **adaptive binned TTE**
mortality_tte <- clean_data_filtered[, .(
  Total = .N,
  Deaths = sum(Death, na.rm = TRUE),
  Mortality_Rate = (sum(Death, na.rm = TRUE) / .N) * 100
), by = TTE_bin]

# 🔹 Plot Adaptive Binned Mortality Rate Curves
p_actual_age <- ggplot(mortality_age, aes(x = age_bin, y = Mortality_Rate, group = 1)) +
  geom_line(color = "black", size = 1) +
  geom_point(color = "black", size = 2) +
  labs(title = "Real Mortality by Age",
       x = "Age (Years)",
       y = "Mortality Rate (%)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x labels

p_actual_concomitant <- ggplot(mortality_concomitant, aes(x = concomitant_bin, y = Mortality_Rate, group = 1)) +
  geom_line(color = "black", size = 1) +
  geom_point(color = "black", size = 2) +
  labs(title = "Real Mortality by Concomitant Drug Count",
       x = "Concomitant Drug Count",
       y = "Mortality Rate (%)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p_actual_tte <- ggplot(mortality_tte, aes(x = TTE_bin, y = Mortality_Rate, group = 1)) +
  geom_line(color = "black", size = 1) +
  geom_point(color = "black", size = 2) +
  labs(title = "Real Mortality by TTE",
       x = "Time to Event (TTE, Days)",
       y = "Mortality Rate (%)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Now bar graphs for our categorical friends

```{r}
# Compute actual mortality by region
mortality_region <- clean_data_filtered[, .(
  Total = .N,
  Deaths = sum(Death, na.rm = TRUE),
  Mortality_Rate = (sum(Death, na.rm = TRUE) / .N) * 100
), by = region]

# Compute actual mortality by sex
mortality_sex <- clean_data_filtered[, .(
  Total = .N,
  Deaths = sum(Death, na.rm = TRUE),
  Mortality_Rate = (sum(Death, na.rm = TRUE) / .N) * 100
), by = sex]

# Compute actual mortality by SCAR type
mortality_scar <- data.table(
  SCAR = c("SJS-TEN", "DRESS", "AGEP", "GBFDE"),
  Total_Patients = c(
    sum(clean_data_filtered$SJSTEN, na.rm = TRUE),
    sum(clean_data_filtered$DRESS, na.rm = TRUE),
    sum(clean_data_filtered$AGEP, na.rm = TRUE),
    sum(clean_data_filtered$GBFDE, na.rm = TRUE)
  ),
  Deaths = c(
    sum(clean_data_filtered$Death[clean_data_filtered$SJSTEN == 1], na.rm = TRUE),
    sum(clean_data_filtered$Death[clean_data_filtered$DRESS == 1], na.rm = TRUE),
    sum(clean_data_filtered$Death[clean_data_filtered$AGEP == 1], na.rm = TRUE),
    sum(clean_data_filtered$Death[clean_data_filtered$GBFDE == 1], na.rm = TRUE)
  )
)

# Calculate actual mortality rate for SCAR
mortality_scar[, Mortality_Rate := (Deaths / Total_Patients) * 100]

# Define SCAR colors
scar_colors <- setNames(brewer.pal(5, "Set1"),  
                        c("SJS-TEN", "DRESS", "AGEP", "GBFDE", "Total SCAR"))

# Bar plots for actual mortality rates
p_actual_region <- ggplot(mortality_region, aes(x = region, y = Mortality_Rate, fill = region)) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Real Mortality by Region",
       x = "Region",
       y = "Mortality Rate (%)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + theme(legend.position = "none")

p_actual_sex <- ggplot(mortality_sex, aes(x = sex, y = Mortality_Rate, fill = sex)) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_manual(values = c("F" = "pink", "M" = "lightblue", "NS" = "lightgreen")) +
  labs(title = "Real Mortality by Sex",
       x = "Sex",
       y = "Mortality Rate (%)") +
  theme_minimal() + theme(legend.position = "none")

p_actual_scar <- ggplot(mortality_scar, aes(x = SCAR, y = Mortality_Rate, fill = SCAR)) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_manual(values = scar_colors) +
  labs(title = "Real Mortality by SCAR Type",
       x = "SCAR Type",
       y = "Mortality Rate (%)") +
  theme_minimal() + theme(legend.position = "none")
```

Final plot together:

```{r}
# Combine the six PDPs into one vertical column
actual_mortality_left <- p_actual_age / p_actual_concomitant / p_actual_tte + 
  plot_layout(ncol = 1)  # Stack vertically

actual_mortality_right <- p_actual_sex / p_actual_region / p_actual_scar + 
  plot_layout(ncol = 1)

actual_mortality_final <- actual_mortality_left | actual_mortality_right
```
Lets do per-SCAR mortality as well! First SJS-TEN:

```{r}
# Have to change this as usual
clean_data_filtered_SJSTEN[, Death := as.numeric(Death == "Yes")]
```

All graphs in one

```{r}
# 🔹 Adaptive Binning (Equal Frequency Bins)
clean_data_filtered_SJSTEN[, age_bin := cut2(age_yr, g = 10)]  # 10 bins based on data distribution
clean_data_filtered_SJSTEN[, TTE_bin := cut2(TTE, g = 10)]
clean_data_filtered_SJSTEN[, concomitant_bin := cut2(concomitant_count, g = 10)]

# Compute mortality rates by **adaptive binned age**
mortality_age_sjsten <- clean_data_filtered_SJSTEN[, .(
  Total = .N,
  Deaths = sum(Death, na.rm = TRUE),
  Mortality_Rate = (sum(Death, na.rm = TRUE) / .N) * 100
), by = age_bin]

# Compute mortality rates by **adaptive binned concomitant drugs**
mortality_concomitant_sjsten <- clean_data_filtered_SJSTEN[, .(
  Total = .N,
  Deaths = sum(Death, na.rm = TRUE),
  Mortality_Rate = (sum(Death, na.rm = TRUE) / .N) * 100
), by = concomitant_bin]

# Compute mortality rates by **adaptive binned TTE**
mortality_tte_sjsten <- clean_data_filtered_SJSTEN[, .(
  Total = .N,
  Deaths = sum(Death, na.rm = TRUE),
  Mortality_Rate = (sum(Death, na.rm = TRUE) / .N) * 100
), by = TTE_bin]

# 🔹 Plot Adaptive Binned Mortality Rate Curves
p_actual_age_sjsten <- ggplot(mortality_age_sjsten, aes(x = age_bin, y = Mortality_Rate, group = 1)) +
  geom_line(color = "black", size = 1) +
  geom_point(color = "black", size = 2) +
  labs(title = "Mortality by Age (SJS-TEN)",
       x = "Age (Years)",
       y = "Mortality Rate (%)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x labels

p_actual_concomitant_sjsten <- ggplot(mortality_concomitant_sjsten, aes(x = concomitant_bin, y = Mortality_Rate, group = 1)) +
  geom_line(color = "black", size = 1) +
  geom_point(color = "black", size = 2) +
  labs(title = "Mortality by #Concom. Drug (SJS-TEN)",
       x = "Concomitant Drug Count",
       y = "Mortality Rate (%)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p_actual_tte_sjsten <- ggplot(mortality_tte_sjsten, aes(x = TTE_bin, y = Mortality_Rate, group = 1)) +
  geom_line(color = "black", size = 1) +
  geom_point(color = "black", size = 2) +
  labs(title = "Mortality by TTE (SJS-TEN)",
       x = "Time to Event (TTE, Days)",
       y = "Mortality Rate (%)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Compute actual mortality by region
mortality_region_sjsten <- clean_data_filtered_SJSTEN[, .(
  Total = .N,
  Deaths = sum(Death, na.rm = TRUE),
  Mortality_Rate = (sum(Death, na.rm = TRUE) / .N) * 100
), by = region]

# Compute actual mortality by sex
mortality_sex_sjsten <- clean_data_filtered_SJSTEN[, .(
  Total = .N,
  Deaths = sum(Death, na.rm = TRUE),
  Mortality_Rate = (sum(Death, na.rm = TRUE) / .N) * 100
), by = sex]

# Bar plots for actual mortality rates
p_actual_region_sjsten <- ggplot(mortality_region_sjsten, aes(x = region, y = Mortality_Rate, fill = region)) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Mortality by Region (SJS-TEN)",
       x = "Region",
       y = "Mortality Rate (%)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + theme(legend.position = "none")

p_actual_sex_sjsten <- ggplot(mortality_sex_sjsten, aes(x = sex, y = Mortality_Rate, fill = sex)) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_manual(values = c("F" = "pink", "M" = "lightblue", "NS" = "lightgreen")) +
  labs(title = "Mortality by Sex (SJS-TEN)",
       x = "Sex",
       y = "Mortality Rate (%)") +
  theme_minimal() + theme(legend.position = "none")
```

Now put all the plots together:

```{r}
actual_mortality_left_sjsten <- p_actual_age_sjsten / p_actual_concomitant_sjsten / p_actual_tte_sjsten + 
  plot_layout(ncol = 1)  # Stack vertically

actual_mortality_right_sjsten <- p_actual_region_sjsten/ p_actual_sex_sjsten  + 
  plot_layout(ncol = 1)

actual_mortality_final_sjsten <- actual_mortality_left_sjsten | actual_mortality_right_sjsten
```
Next up, DRESS:

```{r}
# Have to change this as usual
clean_data_filtered_DRESS[, Death := as.numeric(Death == "Yes")]
```

All graphs in one

```{r}
# 🔹 Adaptive Binning (Equal Frequency Bins)
clean_data_filtered_DRESS[, age_bin := cut2(age_yr, g = 10)]  # 10 bins based on data distribution
clean_data_filtered_DRESS[, TTE_bin := cut2(TTE, g = 10)]
clean_data_filtered_DRESS[, concomitant_bin := cut2(concomitant_count, g = 10)]

# Compute mortality rates by **adaptive binned age**
mortality_age_dress <- clean_data_filtered_DRESS[, .(
  Total = .N,
  Deaths = sum(Death, na.rm = TRUE),
  Mortality_Rate = (sum(Death, na.rm = TRUE) / .N) * 100
), by = age_bin]

# Compute mortality rates by **adaptive binned concomitant drugs**
mortality_concomitant_dress <- clean_data_filtered_DRESS[, .(
  Total = .N,
  Deaths = sum(Death, na.rm = TRUE),
  Mortality_Rate = (sum(Death, na.rm = TRUE) / .N) * 100
), by = concomitant_bin]

# Compute mortality rates by **adaptive binned TTE**
mortality_tte_dress <- clean_data_filtered_DRESS[, .(
  Total = .N,
  Deaths = sum(Death, na.rm = TRUE),
  Mortality_Rate = (sum(Death, na.rm = TRUE) / .N) * 100
), by = TTE_bin]

# 🔹 Plot Adaptive Binned Mortality Rate Curves
p_actual_age_dress <- ggplot(mortality_age_dress, aes(x = age_bin, y = Mortality_Rate, group = 1)) +
  geom_line(color = "black", size = 1) +
  geom_point(color = "black", size = 2) +
  labs(title = "Mortality by Age (DRESS)",
       x = "Age (Years)",
       y = "Mortality Rate (%)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x labels

p_actual_concomitant_dress <- ggplot(mortality_concomitant_dress, aes(x = concomitant_bin, y = Mortality_Rate, group = 1)) +
  geom_line(color = "black", size = 1) +
  geom_point(color = "black", size = 2) +
  labs(title = "Mortality by #Concom. Drug (DRESS)",
       x = "Concomitant Drug Count",
       y = "Mortality Rate (%)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p_actual_tte_dress <- ggplot(mortality_tte_dress, aes(x = TTE_bin, y = Mortality_Rate, group = 1)) +
  geom_line(color = "black", size = 1) +
  geom_point(color = "black", size = 2) +
  labs(title = "Mortality by TTE (DRESS)",
       x = "Time to Event (TTE, Days)",
       y = "Mortality Rate (%)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Compute actual mortality by region
mortality_region_dress <- clean_data_filtered_DRESS[, .(
  Total = .N,
  Deaths = sum(Death, na.rm = TRUE),
  Mortality_Rate = (sum(Death, na.rm = TRUE) / .N) * 100
), by = region]

# Compute actual mortality by sex
mortality_sex_dress <- clean_data_filtered_DRESS[, .(
  Total = .N,
  Deaths = sum(Death, na.rm = TRUE),
  Mortality_Rate = (sum(Death, na.rm = TRUE) / .N) * 100
), by = sex]

# Bar plots for actual mortality rates
p_actual_region_dress <- ggplot(mortality_region_dress, aes(x = region, y = Mortality_Rate, fill = region)) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Mortality by Region (DRESS)",
       x = "Region",
       y = "Mortality Rate (%)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + theme(legend.position = "none")

p_actual_sex_dress <- ggplot(mortality_sex_dress, aes(x = sex, y = Mortality_Rate, fill = sex)) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_manual(values = c("F" = "pink", "M" = "lightblue", "NS" = "lightgreen")) +
  labs(title = "Mortality by Sex (DRESS)",
       x = "Sex",
       y = "Mortality Rate (%)") +
  theme_minimal() + theme(legend.position = "none")
```

Now put all the plots together:

```{r}
actual_mortality_left_dress <- p_actual_age_dress / p_actual_concomitant_dress / p_actual_tte_dress + 
  plot_layout(ncol = 1)  # Stack vertically

actual_mortality_right_dress <- p_actual_region_dress/ p_actual_sex_dress  + 
  plot_layout(ncol = 1)

actual_mortality_final_dress <- actual_mortality_left_dress | actual_mortality_right_dress
```
Finally, AGEP:

```{r}
# Have to change this as usual
clean_data_filtered_AGEP[, Death := as.numeric(Death == "Yes")]
```

All graphs in one

```{r}
# 🔹 Adaptive Binning (Equal Frequency Bins)
clean_data_filtered_AGEP[, age_bin := cut2(age_yr, g = 10)]  # 10 bins based on data distribution
clean_data_filtered_AGEP[, TTE_bin := cut2(TTE, g = 10)]
clean_data_filtered_AGEP[, concomitant_bin := cut2(concomitant_count, g = 10)]

# Compute mortality rates by **adaptive binned age**
mortality_age_agep <- clean_data_filtered_AGEP[, .(
  Total = .N,
  Deaths = sum(Death, na.rm = TRUE),
  Mortality_Rate = (sum(Death, na.rm = TRUE) / .N) * 100
), by = age_bin]

# Compute mortality rates by **adaptive binned concomitant drugs**
mortality_concomitant_agep <- clean_data_filtered_AGEP[, .(
  Total = .N,
  Deaths = sum(Death, na.rm = TRUE),
  Mortality_Rate = (sum(Death, na.rm = TRUE) / .N) * 100
), by = concomitant_bin]

# Compute mortality rates by **adaptive binned TTE**
mortality_tte_agep <- clean_data_filtered_AGEP[, .(
  Total = .N,
  Deaths = sum(Death, na.rm = TRUE),
  Mortality_Rate = (sum(Death, na.rm = TRUE) / .N) * 100
), by = TTE_bin]

# 🔹 Plot Adaptive Binned Mortality Rate Curves
p_actual_age_agep <- ggplot(mortality_age_agep, aes(x = age_bin, y = Mortality_Rate, group = 1)) +
  geom_line(color = "black", size = 1) +
  geom_point(color = "black", size = 2) +
  labs(title = "Mortality by Age (AGEP)",
       x = "Age (Years)",
       y = "Mortality Rate (%)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x labels

p_actual_concomitant_agep <- ggplot(mortality_concomitant_agep, aes(x = concomitant_bin, y = Mortality_Rate, group = 1)) +
  geom_line(color = "black", size = 1) +
  geom_point(color = "black", size = 2) +
  labs(title = "Mortality by #Concom. Drug (AGEP)",
       x = "Concomitant Drug Count",
       y = "Mortality Rate (%)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p_actual_tte_agep <- ggplot(mortality_tte_agep, aes(x = TTE_bin, y = Mortality_Rate, group = 1)) +
  geom_line(color = "black", size = 1) +
  geom_point(color = "black", size = 2) +
  labs(title = "Mortality by TTE (AGEP)",
       x = "Time to Event (TTE, Days)",
       y = "Mortality Rate (%)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Compute actual mortality by region
mortality_region_agep <- clean_data_filtered_AGEP[, .(
  Total = .N,
  Deaths = sum(Death, na.rm = TRUE),
  Mortality_Rate = (sum(Death, na.rm = TRUE) / .N) * 100
), by = region]

# Compute actual mortality by sex
mortality_sex_agep <- clean_data_filtered_AGEP[, .(
  Total = .N,
  Deaths = sum(Death, na.rm = TRUE),
  Mortality_Rate = (sum(Death, na.rm = TRUE) / .N) * 100
), by = sex]

# Bar plots for actual mortality rates
p_actual_region_agep <- ggplot(mortality_region_agep, aes(x = region, y = Mortality_Rate, fill = region)) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Mortality by Region (AGEP)",
       x = "Region",
       y = "Mortality Rate (%)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + theme(legend.position = "none")

p_actual_sex_agep <- ggplot(mortality_sex_agep, aes(x = sex, y = Mortality_Rate, fill = sex)) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_manual(values = c("F" = "pink", "M" = "lightblue", "NS" = "lightgreen")) +
  labs(title = "Mortality by Sex (AGEP)",
       x = "Sex",
       y = "Mortality Rate (%)") +
  theme_minimal() + theme(legend.position = "none")
```

Now put all the plots together:

```{r}
actual_mortality_left_agep <- p_actual_age_agep / p_actual_concomitant_agep / p_actual_tte_agep + 
  plot_layout(ncol = 1)  # Stack vertically

actual_mortality_right_agep <- p_actual_region_agep / p_actual_sex_agep  + 
  plot_layout(ncol = 1)

actual_mortality_final_agep <- actual_mortality_left_agep | actual_mortality_right_agep
```


###Univariate analysis

Let's test the categorical predictors first, for Chi-square:

```{r}
# Chi-square test for categorical variables vs. mortality
chisq_sex <- chisq.test(table(clean_data_filtered$sex, clean_data_filtered$Death))
chisq_age_group <- chisq.test(table(clean_data_filtered$age_group, clean_data_filtered$Death))
chisq_region <- chisq.test(table(clean_data_filtered$region, clean_data_filtered$Death))
chisq_causative_drug <- chisq.test(table(clean_data_filtered$causative_drug, clean_data_filtered$Death))

# Print p-values
list(
  Sex_p_value = chisq_sex$p.value,
  Age_Group_p_value = chisq_age_group$p.value,
  Region_p_value = chisq_region$p.value,
  Causative_Drug_p_value = chisq_causative_drug$p.value
)
```


Then, use the wilcoxon test for continuous predictors:

```{r}
# Mann-Whitney U test for TTE (since TTE is often skewed)
wilcox_tte <- wilcox.test(TTE ~ Death, data = clean_data_filtered)

# Mann-Whitney U test for concomitant count
wilcox_concomitant <- wilcox.test(concomitant_count ~ Death, data = clean_data_filtered)

# Mann-Whitney U test for age
wilcox_age <- wilcox.test(age_yr ~ Death, data = clean_data_filtered)

# Print results
list(
  TTE_p_value = wilcox_tte$p.value,
  Concomitant_Count_p_value = wilcox_concomitant$p.value,
  age_p_value = wilcox_age$p.value
)
```

#RANDOM FORESTS

we're going to regenerate this variable, because it has certain requirements (it previously had death made numeric, but we need it as a factor):

```{r}
# Filter out rows where Death is NA
clean_data_filtered <- clean_data_SCAR_mortality[!is.na(Death)]

#Subset data
clean_data_filtered_SJSTEN<- clean_data_filtered[clean_data_filtered$SJSTEN,]
clean_data_filtered_DRESS<- clean_data_filtered[clean_data_filtered$DRESS,]
clean_data_filtered_AGEP<- clean_data_filtered[clean_data_filtered$AGEP,]
```

##Interaction-Enhanced RF

```{r}
# Identify the top 50 most common drugs
top_drugs <- names(sort(table(clean_data_filtered$causative_drug), decreasing = TRUE)[1:50])

# Recode rare drugs as "Other"
clean_data_filtered[, causative_drug := ifelse(causative_drug %in% top_drugs, causative_drug, "Other")]

# Convert to factor
clean_data_filtered[, causative_drug := as.factor(causative_drug)]

# Define SCAR group for grouped importance
grouped_vars <- list(
  "SCAR_Combined" = c("SJSTEN", "DRESS", "AGEP", "GBFDE")
)

# Define hyperparameter grid for `ranger`
tune_grid <- expand.grid(
  mtry = c(2, round(sqrt(ncol(clean_data_filtered) - 1)), round(log2(ncol(clean_data_filtered) - 1))),
  min.node.size = c(1, 5, 10),  # Controls tree depth
  splitrule = c("gini", "extratrees")  # Try both Gini and ExtraTrees
)

# Enable parallel processing
num_cores <- detectCores() - 1  # Use all but one core
cl <- makeCluster(num_cores)
registerDoParallel(cl)

# Define cross-validation strategy
control <- trainControl(
  method = "cv", 
  number = 5,
  classProbs = TRUE,
  verboseIter = TRUE  # Enables progress messages
)

# Run grid search with parallel processing
rf_tuned_interaction <- train(
  Death ~ age_yr + sex + region + causative_drug + TTE + concomitant_count + 
          SJSTEN + DRESS + AGEP + GBFDE, 
  data = clean_data_filtered, 
  method = "ranger",
  trControl = control,
  tuneGrid = tune_grid,
  num.trees = 500,  # Number of trees
  importance = "permutation",  # Enables variable importance computation
  seed = 42
)

# Stop parallel processing
stopCluster(cl)

# View best hyperparameters
print(rf_tuned_interaction$bestTune)

# Train final optimized model
rf_final_interaction <- ranger(
  formula = Death ~ age_yr + sex + region + causative_drug + TTE + concomitant_count + 
            SJSTEN + DRESS + AGEP + GBFDE,
  data = clean_data_filtered,
  num.trees = 500,
  mtry = rf_tuned_interaction$bestTune$mtry,
  min.node.size = rf_tuned_interaction$bestTune$min.node.size,
  splitrule = rf_tuned_interaction$bestTune$splitrule,
  probability = TRUE,  # Enables probability predictions
  importance = "permutation",
  seed = 42
)

# Extract variable importance scores
vimp_results <- as.data.table(rf_final_interaction$variable.importance, keep.rownames = TRUE)
setnames(vimp_results, c("Variable", "Importance"))

# Compute Joint VIMP for SCAR Types (Summing Individual Contributions)
scar_vimp <- vimp_results[Variable %in% c("SJSTEN", "DRESS", "AGEP", "GBFDE"), .(
  Variable = "SCAR_Combined",
  Importance = sum(Importance)  # Summing importance scores
)]

# Append Joint SCAR Importance to VIMP Results
vimp_results <- rbind(vimp_results, scar_vimp)

# Print variable importance table (including joint SCAR VIMP)
print(vimp_results)

# Plot VIMP (Including Grouped SCAR Importance)
ggplot(vimp_results, aes(x = reorder(Variable, Importance), y = Importance)) +
  geom_bar(stat = "identity", fill = "steelblue", color = "black") +
  coord_flip() +
  labs(title = "Variable Importance (VIMP) in Mortality Prediction",
       x = "Variable",
       y = "VIMP Score") +
  theme_minimal()
```

Excellent, it works!

##VALIDATION

We'll have to validate this data in particular ways. First, calculate the OOB error rate:

```{r}
# OOB Error Rate
oob_error <- rf_final_interaction$prediction.error  # Extract OOB error
print(paste("OOB Error Rate:", oob_error))
```
Slightly better than the other RF.

```{r}
# Get OOB Predicted Probabilities & True Labels
oob_predictions <- rf_final_interaction$predictions  # Extract predicted probabilities

# Ensure Death labels are a binary factor with correct levels
true_labels <- factor(clean_data_filtered$Death, levels = c("No", "Yes"))  

# Compute AUC using OOB probabilities
roc_curve <- roc(response = true_labels, predictor = oob_predictions[, "Yes"], levels = c("No", "Yes"))

# Extract AUC value
auc_value <- auc(roc_curve)

# Print the corrected AUC value
print(paste("Corrected AUC:", round(auc_value, 3)))

# Optional: Plot the ROC Curve
plot(roc_curve, col = "blue", main = "ROC Curve for Random Forest Model")
```
Interesting AUC. We will select this model for future work.

Lets put these metrics in a table:

```{r}
# Define the metrics
metrics <- data.table(
  Metric = c("OOB Error/Brier Score", "AUC"),
  Value = c(round(oob_error, 4), round(auc_value, 4))
)

# Convert to a gt table for better visualization
metric_table <- gt(metrics) %>%
  tab_header(title = "Model Performance Metrics") %>%
  cols_label(
    Metric = "Metric",
    Value = "Value"
  ) %>%
  fmt_number(columns = Value, decimals = 4) %>%
  tab_options(
    table.font.size = px(14),
    heading.title.font.size = px(16),
    heading.title.font.weight = "bold"
  )

# Print the table
metric_table
```

###Reliability plots

```{r}
# Extract OOB predicted probabilities (probability of Death = "Yes")
clean_data_filtered[, Pred_Death_Prob := rf_final_interaction$predictions[, "Yes"]]

# Check summary of predictions
summary(clean_data_filtered$Pred_Death_Prob)
```
```{r}
hist(clean_data_filtered$Pred_Death_Prob, breaks = 20, 
     main = "Histogram of Predicted Mortality Probabilities", 
     col = "skyblue", xlab = "Predicted Probability")
```



Now create the reliability plot

```{r}
# Step 1: Create bins for predicted probabilities (10 deciles)
clean_data_filtered[, Pred_Bin := cut2(Pred_Death_Prob, g = 10)]  # 10 bins with roughly equal counts

# Step 2: Compute mean predicted probability & actual mortality rate per bin
calibration_data <- clean_data_filtered[, .(
  Mean_Pred_Prob = mean(Pred_Death_Prob, na.rm = TRUE),
  Actual_Death_Rate = mean(Death == "Yes", na.rm = TRUE)
), by = Pred_Bin]

# Step 3: Sort data for smooth curve plotting
calibration_data <- calibration_data[order(Mean_Pred_Prob)]

# Step 4: Plot calibration curve
ggplot(calibration_data, aes(x = Mean_Pred_Prob, y = Actual_Death_Rate)) +
  geom_point(size = 3, color = "gold1") +  # Scatter points
  geom_line(color = "gold1", size = 1) +  # Connect points to form a curve
  geom_smooth(method = "loess", se = FALSE, color = "gold1", linetype = "dashed") +  # Smoothed trend
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "black") +  # Perfect calibration line
  scale_x_continuous(labels = percent, limits = c(0, 1)) +
  scale_y_continuous(labels = percent, limits = c(0, 1)) +
  labs(title = "Calibration Plot (Reliability Curve)",
       x = "Predicted Mortality Probability",
       y = "Observed Mortality Rate") +
  theme_minimal()
```
Now we can compute a Brier score:

```{r}
# Compute the Brier Score
brier_score <- mean((clean_data_filtered$Pred_Death_Prob - as.numeric(clean_data_filtered$Death == "Yes"))^2, na.rm = TRUE)

# Print result
print(paste("Brier Score:", round(brier_score, 4)))
```
###Calibration by Individual SCAR

We'll compute an AUC, Brier Score, and Calibration Plots for each SCAR. First, Brier score:

```{r}
# Function to compute Brier Score for each SCAR type using its boolean column
compute_brier_score <- function(scar_column) {
  scar_subset <- clean_data_filtered[get(scar_column) == 1]  # Subset by SCAR type
  mean((scar_subset$Pred_Death_Prob - as.numeric(scar_subset$Death == "Yes"))^2, na.rm = TRUE)
}

# Compute Brier Scores for each SCAR
brier_scores <- data.table(
  SCAR = c("SJS-TEN", "DRESS", "AGEP", "GBFDE"),
  Brier_Score = c(
    compute_brier_score("SJSTEN"),
    compute_brier_score("DRESS"),
    compute_brier_score("AGEP"),
    compute_brier_score("GBFDE")
  )
)

# Print Brier Scores
print(brier_scores)
```



Not bad, though for SJS-TEN it's kind of high.

```{r}
# Your predefined SCAR colors
scar_colors <- setNames(brewer.pal(5, "Set1"),  
                        c("SJS-TEN", "DRESS", "AGEP", "GBFDE", "Total SCAR"))

# Function to compute calibration data for a given SCAR type using boolean columns
compute_calibration_data <- function(scar_column, scar_label) {
  scar_subset <- clean_data_filtered[get(scar_column) == 1]  # Subset by SCAR type
  
  # Exclude GBFDE and cases with fewer than 10 patients
  if (scar_label == "GBFDE" || nrow(scar_subset) < 10) return(NULL)

  # Add small jitter to Pred_Death_Prob to prevent duplicate quantiles
  scar_subset[, Pred_Death_Prob := Pred_Death_Prob + runif(.N, -1e-6, 1e-6)]
  
  # Check number of unique values
  if (length(unique(scar_subset$Pred_Death_Prob)) < 10) {
    # Use alternative binning if too few unique values
    scar_subset[, Pred_Bin := cut2(Pred_Death_Prob, g = 10)]
  } else {
    # Standard binning (10 deciles)
    scar_subset[, Pred_Bin := cut(
      Pred_Death_Prob, 
      breaks = unique(quantile(Pred_Death_Prob, probs = seq(0, 1, 0.1), na.rm = TRUE)),
      include.lowest = TRUE, labels = FALSE
    )]
  }

  # Compute mean predicted probability & actual mortality rate per bin
  calibration_data <- scar_subset[, .(
    Mean_Pred_Prob = mean(Pred_Death_Prob, na.rm = TRUE),
    Actual_Death_Rate = mean(Death == "Yes", na.rm = TRUE)
  ), by = Pred_Bin]

  # Sort for smooth curve plotting
  calibration_data <- calibration_data[order(Mean_Pred_Prob)]
  calibration_data[, SCAR := scar_label]  # Add SCAR label
  
  return(calibration_data)
}

# Compute calibration data for each SCAR type (excluding GBFDE)
calibration_data_list <- rbindlist(list(
  compute_calibration_data("SJSTEN", "SJS-TEN"),
  compute_calibration_data("DRESS", "DRESS"),
  compute_calibration_data("AGEP", "AGEP")
), use.names = TRUE, fill = TRUE)

# Plot Calibration Curves using Your SCAR Colors (Excluding GBFDE)
ggplot(calibration_data_list, aes(x = Mean_Pred_Prob, y = Actual_Death_Rate, color = SCAR)) +
  geom_point(size = 3) +  # Scatter points
  geom_line(size = 1) +  # Connect points to form a curve
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "black") +  # Perfect calibration
  scale_x_continuous(labels = percent, limits = c(0, 1)) +
  scale_y_continuous(labels = percent, limits = c(0, 1)) +
  labs(title = "Calibration Curves for SCAR Types (Excluding GBFDE)",
       x = "Predicted Mortality Probability",
       y = "Observed Mortality Rate") +
  scale_color_manual(values = scar_colors[c("SJS-TEN", "DRESS", "AGEP")]) +  # Use only relevant colors
  theme_minimal()
```
Next, we can analyze AUC on a per-SCAR basis

```{r}
# Function to compute AUC for a given SCAR type
compute_auc <- function(scar_column, scar_label) {
  scar_subset <- clean_data_filtered[get(scar_column) == 1]  # Subset data for this SCAR type
  
  # Ensure there are enough data points
  if (nrow(scar_subset) < 10) return(NULL)
  
  # Convert Death to a numeric binary (1 = Yes, 0 = No)
  true_labels <- as.numeric(scar_subset$Death == "Yes")
  
  # Compute AUC using predicted probabilities
  roc_curve <- roc(response = true_labels, predictor = scar_subset$Pred_Death_Prob)
  auc_value <- auc(roc_curve)
  
  return(data.table(SCAR = scar_label, AUC = auc_value))
}

# Compute AUC for each SCAR type (excluding GBFDE due to low sample size)
auc_results <- rbindlist(list(
  compute_auc("SJSTEN", "SJS-TEN"),
  compute_auc("DRESS", "DRESS"),
  compute_auc("AGEP", "AGEP")
), use.names = TRUE, fill = TRUE)

# Print AUC values for each SCAR type
print(auc_results)

# Plot AUC bar chart
ggplot(auc_results, aes(x = SCAR, y = AUC, fill = SCAR)) +
  geom_bar(stat = "identity", color = "black") + 
  scale_fill_manual(values = scar_colors[c("SJS-TEN", "DRESS", "AGEP")]) +  # Use SCAR colors
  labs(title = "AUC Comparison Across SCAR Types",
       x = "SCAR Type",
       y = "AUC Score") +
  theme_minimal()
```

##VARAIBLE IMPORTANCE

###Variable Importance (by Bootstrap)

```{r}
# Define number of bootstrap iterations
num_bootstraps <- 100

# Store VIMP results from bootstraps
vimp_list <- vector("list", num_bootstraps)

# Bootstrapping loop
for (i in 1:num_bootstraps) {
  cat("Bootstrap iteration:", i, "\n")
  
  # Sample data with replacement
  bootstrap_sample <- clean_data_filtered[sample(.N, replace = TRUE)]
  
  # Train a new random forest model
  rf_bootstrap <- ranger(
    Death ~ age_yr + sex + region + causative_drug + TTE + concomitant_count + 
            SJSTEN + DRESS + AGEP + GBFDE, 
    data = bootstrap_sample, 
    num.trees = 500,
    importance = "permutation",
    probability = TRUE
  )
  
  # Store variable importance scores
  vimp_list[[i]] <- as.data.table(rf_bootstrap$variable.importance, keep.rownames = TRUE)
}
```

Now then, the boostrap is complete. Plot the results:

```{r}
# Combine all bootstrap VIMP results
vimp_results <- rbindlist(vimp_list, use.names = TRUE, fill = TRUE)
setnames(vimp_results, c("Variable", "Importance"))

# Compute summary statistics for each variable
vimp_summary <- vimp_results[, .(
  Mean_VIMP = mean(Importance),
  SD_VIMP = sd(Importance),
  Lower_CI = quantile(Importance, 0.025),
  Upper_CI = quantile(Importance, 0.975)
), by = Variable]


# Print bootstrapped VIMP summary with confidence intervals
print(vimp_summary)

# **Plot Bootstrapped VIMP (Including Individual SCAR VIMP, Removing Joint SCAR VIMP)**
pVimpBootstrapIndividual <- ggplot(vimp_summary, aes(x = reorder(Variable, Mean_VIMP), y = Mean_VIMP)) +
  
  # **Boxplot for all variables (INCLUDING SCAR subtypes)**
  geom_boxplot(data = vimp_results, 
               aes(x = reorder(Variable, Importance), y = Importance),
               fill = "lightblue", color = "black", outlier.shape = NA) +
  
  # **Jittered points for individual VIMP values**
  geom_jitter(data = vimp_results, 
              aes(x = Variable, y = Importance), 
              width = 0.2, alpha = 0.3, color = "darkblue") +
  
  # **Highlight individual SCAR VIMP values in red**
  geom_boxplot(data = vimp_results[Variable %in% c("SJSTEN", "DRESS", "AGEP", "GBFDE")], 
               aes(x = Variable, y = Importance), 
               fill = "red", color = "black", outlier.shape = NA) +
  
  geom_jitter(data = vimp_results[Variable %in% c("SJSTEN", "DRESS", "AGEP", "GBFDE")], 
              aes(x = Variable, y = Importance), 
              width = 0.2, alpha = 0.3, color = "darkred") +  
  
  # **Flip coordinates for readability**
  coord_flip() +
  
  # **Titles and Labels**
  labs(title = "Bootstrapped Variable Importance (VIMP)",
       x = "Variable",
       y = "Importance Score") +
  
  # **Minimalist Theme**
  theme_minimal()
```


###VIVI Plots

```{r}
# Define feature names (excluding the response variable "Death")
feature_names <- c("age_yr", "sex", "region", "causative_drug", 
                   "TTE", "concomitant_count", "SJSTEN", "DRESS", "AGEP", "GBFDE")

# Ensure Death is a factor
clean_data_filtered[, Death := as.factor(Death)]

# 🔹 Compute VIVI matrix
vivi_results <- vivi(
  data = clean_data_filtered[, c("Death", feature_names), with = FALSE],  # Data with response + predictors
  fit = rf_final_interaction,  # Trained ranger random forest model
  response = "Death",  # Response variable
  gridSize = 50,  # Grid size for evaluating predictions
  importanceType = "permutation",
  nmax = NULL,  # Use all data rows
  reorder = TRUE,  # Reorder matrix for better visualization
  class = "Yes",  # Factor level of the response variable to analyze
  numPerm = 4,  # Number of permutations for importance
  showVimpError = FALSE  # Suppress standard error printout
)

# 🔹 Convert `vivi_results` matrix to a long format data table
vivi_matrix <- as.data.table(melt(vivi_results, varnames = c("Var1", "Var2"), value.name = "Interaction"))
vivi_matrix[, Interaction := round(Interaction, 3)]  # Round values for display

# 🔹 Generate the VIVI Heatmap with Rotated Labels
vivi_plot <- viviHeatmap(vivi_results) +
  ggtitle("Variable Importance & Interaction Strengths") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 0))  # Rotate x-axis labels

# 🔹 Overlay text labels onto heatmap
pVIVI <- vivi_plot + 
  geom_text(data = vivi_matrix, aes(x = Var1, y = Var2, label = Interaction), 
            color = "black", size = 2.5)

```


###PDP Plots

We'll start with a PDP plot with age:

```{r}
pdp_age <- partial(
  object = rf_final_interaction, 
  pred.var = "age_yr", 
  train = clean_data_filtered,
  grid.resolution = 100,  # Number of points in PDP
  prob = TRUE  # Ensures output is probability-based
)

# Convert to data frame for plotting
pdp_age_df <- as.data.frame(pdp_age)

# Plot PDP for age
pdp_age_plot <- ggplot(pdp_age_df, aes(x = age_yr, y = (1 - yhat) * 100)) +
  geom_line(color = "black", size = 1) +  # PDP line
  geom_rug(data = clean_data_filtered, aes(x = age_yr), sides = "b", alpha = 0.3, inherit.aes = FALSE) +  # Add rug
  labs(title = "PDP for Age",
       x = "Age (Years)",
       y = "Pred. Mortality (%)") +
  theme_minimal()



#Display the plot
pdp_age_plot

```

Then, TTE:

```{r}
# Compute PDP for TTE using the pdp package
pdp_tte <- partial(
  object = rf_final_interaction, 
  pred.var = "TTE", 
  train = clean_data_filtered,
  grid.resolution = 100,  # Number of points in PDP
  prob = TRUE  # Ensures output is probability-based
)

# Convert to data frame for plotting
pdp_tte_df <- as.data.frame(pdp_tte)

# Add jitter to TTE values before plotting
jittered_tte <- clean_data_filtered %>%
  mutate(jittered_x = TTE + runif(n(), min = -0.5, max = 0.5))  # Small jitter to avoid overlap

# Plot PDP for TTE (using 1 - yhat)
pdp_tte_plot <- ggplot(pdp_tte_df, aes(x = TTE, y = (1 - yhat) * 100)) +
  geom_line(color = "black", size = 1) +  # PDP line
  geom_rug(data = jittered_tte, aes(x = jittered_x), 
           sides = "b", alpha = 0.3, inherit.aes = FALSE) +  # Jittered rug plot
  labs(title = "PDP for TTE",
       x = "Time to Event (TTE), Days",
       y = "Pred. Mortality (%)") +
  theme_minimal() +
  ylim(0, NA)  # Ensure the y-axis starts at 0

# Display the plot
pdp_tte_plot
```

Next concomitant_count:

```{r}
# Compute PDP for Concomitant Medication Count using the pdp package
pdp_concomitant <- partial(
  object = rf_final_interaction, 
  pred.var = "concomitant_count", 
  train = clean_data_filtered,
  grid.resolution = 100,  # Number of points in PDP
  prob = TRUE  # Ensures output is probability-based
)

# Convert to data frame for plotting
pdp_concomitant_df <- as.data.frame(pdp_concomitant)

# Add jitter to concomitant_count values before plotting
jittered_concomitant <- clean_data_filtered %>%
  mutate(jittered_x = concomitant_count + runif(n(), min = -0.2, max = 0.2))  # Manual jitter

# Plot PDP for Concomitant Medication Count (using 1 - yhat)
pdp_concomitant_plot <- ggplot(pdp_concomitant_df, aes(x = concomitant_count, y = (1 - yhat) * 100)) +
  geom_line(color = "black", size = 1) +  # PDP line
  geom_rug(data = jittered_concomitant, aes(x = jittered_x), 
           sides = "b", alpha = 0.3, inherit.aes = FALSE) +  
  labs(title = "PDP for Concomitant Med. Count",
       x = "Number of Concomitant Medications",
       y = "Pred. Mortality (%)") +
  theme_minimal() +
  ylim(0, NA)  # Ensure the y-axis starts at 0

# Display plot
pdp_concomitant_plot
```

Then, sex:

```{r}
# Compute Partial Dependence for sex
pdp_sex <- partial(
  object = rf_final_interaction,  # Trained random forest model
  pred.var = "sex",               # Feature of interest
  train = clean_data_filtered,    # Full dataset (ensures averaging over covariates)
  prob = TRUE                     # Use probabilities instead of logit scores
)

# Plot the PDP
pdp_sex_plot <- ggplot(pdp_sex, aes(x = sex, y = (1-yhat)*100, fill = sex)) +
  geom_bar(stat = "identity", color = "black") +
scale_fill_manual(values = c("pink", "lightblue", "lightgreen")) +  labs(title = "PDP for Sex",
       x = "Sex",
       y = "Pred. Mortality (%)") +
  theme_minimal() +
  theme(legend.position = "none") +
  ylim(0, NA)

#Display the plot
pdp_sex_plot
```



Then for region:

```{r}
# Compute PDP for Region using the pdp package
pdp_region <- partial(
  object = rf_final_interaction, 
  pred.var = "region", 
  train = clean_data_filtered,
  grid.resolution = length(unique(clean_data_filtered$region)),  # Number of points in PDP
  prob = TRUE  # Ensures output is probability-based
)

# Convert to data frame for plotting
pdp_region_df <- as.data.frame(pdp_region)

# Plot PDP for Region (with y-axis starting at 0)
pdp_region_plot <- ggplot(pdp_region_df, aes(x = region, y = (1 - yhat) * 100, fill = region)) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "PDP for Region",
       x = "Region",
       y = "Pred. Mortality (%)") +
  theme_minimal() +
  theme(legend.position = "none",  # Hide legend since x-axis has labels
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3, size = 8)) +  # Rotate x-axis labels
  ylim(0, NA)  # Ensures ymin starts at 0 while allowing ymax to adjust automatically

# Show the plot
pdp_region_plot
```


Now, for the computed mortality for each SCAR;

```{r}
# Define SCAR conditions to evaluate
scar_types <- c("SJSTEN", "DRESS", "AGEP", "GBFDE")

# Define consistent SCAR colors
scar_colors <- setNames(brewer.pal(5, "Set1"), c("SJSTEN", "DRESS", "AGEP", "GBFDE", "Total SCAR"))

# Initialize an empty list to store PDP results
pdp_scar_list <- list()

# Compute PDP for each SCAR type
for (scar in scar_types) {
  
  # Generate Partial Dependence Data
  pdp_result <- partial(
    object = rf_final_interaction,  # Trained RF model
    pred.var = scar,                # SCAR variable of interest
    train = clean_data_filtered,    # Full dataset for averaging across covariates
    prob = TRUE,                    # Ensure probability-based interpretation
    categorical = TRUE              # Explicitly treat SCAR as categorical
  )
  
  # Rename columns for consistency
  colnames(pdp_result) <- c("SCAR_Value", "Predicted_Prob")
  
  # Filter for SCAR = TRUE (1) only
  pdp_result <- pdp_result[pdp_result$SCAR_Value == 1, ]
  
  # Label the SCAR type
  pdp_result$SCAR_Type <- scar
  
  # Store in list
  pdp_scar_list[[scar]] <- pdp_result
}

# Combine PDP results into a single data frame
pdp_data_scar <- do.call(rbind, pdp_scar_list)

# Create the SCAR PDP Bar Plot
pdp_scar_plot <- ggplot(pdp_data_scar, aes(x = SCAR_Type, y = 100 * (1-Predicted_Prob), fill = SCAR_Type)) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_manual(values = scar_colors) +  # Apply custom SCAR colors
  labs(title = "Predicted Mortality by SCAR Type",
       x = "SCAR Type",
       y = "Pred. Mortality (%)") +
  theme_minimal() +
  theme(legend.position = "none") +  # Hide legend since labels are on x-axis
  ylim(0, NA)  # Ensure y-axis starts at 0 while allowing ymax to adjust automatically

# Display the plot
print(pdp_scar_plot)
```

###FINAL PLOTS

Now we can combine them in one big plot!

```{r}
# Combine the six PDPs into one vertical column
pdp_column1 <- pdp_age_plot / pdp_concomitant_plot / pdp_tte_plot + 
  plot_layout(ncol = 1)  # Stack vertically

pdp_column2 <- pdp_sex_plot / pdp_region_plot/ pdp_scar_plot + 
  plot_layout(ncol = 1)

# Combine into the left half of the figure
fig8_left_column <- pdp_column1 | pdp_column2

# Combine heatmap and metric table into the second column
fig8_right_column <- pVIVI / metric_table + 
  plot_layout(ncol = 1, heights = c(3, 1))  # Adjust heights (more space for heatmap)

# Or we can use the bootstrap VIMP instead of the VIVI plot
fig8_right_column_alt <- pVimpBootstrapIndividual / metric_table + 
  plot_layout(ncol = 1, heights = c(3, 1))  # Adjust heights

# Combine them
fig8_complete <- fig8_left_column | fig8_right_column
fig8_complete_alt <- fig8_left_column | fig8_right_column_alt

```


##Random forest by Individual SCAR

We will perform separate RFs for each individual SCAR as well, to remove the effect of that variable.

First, SJS-TEN only:

```{r}
# Define hyperparameter grid for `ranger`
tune_grid <- expand.grid(
  mtry = c(2, round(sqrt(ncol(clean_data_filtered_SJSTEN) - 1)), round(log2(ncol(clean_data_filtered_SJSTEN) - 1))),
  min.node.size = c(1, 5, 10),  # Controls tree depth
  splitrule = c("gini", "extratrees")  # Try both Gini and ExtraTrees
)

# Enable parallel processing
num_cores <- detectCores() - 1  # Use all but one core
cl <- makeCluster(num_cores)
registerDoParallel(cl)

# Define cross-validation strategy
control <- trainControl(
  method = "cv", 
  number = 5,
  classProbs = TRUE,
  verboseIter = TRUE  # Enables progress messages
)

# Run grid search with parallel processing
rf_tuned_interaction_sjsten <- train(
  Death ~ age_yr + sex + region + causative_drug + TTE + concomitant_count, 
  data = clean_data_filtered_SJSTEN, 
  method = "ranger",
  trControl = control,
  tuneGrid = tune_grid,
  num.trees = 500,  # Number of trees
  importance = "permutation",  # Enables variable importance computation
  seed = 42
)

# Stop parallel processing
stopCluster(cl)

# View best hyperparameters
print(rf_tuned_interaction_sjsten$bestTune)

# Train final optimized model
rf_final_interaction_sjsten <- ranger(
  formula = Death ~ age_yr + sex + region + causative_drug + TTE + concomitant_count,
  data = clean_data_filtered_SJSTEN,
  num.trees = 500,
  mtry = rf_tuned_interaction_sjsten$bestTune$mtry,
  min.node.size = rf_tuned_interaction_sjsten$bestTune$min.node.size,
  splitrule = rf_tuned_interaction_sjsten$bestTune$splitrule,
  probability = TRUE,  # Enables probability predictions
  importance = "permutation",
  seed = 42
)

# Extract variable importance scores
vimp_results_sjsten <- as.data.table(rf_final_interaction_sjsten$variable.importance, keep.rownames = TRUE)
setnames(vimp_results_sjsten, c("Variable", "Importance"))

# Print variable importance table (including joint SCAR VIMP)
print(vimp_results_sjsten)

# Plot VIMP (Including Grouped SCAR Importance)
ggplot(vimp_results_sjsten, aes(x = reorder(Variable, Importance), y = Importance)) +
  geom_bar(stat = "identity", fill = "steelblue", color = "black") +
  coord_flip() +
  labs(title = "Variable Importance (VIMP) in Mortality Prediction, SJS-TEN",
       x = "Variable",
       y = "VIMP Score") +
  theme_minimal()
```


Then, DRESS only:

```{r}
# Define hyperparameter grid for `ranger`
tune_grid <- expand.grid(
  mtry = c(2, round(sqrt(ncol(clean_data_filtered_DRESS) - 1)), round(log2(ncol(clean_data_filtered_DRESS) - 1))),
  min.node.size = c(1, 5, 10),  # Controls tree depth
  splitrule = c("gini", "extratrees")  # Try both Gini and ExtraTrees
)

# Enable parallel processing
num_cores <- detectCores() - 1  # Use all but one core
cl <- makeCluster(num_cores)
registerDoParallel(cl)

# Define cross-validation strategy
control <- trainControl(
  method = "cv", 
  number = 5,
  classProbs = TRUE,
  verboseIter = TRUE  # Enables progress messages
)

# Run grid search with parallel processing
rf_tuned_interaction_dress <- train(
  Death ~ age_yr + sex + region + causative_drug + TTE + concomitant_count, 
  data = clean_data_filtered_DRESS, 
  method = "ranger",
  trControl = control,
  tuneGrid = tune_grid,
  num.trees = 500,  # Number of trees
  importance = "permutation",  # Enables variable importance computation
  seed = 42
)

# Stop parallel processing
stopCluster(cl)

# View best hyperparameters
print(rf_tuned_interaction_dress$bestTune)

# Train final optimized model
rf_final_interaction_dress <- ranger(
  formula = Death ~ age_yr + sex + region + causative_drug + TTE + concomitant_count,
  data = clean_data_filtered_DRESS,
  num.trees = 500,
  mtry = rf_tuned_interaction_dress$bestTune$mtry,
  min.node.size = rf_tuned_interaction_dress$bestTune$min.node.size,
  splitrule = rf_tuned_interaction_dress$bestTune$splitrule,
  probability = TRUE,  # Enables probability predictions
  importance = "permutation",
  seed = 42
)

# Extract variable importance scores
vimp_results_dress <- as.data.table(rf_final_interaction_dress$variable.importance, keep.rownames = TRUE)
setnames(vimp_results_dress, c("Variable", "Importance"))

# Print variable importance table
print(vimp_results_dress)

# Plot VIMP
ggplot(vimp_results_dress, aes(x = reorder(Variable, Importance), y = Importance)) +
  geom_bar(stat = "identity", fill = "steelblue", color = "black") +
  coord_flip() +
  labs(title = "Variable Importance (VIMP) in Mortality Prediction, DRESS",
       x = "Variable",
       y = "VIMP Score") +
  theme_minimal()
```

Finally, AGEP only:

```{r}
# Define hyperparameter grid for `ranger`
tune_grid <- expand.grid(
  mtry = c(2, round(sqrt(ncol(clean_data_filtered_AGEP) - 1)), round(log2(ncol(clean_data_filtered_AGEP) - 1))),
  min.node.size = c(1, 5, 10),  # Controls tree depth
  splitrule = c("gini", "extratrees")  # Try both Gini and ExtraTrees
)

# Enable parallel processing
num_cores <- detectCores() - 1  # Use all but one core
cl <- makeCluster(num_cores)
registerDoParallel(cl)

# Define cross-validation strategy
control <- trainControl(
  method = "cv", 
  number = 5,
  classProbs = TRUE,
  verboseIter = TRUE  # Enables progress messages
)

# Run grid search with parallel processing
rf_tuned_interaction_agep <- train(
  Death ~ age_yr + sex + region + causative_drug + TTE + concomitant_count, 
  data = clean_data_filtered_AGEP, 
  method = "ranger",
  trControl = control,
  tuneGrid = tune_grid,
  num.trees = 500,  # Number of trees
  importance = "permutation",  # Enables variable importance computation
  seed = 42
)

# Stop parallel processing
stopCluster(cl)

# View best hyperparameters
print(rf_tuned_interaction_agep$bestTune)

# Train final optimized model
rf_final_interaction_agep <- ranger(
  formula = Death ~ age_yr + sex + region + causative_drug + TTE + concomitant_count,
  data = clean_data_filtered_AGEP,
  num.trees = 500,
  mtry = rf_tuned_interaction_agep$bestTune$mtry,
  min.node.size = rf_tuned_interaction_agep$bestTune$min.node.size,
  splitrule = rf_tuned_interaction_agep$bestTune$splitrule,
  probability = TRUE,  # Enables probability predictions
  importance = "permutation",
  seed = 42
)

# Extract variable importance scores
vimp_results_agep <- as.data.table(rf_final_interaction_agep$variable.importance, keep.rownames = TRUE)
setnames(vimp_results_agep, c("Variable", "Importance"))

# Print variable importance table
print(vimp_results_agep)

# Plot VIMP
ggplot(vimp_results_agep, aes(x = reorder(Variable, Importance), y = Importance)) +
  geom_bar(stat = "identity", fill = "steelblue", color = "black") +
  coord_flip() +
  labs(title = "Variable Importance (VIMP) in Mortality Prediction, AGEP",
       x = "Variable",
       y = "VIMP Score") +
  theme_minimal()
```

###Validation

Now we validate these:

```{r}
# OOB Error Rate
oob_error_sjsten <- rf_final_interaction_sjsten$prediction.error
oob_error_dress <- rf_final_interaction_dress$prediction.error
oob_error_agep <- rf_final_interaction_agep$prediction.error

# Print results
print(paste("OOB Error Rate SJS-TEN:", oob_error_sjsten))
print(paste("OOB Error Rate DRESS:", oob_error_dress))
print(paste("OOB Error Rate AGEP:", oob_error_agep))
```
Then get AUCs

```{r}
# Get OOB Predicted Probabilities & True Labels
oob_predictions_sjsten <- rf_final_interaction_sjsten$predictions
oob_predictions_dress <- rf_final_interaction_dress$predictions
oob_predictions_agep <- rf_final_interaction_agep$predictions

# Ensure Death labels are a binary factor with correct levels
true_labels_sjsten <- factor(clean_data_filtered_SJSTEN$Death, levels = c("No", "Yes"))
true_labels_dress <- factor(clean_data_filtered_DRESS$Death, levels = c("No", "Yes"))  
true_labels_agep <- factor(clean_data_filtered_AGEP$Death, levels = c("No", "Yes"))  

# Compute AUC using OOB probabilities
roc_curve_sjsten <- roc(response = true_labels_sjsten, predictor = oob_predictions_sjsten[, "Yes"], levels = c("No", "Yes"))
roc_curve_dress <- roc(response = true_labels_dress, predictor = oob_predictions_dress[, "Yes"], levels = c("No", "Yes"))
roc_curve_agep <- roc(response = true_labels_agep, predictor = oob_predictions_agep[, "Yes"], levels = c("No", "Yes"))

# Extract AUC value
auc_value_sjsten <- auc(roc_curve_sjsten)
auc_value_dress <- auc(roc_curve_dress)
auc_value_agep <- auc(roc_curve_agep)

# Print the corrected AUC value
print(paste("Corrected AUC (SJS-TEN):", round(auc_value_sjsten, 3)))
print(paste("Corrected AUC (DRESS):", round(auc_value_dress, 3)))
print(paste("Corrected AUC (AGEP):", round(auc_value_agep, 3)))
```
Then get Brier scores:

```{r}
# Compute the Brier Score
brier_score_sjsten <- mean((clean_data_filtered_SJSTEN$Pred_Death_Prob - as.numeric(clean_data_filtered_SJSTEN$Death == "Yes"))^2, na.rm = TRUE)
brier_score_dress <- mean((clean_data_filtered_DRESS$Pred_Death_Prob - as.numeric(clean_data_filtered_DRESS$Death == "Yes"))^2, na.rm = TRUE)
brier_score_agep <- mean((clean_data_filtered_AGEP$Pred_Death_Prob - as.numeric(clean_data_filtered_AGEP$Death == "Yes"))^2, na.rm = TRUE)

# Print result
print(paste("Brier Score (SJS-TEN):", round(brier_score_sjsten, 4)))
print(paste("Brier Score (DRESS):", round(brier_score_dress, 4)))
print(paste("Brier Score (AGEP):", round(brier_score_agep, 4)))
```

### Variable Importance (by Bootstrap)

Let's get our bootstrap VIMPs for each individual SCAR RF

```{r}
# Define number of bootstrap iterations
num_bootstraps <- 100

# Store VIMP results from bootstraps
vimp_list_sjsten <- vector("list", num_bootstraps)
vimp_list_dress <- vector("list", num_bootstraps)
vimp_list_agep <- vector("list", num_bootstraps)

# Bootstrapping loop
for (i in 1:num_bootstraps) {
  cat("Bootstrap iteration:", i, "\n")
  
  # Sample data with replacement
  bootstrap_sample_sjsten <- clean_data_filtered_SJSTEN[sample(.N, replace = TRUE)]
    bootstrap_sample_dress <- clean_data_filtered_DRESS[sample(.N, replace = TRUE)]
    bootstrap_sample_agep <- clean_data_filtered_AGEP[sample(.N, replace = TRUE)]
  
  # Train a new random forest model
  rf_bootstrap_sjsten <- ranger(
    Death ~ age_yr + sex + region + causative_drug + TTE + concomitant_count,
    data = bootstrap_sample_sjsten, 
    num.trees = 500,
    importance = "permutation",
    probability = TRUE
  )
  rf_bootstrap_dress <- ranger(
    Death ~ age_yr + sex + region + causative_drug + TTE + concomitant_count,
    data = bootstrap_sample_dress, 
    num.trees = 500,
    importance = "permutation",
    probability = TRUE
  )
  rf_bootstrap_agep <- ranger(
    Death ~ age_yr + sex + region + causative_drug + TTE + concomitant_count,
    data = bootstrap_sample_agep, 
    num.trees = 500,
    importance = "permutation",
    probability = TRUE
  )
  
  # Store variable importance scores
  vimp_list_sjsten[[i]] <- as.data.table(rf_bootstrap_sjsten$variable.importance, keep.rownames = TRUE)
  vimp_list_dress[[i]] <- as.data.table(rf_bootstrap_dress$variable.importance, keep.rownames = TRUE)
  vimp_list_agep[[i]] <- as.data.table(rf_bootstrap_agep$variable.importance, keep.rownames = TRUE)
}
```
Now plot results for SJS-TEN

```{r}
# Combine all bootstrap VIMP results
vimp_results_sjsten <- rbindlist(vimp_list_sjsten, use.names = TRUE, fill = TRUE)
setnames(vimp_results_sjsten, c("Variable", "Importance"))

# Compute summary statistics for each variable
vimp_summary_sjsten <- vimp_results_sjsten[, .(
  Mean_VIMP = mean(Importance),
  SD_VIMP = sd(Importance),
  Lower_CI = quantile(Importance, 0.025),
  Upper_CI = quantile(Importance, 0.975)
), by = Variable]


# Print bootstrapped VIMP summary with confidence intervals
print(vimp_summary_sjsten)

# **Plot Bootstrapped VIMP (Including Individual SCAR VIMP, Removing Joint SCAR VIMP)**
pVimpBootstrapSJSTEN <- ggplot(vimp_summary_sjsten, aes(x = reorder(Variable, Mean_VIMP), y = Mean_VIMP)) +
  
  # **Boxplot for all variables (INCLUDING SCAR subtypes)**
  geom_boxplot(data = vimp_results_sjsten, 
               aes(x = reorder(Variable, Importance), y = Importance),
               fill = "lightblue", color = "black", outlier.shape = NA) +
  
  # **Jittered points for individual VIMP values**
  geom_jitter(data = vimp_results_sjsten, 
              aes(x = Variable, y = Importance), 
              width = 0.2, alpha = 0.3, color = "darkblue") +

  # **Flip coordinates for readability**
  coord_flip() +
  
  # **Titles and Labels**
  labs(title = "Bootstrapped Variable Importance (VIMP), SJS-TEN",
       x = "Variable",
       y = "Importance Score") +
  
  # **Minimalist Theme**
  theme_minimal()
```
Then do DRESS:

```{r}
# Combine all bootstrap VIMP results
vimp_results_dress <- rbindlist(vimp_list_dress, use.names = TRUE, fill = TRUE)
setnames(vimp_results_dress, c("Variable", "Importance"))

# Compute summary statistics for each variable
vimp_summary_dress <- vimp_results_dress[, .(
  Mean_VIMP = mean(Importance),
  SD_VIMP = sd(Importance),
  Lower_CI = quantile(Importance, 0.025),
  Upper_CI = quantile(Importance, 0.975)
), by = Variable]


# Print bootstrapped VIMP summary with confidence intervals
print(vimp_summary_dress)

# **Plot Bootstrapped VIMP (Including Individual SCAR VIMP, Removing Joint SCAR VIMP)**
pVimpBootstrapDRESS <- ggplot(vimp_summary_dress, aes(x = reorder(Variable, Mean_VIMP), y = Mean_VIMP)) +
  
  # **Boxplot for all variables (INCLUDING SCAR subtypes)**
  geom_boxplot(data = vimp_results_dress, 
               aes(x = reorder(Variable, Importance), y = Importance),
               fill = "lightblue", color = "black", outlier.shape = NA) +
  
  # **Jittered points for individual VIMP values**
  geom_jitter(data = vimp_results_dress, 
              aes(x = Variable, y = Importance), 
              width = 0.2, alpha = 0.3, color = "darkblue") +

  # **Flip coordinates for readability**
  coord_flip() +
  
  # **Titles and Labels**
  labs(title = "Bootstrapped Variable Importance (VIMP), DRESS",
       x = "Variable",
       y = "Importance Score") +
  
  # **Minimalist Theme**
  theme_minimal()
```

Then do AGEP:

```{r}
# Combine all bootstrap VIMP results
vimp_results_agep <- rbindlist(vimp_list_agep, use.names = TRUE, fill = TRUE)
setnames(vimp_results_agep, c("Variable", "Importance"))

# Compute summary statistics for each variable
vimp_summary_agep <- vimp_results_agep[, .(
  Mean_VIMP = mean(Importance),
  SD_VIMP = sd(Importance),
  Lower_CI = quantile(Importance, 0.025),
  Upper_CI = quantile(Importance, 0.975)
), by = Variable]


# Print bootstrapped VIMP summary with confidence intervals
print(vimp_summary_agep)

# **Plot Bootstrapped VIMP**
pVimpBootstrapAGEP <- ggplot(vimp_summary_agep, aes(x = reorder(Variable, Mean_VIMP), y = Mean_VIMP)) +
  
  # **Boxplot for all variables (INCLUDING SCAR subtypes)**
  geom_boxplot(data = vimp_results_agep, 
               aes(x = reorder(Variable, Importance), y = Importance),
               fill = "lightblue", color = "black", outlier.shape = NA) +
  
  # **Jittered points for individual VIMP values**
  geom_jitter(data = vimp_results_agep, 
              aes(x = Variable, y = Importance), 
              width = 0.2, alpha = 0.3, color = "darkblue") +

  # **Flip coordinates for readability**
  coord_flip() +
  
  # **Titles and Labels**
  labs(title = "Bootstrapped Variable Importance (VIMP), AGEP",
       x = "Variable",
       y = "Importance Score") +
  
  # **Minimalist Theme**
  theme_minimal()
```

###VIVI Plots

VIVI plots are a method of simultaneously plotting VIMPs and VINTs (which measure the strength of the interaction between two variables). Do the VIVI Plot for SJS-TEN:

```{r}
# Define feature names (excluding the response variable "Death")
feature_names <- c("age_yr", "sex", "region", "causative_drug", "TTE", "concomitant_count")

# Ensure Death is a factor
clean_data_filtered_SJSTEN[, Death := as.factor(Death)]

# 🔹 Compute VIVI matrix
vivi_results_sjsten <- vivi(
  data = clean_data_filtered_SJSTEN[, c("Death", feature_names), with = FALSE],  # Data with response + predictors
  fit = rf_final_interaction_sjsten,  # Trained ranger random forest model
  response = "Death",  # Response variable
  gridSize = 50,  # Grid size for evaluating predictions
  importanceType = "permutation",
  nmax = NULL,  # Use all data rows
  reorder = TRUE,  # Reorder matrix for better visualization
  class = "Yes",  # Factor level of the response variable to analyze
  numPerm = 4,  # Number of permutations for importance
  showVimpError = FALSE  # Suppress standard error printout
)

# 🔹 Convert `vivi_results` matrix to a long format data table
vivi_matrix_sjsten <- as.data.table(melt(vivi_results_sjsten, varnames = c("Var1", "Var2"), value.name = "Interaction"))
vivi_matrix_sjsten[, Interaction := round(Interaction, 3)]  # Round values for display

# 🔹 Generate the VIVI Heatmap with Text Labels
vivi_plot_sjsten <- viviHeatmap(vivi_results_sjsten) +
  ggtitle("Variable Importance & Interaction Strengths, SJS-TEN") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 0))  # Rotate x-axis labels

# 🔹 Overlay text labels onto heatmap
pVIVI_sjsten <- vivi_plot_sjsten + 
  geom_text(data = vivi_matrix_sjsten, aes(x = Var1, y = Var2, label = Interaction), 
            color = "black", size = 4)
```


Next, DRESS:

```{r}
# Define feature names (excluding the response variable "Death")
feature_names <- c("age_yr", "sex", "region", "causative_drug", "TTE", "concomitant_count")

# Ensure Death is a factor
clean_data_filtered_DRESS[, Death := as.factor(Death)]

# 🔹 Compute VIVI matrix
vivi_results_dress <- vivi(
  data = clean_data_filtered_DRESS[, c("Death", feature_names), with = FALSE],  # Data with response + predictors
  fit = rf_final_interaction_dress,  # Trained ranger random forest model
  response = "Death",  # Response variable
  gridSize = 50,  # Grid size for evaluating predictions
  importanceType = "permutation",
  nmax = NULL,  # Use all data rows
  reorder = TRUE,  # Reorder matrix for better visualization
  class = "Yes",  # Factor level of the response variable to analyze
  numPerm = 4,  # Number of permutations for importance
  showVimpError = FALSE  # Suppress standard error printout
)

# 🔹 Convert `vivi_results` matrix to a long format data table
vivi_matrix_dress <- as.data.table(melt(vivi_results_dress, varnames = c("Var1", "Var2"), value.name = "Interaction"))
vivi_matrix_dress[, Interaction := round(Interaction, 3)]  # Round values for display

# 🔹 Generate the VIVI Heatmap with Text Labels
vivi_plot_dress <- viviHeatmap(vivi_results_dress) +
  ggtitle("Variable Importance & Interaction Strengths, DRESS") +
  theme(axis.text.x = element_text(angle = 45, hjust = 0))  # Rotate x-axis labels +
  theme_minimal()

# 🔹 Overlay text labels onto heatmap
pVIVI_dress <- vivi_plot_dress + 
  geom_text(data = vivi_matrix_dress, aes(x = Var1, y = Var2, label = Interaction), 
            color = "black", size = 4)
```


Now AGEP:

```{r}
# Define feature names (excluding the response variable "Death")
feature_names <- c("age_yr", "sex", "region", "causative_drug", "TTE", "concomitant_count")

# Ensure Death is a factor
clean_data_filtered_AGEP[, Death := as.factor(Death)]

# 🔹 Compute VIVI matrix
vivi_results_agep <- vivi(
  data = clean_data_filtered_AGEP[, c("Death", feature_names), with = FALSE],  # Data with response + predictors
  fit = rf_final_interaction_agep,  # Trained ranger random forest model
  response = "Death",  # Response variable
  gridSize = 50,  # Grid size for evaluating predictions
  importanceType = "permutation",
  nmax = NULL,  # Use all data rows
  reorder = TRUE,  # Reorder matrix for better visualization
  class = "Yes",  # Factor level of the response variable to analyze
  numPerm = 4,  # Number of permutations for importance
  showVimpError = FALSE  # Suppress standard error printout
)

# 🔹 Convert `vivi_results` matrix to a long format data table
vivi_matrix_agep <- as.data.table(melt(vivi_results_agep, varnames = c("Var1", "Var2"), value.name = "Interaction"))
vivi_matrix_agep[, Interaction := round(Interaction, 3)]  # Round values for display

# 🔹 Generate the VIVI Heatmap with Text Labels
vivi_plot_agep <- viviHeatmap(vivi_results_agep) +
  ggtitle("Variable Importance & Interaction Strengths, AGEP") +
  theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 0))  # Rotate x-axis labels

# 🔹 Overlay text labels onto heatmap
pVIVI_agep <- vivi_plot_agep + 
  geom_text(data = vivi_matrix_agep, aes(x = Var1, y = Var2, label = Interaction), 
            color = "black", size = 4)
```

###PDP Plots (SJS-TEN)

We'll generate the same plots for SJS-TEN:

```{r}
# Define PDP for Age (SJS-TEN)
pdp_age_sjs <- partial(
  object = rf_final_interaction_sjsten, 
  pred.var = "age_yr", 
  train = clean_data_filtered_SJSTEN,
  grid.resolution = 100, 
  prob = TRUE
)

# Add jitter to age values before plotting
jittered_age_sjs <- clean_data_filtered_SJSTEN %>%
  mutate(jittered_x = age_yr + runif(n(), min = -0.5, max = 0.5))  # Small jitter to avoid overlap

# Convert to data frame and plot
pdp_age_sjs_plot <- ggplot(as.data.frame(pdp_age_sjs), aes(x = age_yr, y = 100 * (1 - yhat))) +
  geom_line(color = "black", size = 1) +  # PDP line
  geom_rug(data = jittered_age_sjs, aes(x = jittered_x), 
           sides = "b", alpha = 0.3, inherit.aes = FALSE) +  # Jittered rug plot
  scale_y_continuous(limits = c(0, NA)) +
  labs(title = "PDP for Age (SJS-TEN)",
       x = "Age (Years)",
       y = "Pred. Mortality (%)") +
  theme_minimal()

# Define PDP for TTE (SJS-TEN)
pdp_tte_sjs <- partial(
  object = rf_final_interaction_sjsten, 
  pred.var = "TTE", 
  train = clean_data_filtered_SJSTEN,
  grid.resolution = 100, 
  prob = TRUE
)

# Add jitter to TTE values before plotting
jittered_tte_sjs <- clean_data_filtered_SJSTEN %>%
  mutate(jittered_x = TTE + runif(n(), min = -0.5, max = 0.5))  # Small jitter to avoid overlap

# Convert to data frame and plot
pdp_tte_sjs_plot <- ggplot(as.data.frame(pdp_tte_sjs), aes(x = TTE, y = 100 * (1 - yhat))) +
  geom_line(color = "black", size = 1) +  # PDP line
  geom_rug(data = jittered_tte_sjs, aes(x = jittered_x), 
           sides = "b", alpha = 0.3, inherit.aes = FALSE) +  # Jittered rug plot
  scale_y_continuous(limits = c(0, NA)) +
  labs(title = "PDP for TTE (SJS-TEN)",
       x = "Time to Event (TTE), Days",
       y = "Pred. Mortality (%)") +
  theme_minimal()

# Define PDP for Concomitant Medications (SJS-TEN)
pdp_concomitant_sjs <- partial(
  object = rf_final_interaction_sjsten, 
  pred.var = "concomitant_count", 
  train = clean_data_filtered_SJSTEN,
  grid.resolution = 100, 
  prob = TRUE
)

# Add jitter to Concomitant Medication Count before plotting
jittered_concomitant_sjs <- clean_data_filtered_SJSTEN %>%
  mutate(jittered_x = concomitant_count + runif(n(), min = -0.2, max = 0.2))  # Small jitter to avoid overlap

# Convert to data frame and plot
pdp_concomitant_sjs_plot <- ggplot(as.data.frame(pdp_concomitant_sjs), aes(x = concomitant_count, y = 100 * (1 - yhat))) +
  geom_line(color = "black", size = 1) +  # PDP line
  geom_rug(data = jittered_concomitant_sjs, aes(x = jittered_x), 
           sides = "b", alpha = 0.3, inherit.aes = FALSE) +  # Jittered rug plot
  scale_y_continuous(limits = c(0, NA)) +
  labs(title = "PDP for Concomitant Med. Count (SJS-TEN)",
       x = "Number of Concomitant Medications",
       y = "Pred. Mortality (%)") +
  theme_minimal()

# Define PDP for Sex (SJS-TEN)
pdp_sex_sjs <- partial(
  object = rf_final_interaction_sjsten, 
  pred.var = "sex", 
  train = clean_data_filtered_SJSTEN,
  prob = TRUE
)

# Convert to data frame and plot
pdp_sex_sjs_plot <- ggplot(as.data.frame(pdp_sex_sjs), aes(x = sex, y = 100 * (1 - yhat), fill = sex)) +
  geom_bar(stat = "identity", color = "black") +
  scale_y_continuous(limits = c(0, NA)) +
  scale_fill_manual(values = c("pink", "lightblue", "lightgreen")) +
  labs(title = "PDP for Sex (SJS-TEN)",
       x = "Sex",
       y = "Pred. Mortality (%)") +
  theme_minimal() +
  theme(legend.position = "none")

# Define PDP for Region (SJS-TEN)
pdp_region_sjs <- partial(
  object = rf_final_interaction_sjsten, 
  pred.var = "region", 
  train = clean_data_filtered_SJSTEN,
  prob = TRUE
)

# Convert to data frame and plot
pdp_region_sjs_plot <- ggplot(as.data.frame(pdp_region_sjs), aes(x = region, y = 100 * (1 - yhat), fill = region)) +
  geom_bar(stat = "identity", color = "black") +
  scale_y_continuous(limits = c(0, NA)) +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "PDP for Region (SJS-TEN)",
       x = "Region",
       y = "Pred. Mortality (%)") +
  theme_minimal() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, hjust = 1))

# Print plots
print(pdp_age_sjs_plot)
print(pdp_tte_sjs_plot)
print(pdp_concomitant_sjs_plot)
print(pdp_sex_sjs_plot)
print(pdp_region_sjs_plot)

```

Create the metric table:

```{r}
# Define the metrics
metrics_sjsten <- data.table(
  Metric = c("OOB Error/Brier Score", "AUC"),
  Value = c(round(oob_error_sjsten, 4), round(auc_value_sjsten, 4))
)

# Convert to a gt table for better visualization
metric_table_sjsten <- gt(metrics_sjsten) %>%
  tab_header(title = "Model Performance Metrics (SJS-TEN)") %>%
  cols_label(
    Metric = "Metric",
    Value = "Value"
  ) %>%
  fmt_number(columns = Value, decimals = 4) %>%
  tab_options(
    table.font.size = px(14),
    heading.title.font.size = px(16),
    heading.title.font.weight = "bold"
  )

# Print the table
metric_table_sjsten
```

Now create the final combined plot

```{r}
# Combine three PDPs into one vertical column
pdp_column1_sjsten <- pdp_age_sjs_plot / pdp_concomitant_sjs_plot / pdp_tte_sjs_plot + 
  plot_layout(ncol = 1)  # Stack vertically

# combine discrete PDPs into one vertical column
pdp_column2_sjsten <- pdp_region_sjs_plot / pdp_sex_sjs_plot 
  plot_layout(ncol = 1)

#Combine column 1 and 2 into the left column
left_column_sjsten <- pdp_column1_sjsten | pdp_column2_sjsten

# Combine heatmap and metric table into the second column
right_column_sjsten <- pVIVI_sjsten / metric_table_sjsten + 
  plot_layout(ncol = 1, heights = c(3, 1))  # Adjust heights (more space for heatmap)

# Or we can use the bootstrap VIMP instead of the VIVI plot
right_column_sjsten_alt <- pVimpBootstrapSJSTEN / metric_table_sjsten + 
  plot_layout(ncol = 1, heights = c(3, 1))  # Adjust heights

#Combine the figures
figS10_complete <- left_column_sjsten | right_column_sjsten
figS10_complete_alt <- left_column_sjsten | right_column_sjsten_alt
```

###PDP Plots (DRESS)

We'll create the same plots for DRESS

```{r}
# Define PDP for Age (DRESS)
pdp_age_dress <- partial(
  object = rf_final_interaction_dress, 
  pred.var = "age_yr", 
  train = clean_data_filtered_DRESS,
  grid.resolution = 100, 
  prob = TRUE
)

# Define PDP for TTE (DRESS)
pdp_tte_dress <- partial(
  object = rf_final_interaction_dress, 
  pred.var = "TTE", 
  train = clean_data_filtered_DRESS,
  grid.resolution = 100, 
  prob = TRUE
)

# Define PDP for Concomitant Medications (DRESS)
pdp_concomitant_dress <- partial(
  object = rf_final_interaction_dress, 
  pred.var = "concomitant_count", 
  train = clean_data_filtered_DRESS,
  grid.resolution = 100, 
  prob = TRUE
)

# Add jitter to age values before plotting
jittered_age_dress <- clean_data_filtered_DRESS %>%
  mutate(jittered_x = age_yr + runif(n(), min = -0.5, max = 0.5))  # Small jitter to avoid overlap

# Convert age to data frame and plot
pdp_age_dress_plot <- ggplot(as.data.frame(pdp_age_dress), aes(x = age_yr, y = 100 * (1 - yhat))) +
  geom_line(color = "black", size = 1) +  # PDP line
  geom_rug(data = jittered_age_dress, aes(x = jittered_x), 
           sides = "b", alpha = 0.3, inherit.aes = FALSE) +  # Jittered rug plot
  scale_y_continuous(limits = c(0, NA)) +
  labs(title = "PDP for Age (DRESS)",
       x = "Age (Years)",
       y = "Pred. Mortality (%)") +
  theme_minimal()

# Add jitter to TTE values before plotting
jittered_tte_dress <- clean_data_filtered_DRESS %>%
  mutate(jittered_x = TTE + runif(n(), min = -1, max = 1))  # Small jitter to avoid overlap

# Convert TTE to data frame and plot
pdp_tte_dress_plot <- ggplot(as.data.frame(pdp_tte_dress), aes(x = TTE, y = 100 * (1 - yhat))) +
  geom_line(color = "black", size = 1) +  # PDP line
  geom_rug(data = jittered_tte_dress, aes(x = jittered_x), 
           sides = "b", alpha = 0.3, inherit.aes = FALSE) +  # Jittered rug plot
  scale_y_continuous(limits = c(0, NA)) +
  labs(title = "PDP for TTE (DRESS)",
       x = "Time to Event (TTE), Days",
       y = "Pred. Mortality (%)") +
  theme_minimal()

# Add jitter to Concomitant Medication Count before plotting
jittered_concomitant_dress <- clean_data_filtered_DRESS %>%
  mutate(jittered_x = concomitant_count + runif(n(), min = -0.2, max = 0.2))  # Small jitter to avoid overlap

# Convert concomitant to data frame and plot
pdp_concomitant_dress_plot <- ggplot(as.data.frame(pdp_concomitant_dress), aes(x = concomitant_count, y = 100 * (1 - yhat))) +
  geom_line(color = "black", size = 1) +  # PDP line
  geom_rug(data = jittered_concomitant_dress, aes(x = jittered_x), 
           sides = "b", alpha = 0.3, inherit.aes = FALSE) +  # Jittered rug plot
  scale_y_continuous(limits = c(0, NA)) +
  labs(title = "PDP for Concomitant Med. Count (DRESS)",
       x = "Number of Concomitant Medications",
       y = "Pred. Mortality (%)") +
  theme_minimal()

# Define PDP for Sex (DRESS)
pdp_sex_dress <- partial(
  object = rf_final_interaction_dress, 
  pred.var = "sex", 
  train = clean_data_filtered_DRESS,
  prob = TRUE
)

# Convert to data frame and plot
pdp_sex_dress_plot <- ggplot(as.data.frame(pdp_sex_dress), aes(x = sex, y = 100 * (1 - yhat), fill = sex)) +
  geom_bar(stat = "identity", color = "black") +
  scale_y_continuous(limits = c(0, NA)) +
  scale_fill_manual(values = c("pink", "lightblue", "lightgreen")) +
  labs(title = "PDP for Sex (DRESS)",
       x = "Sex",
       y = "Pred. Mortality (%)") +
  theme_minimal() +
  theme(legend.position = "none")

# Define PDP for Region (DRESS)
pdp_region_dress <- partial(
  object = rf_final_interaction_dress, 
  pred.var = "region", 
  train = clean_data_filtered_DRESS,
  prob = TRUE
)

# Convert to data frame and plot
pdp_region_dress_plot <- ggplot(as.data.frame(pdp_region_dress), aes(x = region, y = 100 * (1 - yhat), fill = region)) +
  geom_bar(stat = "identity", color = "black") +
  scale_y_continuous(limits = c(0, NA)) +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "PDP for Region (DRESS)",
       x = "Region",
       y = "Pred. Mortality (%)") +
  theme_minimal() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, hjust = 1))

# Print plots
print(pdp_age_dress_plot)
print(pdp_tte_dress_plot)
print(pdp_concomitant_dress_plot)
print(pdp_sex_dress_plot)
print(pdp_region_dress_plot)
```

Create the metric table:

```{r}
# Define the metrics
metrics_dress <- data.table(
  Metric = c("OOB Error/Brier Score", "AUC"),
  Value = c(round(oob_error_dress, 4), round(auc_value_dress, 4))
)

# Convert to a gt table for better visualization
metric_table_dress <- gt(metrics_dress) %>%
  tab_header(title = "Model Performance Metrics (DRESS)") %>%
  cols_label(
    Metric = "Metric",
    Value = "Value"
  ) %>%
  fmt_number(columns = Value, decimals = 4) %>%
  tab_options(
    table.font.size = px(14),
    heading.title.font.size = px(16),
    heading.title.font.weight = "bold"
  )

# Print the table
metric_table_dress
```

Now create the final combined plot

```{r}
# Combine three PDPs into one vertical column
pdp_column1_dress <- pdp_age_dress_plot / pdp_concomitant_dress_plot / pdp_tte_dress_plot + 
  plot_layout(ncol = 1)  # Stack vertically

# Combine discrete PDPs into one vertical column
pdp_column2_dress <- pdp_region_dress_plot / pdp_sex_dress_plot + 
  plot_layout(ncol = 1)

# Combine column 1 and 2 into the left column
left_column_dress <- pdp_column1_dress | pdp_column2_dress

# Combine heatmap and metric table into the second column
right_column_dress <- pVIVI_dress / metric_table_dress + 
  plot_layout(ncol = 1, heights = c(3, 1))  # Adjust heights (more space for heatmap)

# Alternative layout with bootstrapped VIMP
right_column_dress_alt <- pVimpBootstrapDRESS / metric_table_dress + 
  plot_layout(ncol = 1, heights = c(3, 1))  # Adjust heights (more space for heatmap)

# Combine the figures
figS11_complete <- left_column_dress | right_column_dress
figS11_complete_alt <- left_column_dress | right_column_dress_alt
```

###PDP Plots (AGEP)

We'll generate the same plots for AGEP:

```{r}
# Define PDP for Age (AGEP)
pdp_age_agep <- partial(
  object = rf_final_interaction_agep, 
  pred.var = "age_yr", 
  train = clean_data_filtered_AGEP,
  grid.resolution = 100, 
  prob = TRUE
)

# Define PDP for TTE (AGEP)
pdp_tte_agep <- partial(
  object = rf_final_interaction_agep, 
  pred.var = "TTE", 
  train = clean_data_filtered_AGEP,
  grid.resolution = 100, 
  prob = TRUE
)

# Define PDP for Concomitant Medications (AGEP)
pdp_concomitant_agep <- partial(
  object = rf_final_interaction_agep, 
  pred.var = "concomitant_count", 
  train = clean_data_filtered_AGEP,
  grid.resolution = 100, 
  prob = TRUE
)

# Add jitter to age values before plotting
jittered_age_agep <- clean_data_filtered_AGEP %>%
  mutate(jittered_x = age_yr + runif(n(), min = -0.5, max = 0.5))  # Small jitter to avoid overlap

# Convert age to data frame and plot
pdp_age_agep_plot <- ggplot(as.data.frame(pdp_age_agep), aes(x = age_yr, y = 100 * (1 - yhat))) +
  geom_line(color = "black", size = 1) +  # PDP line
  geom_rug(data = jittered_age_agep, aes(x = jittered_x), 
           sides = "b", alpha = 0.3, inherit.aes = FALSE) +  # Jittered rug plot
  scale_y_continuous(limits = c(0, NA)) +
  labs(title = "PDP for Age (AGEP)",
       x = "Age (Years)",
       y = "Pred. Mortality (%)") +
  theme_minimal()

# Add jitter to TTE values before plotting
jittered_tte_agep <- clean_data_filtered_AGEP %>%
  mutate(jittered_x = TTE + runif(n(), min = -1, max = 1))  # Small jitter to avoid overlap

# Convert TTE to data frame and plot
pdp_tte_agep_plot <- ggplot(as.data.frame(pdp_tte_agep), aes(x = TTE, y = 100 * (1 - yhat))) +
  geom_line(color = "black", size = 1) +  # PDP line
  geom_rug(data = jittered_tte_agep, aes(x = jittered_x), 
           sides = "b", alpha = 0.3, inherit.aes = FALSE) +  # Jittered rug plot
  scale_y_continuous(limits = c(0, NA)) +
  labs(title = "PDP for TTE (AGEP)",
       x = "Time to Event (TTE), Days",
       y = "Pred. Mortality (%)") +
  theme_minimal()

# Add jitter to Concomitant Medication Count before plotting
jittered_concomitant_agep <- clean_data_filtered_AGEP %>%
  mutate(jittered_x = concomitant_count + runif(n(), min = -0.2, max = 0.2))  # Small jitter to avoid overlap

# Convert concomitant to data frame and plot
pdp_concomitant_agep_plot <- ggplot(as.data.frame(pdp_concomitant_agep), aes(x = concomitant_count, y = 100 * (1 - yhat))) +
  geom_line(color = "black", size = 1) +  # PDP line
  geom_rug(data = jittered_concomitant_agep, aes(x = jittered_x), 
           sides = "b", alpha = 0.3, inherit.aes = FALSE) +  # Jittered rug plot
  scale_y_continuous(limits = c(0, NA)) +
  labs(title = "PDP for Concomitant Med. Count (AGEP)",
       x = "Number of Concomitant Medications",
       y = "Pred. Mortality (%)") +
  theme_minimal()

# Define PDP for Sex (AGEP)
pdp_sex_agep <- partial(
  object = rf_final_interaction_agep, 
  pred.var = "sex", 
  train = clean_data_filtered_AGEP,
  prob = TRUE
)

# Convert to data frame and plot
pdp_sex_agep_plot <- ggplot(as.data.frame(pdp_sex_agep), aes(x = sex, y = 100 * (1 - yhat), fill = sex)) +
  geom_bar(stat = "identity", color = "black") +
  scale_y_continuous(limits = c(0, NA)) +
  scale_fill_manual(values = c("pink", "lightblue", "lightgreen")) +
  labs(title = "PDP for Sex (AGEP)",
       x = "Sex",
       y = "Pred. Mortality (%)") +
  theme_minimal() +
  theme(legend.position = "none")

# Define PDP for Region (AGEP)
pdp_region_agep <- partial(
  object = rf_final_interaction_agep, 
  pred.var = "region", 
  train = clean_data_filtered_AGEP,
  prob = TRUE
)

# Convert to data frame and plot
pdp_region_agep_plot <- ggplot(as.data.frame(pdp_region_agep), aes(x = region, y = 100 * (1 - yhat), fill = region)) +
  geom_bar(stat = "identity", color = "black") +
  scale_y_continuous(limits = c(0, NA)) +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "PDP for Region (AGEP)",
       x = "Region",
       y = "Pred. Mortality (%)") +
  theme_minimal() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, hjust = 1))

# Print plots (Optional)
print(pdp_age_agep_plot)
print(pdp_tte_agep_plot)
print(pdp_concomitant_agep_plot)
print(pdp_sex_agep_plot)
print(pdp_region_agep_plot)
```

Create the metric table:

```{r}
# Define the metrics
metrics_agep <- data.table(
  Metric = c("OOB Error/Brier Score", "AUC"),
  Value = c(round(oob_error_agep, 4), round(auc_value_agep, 4))
)

# Convert to a gt table for better visualization
metric_table_agep <- gt(metrics_agep) %>%
  tab_header(title = "Model Performance Metrics (AGEP)") %>%
  cols_label(
    Metric = "Metric",
    Value = "Value"
  ) %>%
  fmt_number(columns = Value, decimals = 4) %>%
  tab_options(
    table.font.size = px(14),
    heading.title.font.size = px(16),
    heading.title.font.weight = "bold"
  )

# Print the table
metric_table_agep
```

Now create the final combined plot

```{r}
# Combine three PDPs into one vertical column
pdp_column1_agep <- pdp_age_agep_plot / pdp_concomitant_agep_plot / pdp_tte_agep_plot + 
  plot_layout(ncol = 1)  # Stack vertically

# combine discrete PDPs into one vertical column
pdp_column2_agep <- pdp_region_agep_plot / pdp_sex_agep_plot 
  plot_layout(ncol = 1)

#Combine column 1 and 2 into the left column
left_column_agep <- pdp_column1_agep | pdp_column2_agep

# Combine heatmap and metric table into the second column
right_column_agep <- pVIVI_agep / metric_table_agep + 
  plot_layout(ncol = 1, heights = c(3, 1))  # Adjust heights (more space for heatmap)

# and the alternative
right_column_agep_alt <- pVimpBootstrapAGEP / metric_table_agep + 
  plot_layout(ncol = 1, heights = c(3, 1))  # Adjust heights

#Combine the figures
figS12_complete <- left_column_agep | right_column_agep
figS12_complete_alt <- left_column_agep | right_column_agep_alt
```
##PERMUTATION TESTING

This is to provide p-values for categorical variables in the mortality dataset. We'll write a function that will find pairwise permutation p-values for each pair of variables in a category.

```{r}
compute_pairwise_pdp_pvalues <- function(category_var, rf_model, data, num_permutations = 1000) {
  
  # Ensure category_var exists in data
  if (!category_var %in% colnames(data)) {
    stop(paste("Error: Variable", category_var, "not found in dataset."))
  }

  # Compute PDP for the categorical variable
  pdp_data <- partial(
    object = rf_model, 
    pred.var = category_var, 
    train = data,
    prob = TRUE
  )

  # Convert PDP data to a data frame
  pdp_df <- as.data.frame(pdp_data)
  colnames(pdp_df) <- c("Category", "Predicted_Prob")

  # Extract unique categories
  categories <- unique(pdp_df$Category)

  # Compute all unique pairwise category comparisons (upper-triangle only)
  category_pairs <- t(combn(categories, 2))  # Generate all category pairs

  # Compute observed pairwise differences (ALWAYS subtract Category_2 from Category_1)
  obs_diffs <- numeric(nrow(category_pairs))
  for (i in 1:nrow(category_pairs)) {
    cat1 <- category_pairs[i, 1]
    cat2 <- category_pairs[i, 2]
    
    mean_cat1 <- mean(pdp_df$Predicted_Prob[pdp_df$Category == cat1])
    mean_cat2 <- mean(pdp_df$Predicted_Prob[pdp_df$Category == cat2])
    
    obs_diffs[i] <- mean_cat1 - mean_cat2  # Ensure direction: Category_1 - Category_2
  }

  # Initialize matrix for storing permuted differences
  permuted_diffs <- matrix(NA, nrow = nrow(category_pairs), ncol = num_permutations)

  # Initialize progress bar
  pb <- txtProgressBar(min = 0, max = num_permutations, style = 3)

  # Permutation testing loop
  for (perm in 1:num_permutations) {
    setTxtProgressBar(pb, perm)  # Update progress bar

    # Shuffle category labels
    permuted_data <- pdp_df
    permuted_data$Category <- sample(permuted_data$Category)

    # Compute permuted pairwise differences
    for (i in 1:nrow(category_pairs)) {
      cat1 <- category_pairs[i, 1]
      cat2 <- category_pairs[i, 2]
      
      mean_cat1 <- mean(permuted_data$Predicted_Prob[permuted_data$Category == cat1])
      mean_cat2 <- mean(permuted_data$Predicted_Prob[permuted_data$Category == cat2])
      
      permuted_diffs[i, perm] <- mean_cat1 - mean_cat2  # Keep the same order
    }
  }

  close(pb)  # Close progress bar

  # Compute one-tailed p-values (alternative hypothesis: observed difference ≥ permuted difference)
  p_values <- numeric(nrow(category_pairs))
  perm_mean <- numeric(nrow(category_pairs))
  perm_sd <- numeric(nrow(category_pairs))

  for (i in 1:nrow(category_pairs)) {
    p_values[i] <- mean(permuted_diffs[i, ] >= obs_diffs[i])  # One-tailed test
    perm_mean[i] <- mean(permuted_diffs[i, ])  # Mean of permuted differences
    perm_sd[i] <- sd(permuted_diffs[i, ])  # Standard deviation of permuted differences
  }

  # Apply Benjamini-Hochberg correction
  p_values_bh <- p.adjust(p_values, method = "BH")

  # Return pairwise p-values as a data frame with separate columns for categories
  pairwise_p_values <- data.frame(
    Category_1 = category_pairs[, 1],
    Category_2 = category_pairs[, 2],
    Observed_Difference = obs_diffs,
    Permuted_Mean = perm_mean,
    Permuted_SD = perm_sd,
    P_Value = p_values,
    P_Value_BH = p_values_bh  # Benjamini-Hochberg corrected p-values
  )

  return(pairwise_p_values)
}
```

Remember that the RF is predicting living, not death, so the magnitude will be backwards. Now apply it:

```{r}
# Compute pairwise p-values for categorical variables
p_values_sex <- compute_pairwise_pdp_pvalues("sex", rf_final_interaction, clean_data_filtered, num_permutations = 10000)
p_values_region <- compute_pairwise_pdp_pvalues("region", rf_final_interaction, clean_data_filtered, num_permutations = 10000)

# Print results
print(p_values_sex)
print(p_values_region)
```

We'll need a different function for SCAR, since it's a series of booleans.

```{r}
compute_scar_pdp_pvalues <- function(rf_model, data, num_permutations = 1000) {
  
  # Define SCAR conditions
  scar_types <- c("SJSTEN", "DRESS", "AGEP", "GBFDE")
  
  # Initialize list to store PDP results
  pdp_scar_list <- list()

  # Compute PDPs for each SCAR type
  for (scar in scar_types) {
    
    pdp_result <- partial(
      object = rf_model,  
      pred.var = scar,    
      train = data,     
      prob = TRUE,        
      categorical = TRUE  
    )
    
    colnames(pdp_result) <- c("SCAR_Value", "Predicted_Prob")
    
    # Keep only SCAR = TRUE (1) results
    pdp_result <- pdp_result[pdp_result$SCAR_Value == 1, ]
    
    # Store SCAR type
    pdp_result$SCAR_Type <- scar
    
    pdp_scar_list[[scar]] <- pdp_result
  }
  
  # Combine PDP results into a single data frame
  pdp_data_scar <- do.call(rbind, pdp_scar_list)
  
  # Convert mortality probability (1 - yhat)
  pdp_data_scar$Predicted_Prob <- 1 - pdp_data_scar$Predicted_Prob
  
  # Compute observed pairwise differences
  scar_pairs <- t(combn(scar_types, 2))  # Generate pairwise SCAR combinations
  obs_diffs <- numeric(nrow(scar_pairs))

  for (i in 1:nrow(scar_pairs)) {
    scar1 <- scar_pairs[i, 1]
    scar2 <- scar_pairs[i, 2]
    
    mean_scar1 <- mean(pdp_data_scar$Predicted_Prob[pdp_data_scar$SCAR_Type == scar1])
    mean_scar2 <- mean(pdp_data_scar$Predicted_Prob[pdp_data_scar$SCAR_Type == scar2])
    
    obs_diffs[i] <- mean_scar1 - mean_scar2  # Difference between SCAR types
  }

  # Initialize permutation results matrix
  permuted_diffs <- matrix(NA, nrow = nrow(scar_pairs), ncol = num_permutations)

  # Progress bar
  pb <- txtProgressBar(min = 0, max = num_permutations, style = 3)

  # Permutation loop
  for (perm in 1:num_permutations) {
    setTxtProgressBar(pb, perm)  

    # Shuffle SCAR labels
    permuted_data <- pdp_data_scar
    permuted_data$SCAR_Type <- sample(permuted_data$SCAR_Type)

    # Compute permuted pairwise differences
    for (i in 1:nrow(scar_pairs)) {
      scar1 <- scar_pairs[i, 1]
      scar2 <- scar_pairs[i, 2]
      
      mean_scar1 <- mean(permuted_data$Predicted_Prob[permuted_data$SCAR_Type == scar1])
      mean_scar2 <- mean(permuted_data$Predicted_Prob[permuted_data$SCAR_Type == scar2])
      
      permuted_diffs[i, perm] <- mean_scar1 - mean_scar2  # Keep order
    }
  }

  close(pb)

  # Compute one-tailed p-values (obs_diff ≥ permuted)
  p_values <- numeric(nrow(scar_pairs))
  perm_mean <- numeric(nrow(scar_pairs))
  perm_sd <- numeric(nrow(scar_pairs))

  for (i in 1:nrow(scar_pairs)) {
    p_values[i] <- mean(permuted_diffs[i, ] >= obs_diffs[i])
    perm_mean[i] <- mean(permuted_diffs[i, ])
    perm_sd[i] <- sd(permuted_diffs[i, ])
  }

  # Apply Benjamini-Hochberg correction
  p_values_bh <- p.adjust(p_values, method = "BH")

  # Return pairwise p-values as a data frame
  pairwise_p_values <- data.frame(
    SCAR_1 = scar_pairs[, 1],
    SCAR_2 = scar_pairs[, 2],
    Observed_Difference = obs_diffs,
    Permuted_Mean = perm_mean,
    Permuted_SD = perm_sd,
    P_Value = p_values,
    P_Value_BH = p_values_bh  # Adjusted p-value
  )

  return(pairwise_p_values)
}
```

Let's run it!

```{r}
# Compute permutation p-values for SCAR-type PDP comparisons
p_values_scar <- compute_scar_pdp_pvalues(rf_final_interaction, clean_data_filtered, 1000)
print(p_values_scar)
```

Very good. Now we can permutation-test for the SCAR-specific RFs

```{r}
# Compute permutation p-values for SEX in each SCAR-specific RF
p_values_sex_sjsten <- compute_pairwise_pdp_pvalues("sex", rf_final_interaction_sjsten, clean_data_filtered_SJSTEN, 1000)
print(p_values_sex_sjsten)

p_values_sex_dress <- compute_pairwise_pdp_pvalues("sex", rf_final_interaction_dress, clean_data_filtered_DRESS, 1000)
print(p_values_sex_dress)

p_values_sex_agep <- compute_pairwise_pdp_pvalues("sex", rf_final_interaction_agep, clean_data_filtered_AGEP, 1000)
print(p_values_sex_agep)

# Compute permutation p-values for REGION in each SCAR-specific RF
p_values_region_sjsten <- compute_pairwise_pdp_pvalues("region", rf_final_interaction_sjsten, clean_data_filtered_SJSTEN, 1000)
print(p_values_region_sjsten)

p_values_region_dress <- compute_pairwise_pdp_pvalues("region", rf_final_interaction_dress, clean_data_filtered_DRESS, 1000)
print(p_values_region_dress)

p_values_region_agep <- compute_pairwise_pdp_pvalues("region", rf_final_interaction_agep, clean_data_filtered_AGEP, 1000)
print(p_values_region_agep)
```